### 2024-03-05 paper/main.tex

\documentclass[twocolumn]{aastex631}

% Packages
\usepackage{microtype}  % ALWAYS!
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{soul}

\definecolor{pink}{RGB}{232,132,161}
\definecolor{yellow}{RGB}{255,213,0}

\newcommand{\kc}[1]{\textcolor{yellow}{\textbf{kc: #1}} }
\newcommand\shadetext[2][]{%
  \setbox0=\hbox{{\special{pdf:literal 7 Tr }#2}}%
  \tikz[baseline=0]\path [#1] \pgfextra{\rlap{\copy0}} (0,-\dp0) rectangle (\wd0,\ht0);% 
  }
\newcommand{\gb}[1]{\shadetext[left color=blue, right color=red, middle color=lime, shading angle=45]{\textbf{g: #1}} }
% \newcommand{\ecite}[1]{\textcolor{pink}{\textbf{: #1}} }
% \newcommand{\e}[1]{\textcolor{yellow}{\textbf{: #1}} }

\newcommand{\remove}[1]{\textcolor{red}{#1}}
\newcommand{\add}[1]{\textcolor{green}{#1}}

\newcommand{\mlg}{\ensuremath{M_{\rm LG}}}
\newcommand{\mmto}{\ensuremath{M_{\rm M31}}}
\newcommand{\mmw}{\ensuremath{M_{\rm MW}}}
\newcommand{\vtan}{\ensuremath{v_\textrm{tan}}}
\newcommand{\vrad}{\ensuremath{v_\textrm{rad}}}
\newcommand{\ms}[1]{\ensuremath{M_{*{#1}}}}
\newcommand{\reflabel}[1]{\ensuremath{^{\mbox{\scriptsize{#1}}}}}
\newcommand{\scsep}{\ensuremath{\rm r_{sep}/r_{vir}}}
\newcommand{\scvel}{\ensuremath{\rm v_{rel}/v_{vir}}}
%\newcommand{\lsubcat}{\textit{Low Mass Subhalo Catalog}}
%\newcommand{\hsubcat}{\textit{High Mass Subhalo Catalog}}
\newcommand{\paircat}{\textit{Full Pair Catalog}}
\newcommand{\Rvir}{\ensuremath{\rm R_{vir}}}

% Style tweaks
% \renewcommand{\twocolumngrid}{\onecolumngrid}
% \setlength{\parindent}{1.1\baselineskip}
% \sloppy\sloppypar\raggedbottom\frenchspacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\shorttitle{Pair fractions in TNG100}
\shortauthors{Chamberlain et al.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{./}{../plots/}}
% \input{math_definitions.tex}

% Affiliations
\newcommand{\affuofa}{University of Arizona, 933 N. Cherry Ave,
    Tucson, AZ 85721, USA}

\newcommand{\affuofu}{Department of Astronomy, University of Utah, Salt Lake City, UT 84112, USA}

\begin{document}

\title{Orbits
}

\author[0000-0001-8765-8670]{Katie~Chamberlain}
\affiliation{\affuofa}

\author[0000-0002-9820-1219]{Ekta~Patel}
\thanks{Hubble Fellow}\affiliation{\affuofu}

\author[0000-0003-0715-2173]{Gurtina Besla}
\affiliation{\affuofa}

\author{others}

\begin{abstract}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{sec:intro}

$z=1.5$ corresponds to snapshot \#40 in the TNG sim
% Introduce RG15, Lotz, etc.

%%% OUTLINE
\section{Sample Data}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits

\subsection{Isolated Pairs}
We utilize the \paircat{} of isolated low and high mass pairs described in ~\citet{Chamberlain2024}. 
In summary, the \paircat{} is a collection of isolated subhalo pairs at each snapshot of the highest resolution run of the $(110.7\,\Mpc)^3$ volume IllustrisTNG simulation, TNG100-1 (hereafter TNG100).

Stellar masses were assigned to Abundance matching 
For the purposes of our study, we will only be using the

\subsubsection{brief recap of last paper}
\texttt{SUBLINK} algorithm~\citep{RG2015}

\subsubsection{Assumption of Isolation}
% justify stuff from last paper

\subsection{Mergers and Orbits}
\subsubsection{Calculation of merger fraction of total sample}
Plot #1: Merger fraction either before or after orbits
\subsubsection{Pulling Orbits}
Orbital data for each pair from the \paircat{} was collected by following each subhalo of the pair both forwards and backwards in time via the \texttt{SUBLINK} algorithm~\citep{RG2015} merger trees.

% interpolation
Plot #2: Orbits plot

\section{Results - The Mass Dependence of Merger Timescales}
% The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled by size of halo, OR that timescales for ow mass and high mass would be different by a lot for the same physical separation choices. (Maybe come up with a functional form to relate the two? )
\subsection{explain subset of data used for this section}
\subsection{Cumulative time}
Plot #3: Cumulative time example calculation plot
\subsection{"time til merger" for selected snapshots for LvH. }
Maybe also (distribution plot 2D hist? as function of selected separation)
\subsection{Cumulative distributions}
\subsubsection{Physical}
Plot #4: Cumulative distribution, low mass vs. high mass
\subsubsection{Scaled}
Plot #5: Cumulative distribution in scaled units, low mass vs. high mass
Plot #3-5?: the cumulative with spread for low vs. high mass

\section{Results - The Redshift Dependence of Merger Timescales}
% point: to show that the amount of time that pairs spend in the same physical separation bins changes by a significant amount between high and low redshift. In order to do pair fraction-merger rate studies appropriately, use Tobs(z), as suggested in Snyder 2017.  
\subsection{explain subset of data used for this section}
\subsection{explain how data in the plot was calculated}
\subsection{Redshift dependence}
Plot #6: Redshift dependence for low and high mass pairs
\subsection{Best fit redshift dependence?/scaling?}
Is $(1+z)^-2$ the best fit?

% \section{Discussion}
% \subsection{Merger Rates for Dwarf Pairs}
% % - What combination of values best reproduces the merger fraction acquired from the sims?
% % Do you need the scale of merger probability as in Ventou to get the match?
% % Vicente & Casteels 2014 comparison (pair fractions plot)
% \subsection{Pairs of Dwarfs That Do Not Merge}
% % can talk about the unmerged fraction merger rates here
% \subsection{Choosing Samples at Lower and Higher Redshift}
% % can also compare to samples to merger fractions and merger rates chosen at z=1 and z=2 (plots you have already)
% \subsection{Comparing with Merger Rates for Massive Galaxy Pairs}
% % compare to massive pairs in both obs window and merger rate
% % compare back to the assumption of isolation
\section{Summary and Conclusions}

% To do:
% -Sample: starting with FoF group cut and always show the total sample (merged and unmerged) since the unmerged sample is so small
% -Keep Fig 1: (two panels) add in a few more orbits to show the diversity and then have a panel to show one example where you shade everything below 100 kpc to show how the total time at S is computed
% -Keep Figure 2 as is 
% -Remove Figure 3
% -Figure 4: remove normalization of both samples so it's clear that the unmerged ones spend a lot of time at high separation and that they are a small percent of the fraction (put this in discussion or an appendix)
% -Remove Fig 5 and state conclusions in words along with the FoF group selection (most pairs have been in the same group for 1.5 Gyr) (can quote percentages from a cumulative version of the plot, or sum)
% -Figure 6: only show the *full* sample with the FoF group cut 
% check the half mass radius of the most massive dwarfs in your sample and justify 10 kpc by saying you want to be able to resolve twice that

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\columnwidth]{plots/2_pull-orbits/example-orbit.pdf}
    \caption{}
    % \label{fig:orbitex}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\columnwidth]{plots/1_merged-fraction/mergedfraction.pdf}
    \caption{}
    % \label{fig:merger-frac}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\columnwidth]{plots/3_timescale-analysis/fullsamples.pdf}
    \caption{}
    % \label{fig:total-time}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\columnwidth]{plots/4_timescales/timevssep_z.pdf}
    \caption{}
    % \label{fig:redshift-evo}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\columnwidth]{plots/5_define-beginning/countvsinfalltime.pdf}
    \caption{}
    % \label{fig:infall-time}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\columnwidth]{plots/4_timescales/timevssep_phys+co.pdf}
    \caption{}
    % \label{fig:phys-comoving}
\end{figure*}

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\columnwidth]{lotz.png}
    \caption{}
    % \label{fig:phys-comoving}
\end{figure*}

\bibliography{refs}{}
\bibliographystyle{aasjournal}

\end{document}

### 2024-03-05 paper/main.tex
Updates from Overleaf

@@
 Plot #1: Merger fraction either before or after orbits
+Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).
+
 \subsubsection{Pulling Orbits}
@@
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
     \centering
-    \includegraphics[width=\columnwidth]{plots/2_pull-orbits/example-orbit.pdf}
-    \caption{}
-    % \label{fig:orbitex}
-\end{figure*}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
+    \caption{The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
+At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
+The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.}
+    \label{fig:fmerge}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{plots/1_merged-fraction/mergedfraction.pdf}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/}
     \caption{}

### 2024-03-06 paper/main.tex

@@
 
+The infall snapshot is defined as ZZZ. 
+We will define the "First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo.
+
+Mergers are XXX 
+Non-mergers are YYY. 
+The merger snapshot is defined ALALAL.
+
+The orbit collection includes mergers and non-mergers, which
+
+Plot #2: Orbits plot
+Figure~\ref{fig:example-orbits} shows a collection of example orbits pulled from the merger trees of low-mass major pairs initially selected at $z=2$.
+
 % interpolation
-Plot #2: Orbits plot
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{plots/bet-on-it/}
-    \caption{}
-    % \label{fig:merger-frac}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/exorbits_physical.pdf}
+    \caption{Selection of example orbits of low-mass major pairs selected at $z=2$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
+    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group halo, while the dashed lines show the parts of the orbits that occur when the Group Halo of the primary and secondary are not the same.
+    (Top) Orbits of pairs that merge before $z=0$.
+Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
+    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued.
+These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
+    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the $\sim2Gyr$ after the end of the simulation. 
+    }
+    \label{fig:example-orbits}
 \end{figure*}

### 2024-03-09 paper/main.tex
Updates from Overleaf

@@
 \newcommand{\scvel}{\ensuremath{\rm v_{rel}/v_{vir}}}
-%\newcommand{\lsubcat}{\textit{Low Mass Subhalo Catalog}}
-%\newcommand{\hsubcat}{\textit{High Mass Subhalo Catalog}}
+
+\newcommand{\rsep}{\ensuremath{\rm r_{sep}}}
 \newcommand{\paircat}{\textit{Full Pair Catalog}}
 \newcommand{\Rvir}{\ensuremath{\rm R_{vir}}}
+\newcommand{\Rphys}{\ensuremath{\rm R_{phys}}}
+\newcommand{\Rsc}{\ensuremath{\rm R_{sc}}}
+\newcommand{\rsep}{\ensuremath{\rm r_{sep}}}
 
@@
 \graphicspath{{./}{../plots/}}
-% \input{math_definitions.tex}
+\input{math_definitions.tex}
 
@@
 
-% interpolation
+\subsubsection{Interpolation}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
+    \caption{The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
+At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
+The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.}
+    \label{fig:fmerge}
+\end{figure}
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{plots/bet-on-it/exorbits_physical.pdf}
+    \caption{Selection of example orbits of low-mass major pairs selected at $z=2$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
+    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group halo, while the dashed lines show the parts of the orbits that occur when the Group Halo of the primary and secondary are not the same.
+    (Top) Orbits of pairs that merge before $z=0$.
+Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
+    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued.
+These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
+    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the $\sim2Gyr$ after the end of the simulation. 
+    }
+    \label{fig:example-orbits}
+\end{figure*}
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{plots/bet-on-it/4}
+    \caption{Selection of example orbits of low-mass major pairs selected at $z=2$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
+    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group halo, while the dashed lines show the parts of the orbits that occur when the Group Halo of the primary and secondary are not the same.
+    (Top) Orbits of pairs that merge before $z=0$.
+Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
+    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued.
+These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
+    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the $\sim2Gyr$ after the end of the simulation. 
+    }
+    \label{fig:example-orbits}
+\end{figure*}
+
+% ###########################################################################################
+% Results II 
+% ###########################################################################################
 
@@
 
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi_int.pdf}
+    \caption{\kc{reverse the direction of z} 
+    The median (solid) and 1st-3rd quartile spread (shaded regions) of the cumulative time that isolated low-mass pairs (left) and high-mass pairs (right) spend with separations \rsep{} between $10\kpc$ and \Rphys{}.
+    The Merging Pairs sample (dark green and dark pink) selected at $z=(1,1.5,2)$ merge before $z=0$, while the All Pairs sample (light green and light pink) include all the pairs from the \paircat{} (mergers and non-mergers).
+    The low-mass (high-mass) merging pairs represent XX\%(XX\%) of the low-mass (high-mass) pair sample at $z=1,1.5, \mbox{and } 2$ respectively. 
+    The sample of All Pairs spend more time at higher separations than the merging pairs, as expected since the non-merging pairs are not likely to have very low separations that would result in mergers prior to $z=0$.
+    For a more direct comparison between the low-mass and high-mass pairs, see Fig.~\ref{fig:phys-vs-scaled}. 
+    }
+    \label{fig:low-vs-high}
+\end{figure*}
+
+\begin{figure*}[htb]
+    \centering
+    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT_int.pdf}
+    \caption{\kc{Add legend to plot, and I think reverse the direction of z} 
+    The median and 1st-3rd quartile spread of the cumulative time that isolated merging pairs, selected at $z=(1,1.5,2)$, spend with separations \rsep{} greater than $10\kpc$ and less than either \Rphys{} or \Rsc{}.
+    %
+    (Top) The median total time spent by isolated merging pairs with separations between $10\kpc$ and \Rphys{}.
+    Low-mass pairs spend a longer amount of time within $\sim200\kpc$ than high-mass pairs at each of these redshifts. 
+    However, high-mass galaxies spend more time at higher separations of $\rsep>200-300\kpc$.
+This is because the elapsed time is measured starting at the infall snapshot of the pair, and high-mass halos have larger radii, the high-mass halo orbits have a broader distribution of separations that extend to larger separations than those of the low-mass pairs.
+    (Bottom) The median total time spent by pairs with separations greater than $\rsep>10\kpc$ and less than a given fraction of the virial radius ($\rsep/\Rvir<\Rsc$) of the pair, which scales with mass and redshift.
+    High-mass pairs spend more time within the same scaled separation than low-mass pairs.
+For example, a high-mass pair selected from the $z=1.5$ sample will spend \kc{XXGyr} within \kc{$1\Rsc\simXX\kpc/$}, while a low-mass pair from the same sample will spend \kc{XXGyr} within \kc{$1\Rsc\simXX\kpc$}.
+    Low-mass and high-mass pairs have similar cumulative time profiles as a function of scaled separation, with a roughly constant offset for all scaled separations. 
+    The high-mass pairs spend a median of \kc{$XX\Gyr$} more than low-mass pairs at every scaled separation at $z=1$ and $z=1.5$, and a median of \kc{$XX\Gyr$} more $z=2$.
+    }
+    \label{fig:phys-vs-scaled}
+\end{figure*}
+
+% ###########################################################################################
+% Results II 
+% ###########################################################################################
 \section{Results - The Redshift Dependence of Merger Timescales}
@@
 
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
-    \caption{The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
-At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
-The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.}
-    \label{fig:fmerge}
-\end{figure}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/exorbits_physical.pdf}
-    \caption{Selection of example orbits of low-mass major pairs selected at $z=2$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
-    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group halo, while the dashed lines show the parts of the orbits that occur when the Group Halo of the primary and secondary are not the same.
-    (Top) Orbits of pairs that merge before $z=0$.
-Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
-    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued.
-These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
-The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
-    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the $\sim2Gyr$ after the end of the simulation. 
-    }
-    \label{fig:example-orbits}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\columnwidth]{plots/3_timescale-analysis/fullsamples.pdf}
-    \caption{}
-    % \label{fig:total-time}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\columnwidth]{plots/4_timescales/timevssep_z.pdf}
-    \caption{}
-    % \label{fig:redshift-evo}
-\end{figure*}
-
 \begin{figure*}[htb]
@@
 
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\columnwidth]{plots/4_timescales/timevssep_phys+co.pdf}
-    \caption{}
-    % \label{fig:phys-comoving}
-\end{figure*}
+% \begin{figure*}[htb]
+%     \centering
+%     \includegraphics[width=\columnwidth]{plots/4_timescales/timevssep_phys+co.pdf}
+%     \caption{}
+%     % \label{fig:phys-comoving}
+% \end{figure*}
 
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\columnwidth]{lotz.png}
-    \caption{}
-    % \label{fig:phys-comoving}
-\end{figure*}
+% \begin{figure*}[htb]
+%     \centering
+%     \includegraphics[width=\columnwidth]{lotz.png}
+%     \caption{}
+%     % \label{fig:phys-comoving}
+% \end{figure*}
 

### 2024-03-09 paper/main.tex

@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/4}
-    \caption{Selection of example orbits of low-mass major pairs selected at $z=2$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
-    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group halo, while the dashed lines show the parts of the orbits that occur when the Group Halo of the primary and secondary are not the same.
-    (Top) Orbits of pairs that merge before $z=0$.
-Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
-    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued.
-These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
-The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
-    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the $\sim2Gyr$ after the end of the simulation. 
+    \includegraphics[width=\textwidth]{plots/bet-on-it/4_examplecombo.png}
+    \caption{ 
     }
-    \label{fig:example-orbits}
+    \label{fig:interpolation}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi_int.pdf}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi.pdf}
     \caption{\kc{reverse the direction of z} 
@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT_int.pdf}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT.pdf}
     \caption{\kc{Add legend to plot, and I think reverse the direction of z} 
@@
 
-\begin{figure*}[htb]
+% \section{Discussion}
+
+\appendix
+\begin{figure}[htb]
     \centering
-    \includegraphics[width=\columnwidth]{plots/5_define-beginning/countvsinfalltime.pdf}
+    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
+\end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
     \caption{}
-    % \label{fig:infall-time}
-\end{figure*}
+\end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
+\end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_number_pairs.png}
+\end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_beforemerger.png}
+\end{figure}
+\begin{figure}[htb]
+    \centering
+    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_untilmerger_phys.png}
+\end{figure}
 

### 2024-03-09 paper/main.tex

@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/exorbits_physical.pdf}
+    \includegraphics[width=0.5\textwidth]{plots/bet-on-it/exorbits_physical.pdf}
     \caption{Selection of example orbits of low-mass major pairs selected at $z=2$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi.pdf}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi.png}
     \caption{\kc{reverse the direction of z} 
@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT.pdf}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT.png}
     \caption{\kc{Add legend to plot, and I think reverse the direction of z} 

### 2024-03-09 paper/main.tex
Updates from Overleaf

@@
 
-\newcommand{\rsep}{\ensuremath{\rm r_{sep}}}
 \newcommand{\paircat}{\textit{Full Pair Catalog}}
@@
 \subsubsection{Calculation of merger fraction of total sample}
-Plot #1: Merger fraction either before or after orbits
+Plot \#1: Merger fraction either before or after orbits
 Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).
@@
 
-Plot #2: Orbits plot
+Plot \#2: Orbits plot
 Figure~\ref{fig:example-orbits} shows a collection of example orbits pulled from the merger trees of low-mass major pairs initially selected at $z=2$.
+
+ Non-mergers: These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
+
+Mergers:  Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
+
+Flybys:
 
@@
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
     \centering
-    \includegraphics[width=0.5\textwidth]{plots/bet-on-it/exorbits_physical.pdf}
-    \caption{Selection of example orbits of low-mass major pairs selected at $z=2$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
-    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group halo, while the dashed lines show the parts of the orbits that occur when the Group Halo of the primary and secondary are not the same.
-    (Top) Orbits of pairs that merge before $z=0$.
-Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
+    \includegraphics[width=\columnwidth, trim={0.4cm 1.75cm 1cm 3.5cm}, clip]{plots/bet-on-it/5_exampleorbits.png}
+    \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
+The vertical grey line marks $z=1.5$ at a Lookback Time of $9.5\,\Gyr$.
+    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group, while the dashed lines show where they are not within the same FoF group. 
+    (Top) Orbits of pairs that merge before $z=0$. Pairs chosen at $z=1.5$ on average merge at \kc{$z=XX$/within XX Gyrs}.
     (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
-These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
-The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
-    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the $\sim2Gyr$ after the end of the simulation. 
+    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation). 
     }
     \label{fig:example-orbits}
-\end{figure*}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/4_examplecombo.png}
-    \caption{ 
-    }
+    \includegraphics[width=0.8\textwidth]{plots/bet-on-it/4_examplecombo_unlabeled.png}
+    \caption{ Caption    }
     \label{fig:interpolation}
@@
 \subsection{Cumulative time}
-Plot #3: Cumulative time example calculation plot
+Plot \#3: Cumulative time example calculation plot
 \subsection{"time til merger" for selected snapshots for LvH. }
@@
 \subsubsection{Physical}
-Plot #4: Cumulative distribution, low mass vs. high mass
+Plot \#4: Cumulative distribution, low mass vs. high mass
 \subsubsection{Scaled}
-Plot #5: Cumulative distribution in scaled units, low mass vs. high mass
-Plot #3-5?: the cumulative with spread for low vs. high mass
+Plot \#5: Cumulative distribution in scaled units, low mass vs. high mass
+Plot \#3-5?: the cumulative with spread for low vs. high mass
 
@@
     High-mass pairs spend more time within the same scaled separation than low-mass pairs.
-For example, a high-mass pair selected from the $z=1.5$ sample will spend \kc{XXGyr} within \kc{$1\Rsc\simXX\kpc/$}, while a low-mass pair from the same sample will spend \kc{XXGyr} within \kc{$1\Rsc\simXX\kpc$}.
+For example, a high-mass pair selected from the $z=1.5$ sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}, while a low-mass pair from the same sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}.
     Low-mass and high-mass pairs have similar cumulative time profiles as a function of scaled separation, with a roughly constant offset for all scaled separations. 
@@
 \subsection{Redshift dependence}
-Plot #6: Redshift dependence for low and high mass pairs
+Plot \#6: Redshift dependence for low and high mass pairs
 \subsection{Best fit redshift dependence?/scaling?}
@@
 
+** 
 \appendix

### 2024-03-14 paper/main.tex
Updates from Overleaf

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\shorttitle{Pair fractions in TNG100}
+\shorttitle{Merger Timescales in TNG100}
 \shortauthors{Chamberlain et al.}
@@
 
-\title{Orbits
+\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxies Across Cosmic Time
 }
@@
 \begin{abstract}
-
+% Statement
+Evolutionary differences between low-mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high-mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) galaxy pairs have been studied in detail for the last few decades, with studies finding differences in star formation rates, gas fractions, and dark matter halo profiles.
+% Problem 
+However, many studies continue to use equivalent separation criteria to select galaxy pairs for pair fraction, merger fraction, and merger rate studies, though it is unclear that such a selection permits an equitable comparison between galaxies of different mass, nor as a function of redshift.
+% Our solution
+We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass galaxy pairs as a function of cosmic time. 
+% What are the timescales? 
+We show that separation limits that vary with the mass and redshift of the system, such as scaling by the virial radius of the host halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$), lead to equivalent merger timescales for Low- and High-mass systems at all redshifts.
+Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$) leads to timescales that differ between low- and high-mass pairs.
+% Why it matters 
+Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates must be done with self-consistent separation criteria that scales as a function of mass and redshift of the system.
+As a result, applying the same merger timescales to different mass systems will lead to a bias that systematically (over/under) predicts low-mass galaxy merger rates. 
 \end{abstract}
@@
 \begin{figure}[htb]
-    \centering
+    \begin{center}
     \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
@@
     \label{fig:fmerge}
+    \end{center}
 \end{figure}
@@
 \begin{figure}[htb]
-    \centering
+    \begin{center}
     \includegraphics[width=\columnwidth, trim={0.4cm 1.75cm 1cm 3.5cm}, clip]{plots/bet-on-it/5_exampleorbits.png}
@@
     \label{fig:example-orbits}
+    \end{center}
 \end{figure}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/bet-on-it/4_examplecombo_unlabeled.png}
-    \caption{ Caption    }
-    \label{fig:interpolation}
-\end{figure*}
 
@@
 
-\section{Results - The Mass Dependence of Merger Timescales}
-% The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled by size of halo, OR that timescales for ow mass and high mass would be different by a lot for the same physical separation choices.
-(Maybe come up with a functional form to relate the two?
-)
-\subsection{explain subset of data used for this section}
-\subsection{Cumulative time}
-Plot \#3: Cumulative time example calculation plot
-\subsection{"time til merger" for selected snapshots for LvH. }
+\section{Results - The Mass and Redshift Dependence of Merger Timescales}
+% The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled with mass and redshift.
+\subsection{"time til merger" for selected snapshots for LvH.}
 Maybe also (distribution plot 2D hist? as function of selected separation)
@@
 \subsubsection{Physical}
-Plot \#4: Cumulative distribution, low mass vs. high mass
 \subsubsection{Scaled}
-Plot \#5: Cumulative distribution in scaled units, low mass vs. high mass
-Plot \#3-5?: the cumulative with spread for low vs. high mass
 
+% \begin{figure*}[htb]
+%     \centering
+%     \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi.png}
+%     \caption{\kc{reverse the direction of z} 
+%     The median (solid) and 1st-3rd quartile spread (shaded regions) of the cumulative time that isolated low-mass pairs (left) and high-mass pairs (right) spend with separations \rsep{} between $10\kpc$ and \Rphys{}.
+%     The Merging Pairs sample (dark green and dark pink) selected at $z=(1,1.5,2)$ merge before $z=0$, while the All Pairs sample (light green and light pink) include all the pairs from the \paircat{} (mergers and non-mergers).
+%     The low-mass (high-mass) merging pairs represent XX\%(XX\%) of the low-mass (high-mass) pair sample at $z=1,1.5, \mbox{and } 2$ respectively. 
+%     The sample of All Pairs spend more time at higher separations than the merging pairs, as expected since the non-merging pairs are not likely to have very low separations that would result in mergers prior to $z=0$.
+%     For a more direct comparison between the low-mass and high-mass pairs, see Fig.~\ref{fig:phys-vs-scaled}. 
+%     }
+%     \label{fig:low-vs-high}
+% \end{figure*}
+
+% \begin{figure*}[htb]
+%     \centering
+%     \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT.png}
+%     \caption{\kc{Add legend to plot, and I think reverse the direction of z} 
+%     The median and 1st-3rd quartile spread of the cumulative time that isolated merging pairs, selected at $z=(1,1.5,2)$, spend with separations \rsep{} greater than $10\kpc$ and less than either \Rphys{} or \Rsc{}.
+%     %
+%     (Top) The median total time spent by isolated merging pairs with separations between $10\kpc$ and \Rphys{}.
+%     Low-mass pairs spend a longer amount of time within $\sim200\kpc$ than high-mass pairs at each of these redshifts. 
+%     However, high-mass galaxies spend more time at higher separations of $\rsep>200-300\kpc$.
+This is because the elapsed time is measured starting at the infall snapshot of the pair, and high-mass halos have larger radii, the high-mass halo orbits have a broader distribution of separations that extend to larger separations than those of the low-mass pairs.
+%     (Bottom) The median total time spent by pairs with separations greater than $\rsep>10\kpc$ and less than a given fraction of the virial radius ($\rsep/\Rvir<\Rsc$) of the pair, which scales with mass and redshift.
+%     High-mass pairs spend more time within the same scaled separation than low-mass pairs.
+For example, a high-mass pair selected from the $z=1.5$ sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}, while a low-mass pair from the same sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}.
+%     Low-mass and high-mass pairs have similar cumulative time profiles as a function of scaled separation, with a roughly constant offset for all scaled separations. 
+%     The high-mass pairs spend a median of \kc{$XX\Gyr$} more than low-mass pairs at every scaled separation at $z=1$ and $z=1.5$, and a median of \kc{$XX\Gyr$} more $z=2$.
+%     }
+%     \label{fig:phys-vs-scaled}
+% \end{figure*}
+
+% ###########################################################################################
+% Results II 
+% ###########################################################################################
+
+% \begin{figure*}[htb]
+%     \centering
+%     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsz.png}
+%     \caption{}
+%     % \label{fig:low-vs-high}
+% \end{figure*}
+% \begin{figure*}[htb]
+%     \centering
+%     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsz-2d.png}
+%     \caption{}
+%     % \label{fig:low-vs-high}
+% \end{figure*}
 \begin{figure*}[htb]
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi.png}
-    \caption{\kc{reverse the direction of z} 
-    The median (solid) and 1st-3rd quartile spread (shaded regions) of the cumulative time that isolated low-mass pairs (left) and high-mass pairs (right) spend with separations \rsep{} between $10\kpc$ and \Rphys{}.
-    The Merging Pairs sample (dark green and dark pink) selected at $z=(1,1.5,2)$ merge before $z=0$, while the All Pairs sample (light green and light pink) include all the pairs from the \paircat{} (mergers and non-mergers).
-    The low-mass (high-mass) merging pairs represent XX\%(XX\%) of the low-mass (high-mass) pair sample at $z=1,1.5, \mbox{and } 2$ respectively. 
-    The sample of All Pairs spend more time at higher separations than the merging pairs, as expected since the non-merging pairs are not likely to have very low separations that would result in mergers prior to $z=0$.
-    For a more direct comparison between the low-mass and high-mass pairs, see Fig.~\ref{fig:phys-vs-scaled}. 
-    }
-    \label{fig:low-vs-high}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger_full_fit.png}
+    \caption{}
+    % \label{fig:low-vs-high}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT.png}
-    \caption{\kc{Add legend to plot, and I think reverse the direction of z} 
-    The median and 1st-3rd quartile spread of the cumulative time that isolated merging pairs, selected at $z=(1,1.5,2)$, spend with separations \rsep{} greater than $10\kpc$ and less than either \Rphys{} or \Rsc{}.
-    %
-    (Top) The median total time spent by isolated merging pairs with separations between $10\kpc$ and \Rphys{}.
-    Low-mass pairs spend a longer amount of time within $\sim200\kpc$ than high-mass pairs at each of these redshifts. 
-    However, high-mass galaxies spend more time at higher separations of $\rsep>200-300\kpc$.
-This is because the elapsed time is measured starting at the infall snapshot of the pair, and high-mass halos have larger radii, the high-mass halo orbits have a broader distribution of separations that extend to larger separations than those of the low-mass pairs.
-    (Bottom) The median total time spent by pairs with separations greater than $\rsep>10\kpc$ and less than a given fraction of the virial radius ($\rsep/\Rvir<\Rsc$) of the pair, which scales with mass and redshift.
-    High-mass pairs spend more time within the same scaled separation than low-mass pairs.
-For example, a high-mass pair selected from the $z=1.5$ sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}, while a low-mass pair from the same sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}.
-    Low-mass and high-mass pairs have similar cumulative time profiles as a function of scaled separation, with a roughly constant offset for all scaled separations. 
-    The high-mass pairs spend a median of \kc{$XX\Gyr$} more than low-mass pairs at every scaled separation at $z=1$ and $z=1.5$, and a median of \kc{$XX\Gyr$} more $z=2$.
-    }
-    \label{fig:phys-vs-scaled}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger.png}
+    \caption{}
+    % \label{fig:low-vs-high}
 \end{figure*}
 
-% ###########################################################################################
-% Results II 
-% ###########################################################################################
-\section{Results - The Redshift Dependence of Merger Timescales}
-% point: to show that the amount of time that pairs spend in the same physical separation bins changes by a significant amount between high and low redshift.
-In order to do pair fraction-merger rate studies appropriately, use Tobs(z), as suggested in Snyder 2017.
-\subsection{explain subset of data used for this section}
-\subsection{explain how data in the plot was calculated}
-\subsection{Redshift dependence}
-Plot \#6: Redshift dependence for low and high mass pairs
-\subsection{Best fit redshift dependence?/scaling?}
-Is $(1+z)^-2$ the best fit?
-
-% \section{Discussion}
+\section{Discussion}
 % \subsection{Merger Rates for Dwarf Pairs}
@@
 % % compare back to the assumption of isolation
+
 \section{Summary and Conclusions}
@@
 \appendix
-\begin{figure}[htb]
+\begin{figure*}[htb]
     \centering
-    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
-\end{figure}
-\begin{figure}[htb]
+    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
+    \caption{}
+    % \label{fig:low-vs-high}
+\end{figure*}
+
+\begin{figure*}[htb]
     \centering
-    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
     \caption{}
-\end{figure}
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
-\end{figure}
+    % \label{fig:low-vs-high}
+\end{figure*}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
+% \end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
+%     \caption{}
+% \end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
+% \end{figure}
 \begin{figure}[htb]
@@
 \end{figure}
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_beforemerger.png}
-\end{figure}
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_untilmerger_phys.png}
-\end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_beforemerger.png}
+% \end{figure}
+% \begin{figure}[htb]
+%     \centering
+%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_untilmerger_phys.png}
+% \end{figure}
 

### 2024-03-17 paper/main.tex
mostly finished the pair sample part of methodology - next up pair sample/calculation specifics

@@
 % Our solution
-We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass galaxy pairs as a function of cosmic time. 
+We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass major pairs as a function of cosmic time. 
 % What are the timescales? 
-We show that separation limits that vary with the mass and redshift of the system, such as scaling by the virial radius of the host halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$), lead to equivalent merger timescales for Low- and High-mass systems at all redshifts.
+We show that separation limits that vary with the mass and redshift of the system, such as scaling by the virial radius of the host halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$), lead to equivalent merger timescales for low- and high-mass systems at all redshifts.
 Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$) leads to timescales that differ between low- and high-mass pairs.
@@
 
-$z=1.5$ corresponds to snapshot \#40 in the TNG sim
+% $z=1.5$ corresponds to snapshot \#40 in the TNG sim
 % Introduce RG15, Lotz, etc.
@@
 %%% OUTLINE
-\section{Sample Data}
+\section{Methodology}
 %recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
+We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-XX$.
 
-\subsection{Isolated Pairs}
-We utilize the \paircat{} of isolated low and high mass pairs described in ~\citet{Chamberlain2024}. 
-In summary, the \paircat{} is a collection of isolated subhalo pairs at each snapshot of the highest resolution run of the $(110.7\,\Mpc)^3$ volume IllustrisTNG simulation, TNG100-1 (hereafter TNG100).
+Our sample consists of major low-mass and high-mass pairs that are isolated but physically associated.
+From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.
 
-Stellar masses were assigned to Abundance matching 
-For the purposes of our study, we will only be using the
+\subsection{Pair orbit sample}
+We begin with a subset of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
+A brief version of the selection routine is described here.
 
-\subsubsection{brief recap of last paper}
-\texttt{SUBLINK} algorithm~\citep{RG2015}
+At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive halos (by stellar mass) from FoF groups with virial mass\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass.
+This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Brynorman1998} and $\rho_c$ is the critical density of the universe at the time calculated.
+The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.}
+\begin{align*}
+        \mbox{\textbf{low mass:}}&\,\rm M_{G} = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
+        \mbox{\textbf{high mass:}}&\, \rm M_{G}=10^{12}- 6.5\times10^{12}\,\Msun.
+\end{align*}
+Utilizing the most massive subhalos from the same FoF group ensures that the pairs are isolated from massive nearby systems that may perturb the dynamical state of the pair.
 
-\subsubsection{Assumption of Isolation}
-% justify stuff from last paper
+We require that subhalos that constitute a pair meet a minimum subhalo mass criteria of 
+\begin{equation*}
+    \mbox{\textbf{minimum subhalo mass:}}\,
+    \Mhalo > 1\times10^{9}\Msun.
+\end{equation*}
+at the snapshot of consideration. 
+For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to trace their mass history~\citep{RG2015}, and find the peak halo mass of each subhalo.
 
-\subsection{Mergers and Orbits}
-\subsubsection{Calculation of merger fraction of total sample}
-Plot \#1: Merger fraction either before or after orbits
-Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).
+Stellar masses are then assigned to each subhalo in the FoF group using the Abundance Matching prescription of \citet{Moster2013}. 
+The peak halo mass and current snapshot are used to calculate the stellar mass of each subhalo, so that the abundance matching prescription is of the form $\ms{}=f(\Mpeak,z_{\mathrm{snap}})$.
 
-\subsubsection{Pulling Orbits}
-Orbital data for each pair from the \paircat{} was collected by following each subhalo of the pair both forwards and backwards in time via the \texttt{SUBLINK} algorithm~\citep{RG2015} merger trees.
+In \citet{Chamberlain2024}, the abundance matching prescription is sampled 1000 times for each subhalo to account for the spread in the abundance matching prescription. 
+However, for the purposes of the present study, we only use the stellar masses given by the median of the abundance matching relationship. 
+This choice allows us to more thoroughly consider the population under study, and also permits a roughly 1:1 monotonic relationship between stellar mass and peak halo mass, so that taking subsamples of the lowest stellar mass objects will result in a subsample with lower peak halo mass as well.
+% is this a good enough reason?
+% else, maybe spread of number of pairs is so small, that we expect the noise of the timescale analysis to be dominated by the snapshot spacing anyways?
 
-The infall snapshot is defined as ZZZ. 
-We will define the "First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo.
+Primary halos are defined as the subhalo with the highest stellar mass ($M_{*1}$) in their FoF group, and secondaries are defined as the second most massive subhalo with stellar mass $M_{*2}$. 
+Our sample of major pairs then consists of all pairs of primary and secondary halos with 
+\begin{align*} 
+\mbox{\textbf{low mass primaries:}}&\, 10^{8}< \rm M_{*1} < 5\times10^{9} \Msun \\ 
+\mbox{\textbf{high mass primaries:}}&\, 5\times 10^{9}< \rm M_{*1} < 10^{11} \Msun\\
+\mbox{\textbf{stellar mass ratio:}}&\,      
+    M_{*2}/M_{*1} > 1/4.
+\end{align*}
+A primary or secondary halo can only be a member of one single pair at a given snapshot, such that a collection of $N$ pairs consists of $N$ unique primaries and $N$ unique secondaries.
 
-Mergers are XXX 
-Non-mergers are YYY. 
-The merger snapshot is defined ALALAL.
+% i am removing the 10kpc component of this because I think it'd just be easier to handle those cases in the 
+% Finally, we will only consider pairs that have physical separations of $\rsep>10\,\kpc$ at the snapshot of selection. 
+% The physical separation cut does not vary with redshift, and is the same for both low-mass and high-mass pairs.
 
-The orbit collection includes mergers and non-mergers, which
+The base sample used for this analysis then consists of the set of all low-mass and high-mass major pairs from each snapshot of the TNG100 simulation.
 
-Plot \#2: Orbits plot
-Figure~\ref{fig:example-orbits} shows a collection of example orbits pulled from the merger trees of low-mass major pairs initially selected at $z=2$.
+\subsubsection{Mergers}
+% Plot \#1: Merger fraction either before or after orbits
+% Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).
 
- Non-mergers: These orbits can have a variety of orbital periods and thus, a number of different pericentric passages.
-The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbits has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
+The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} catalogs, which contain information about every FoF group and subhalo at each snapshot.
+The \texttt{SUBLINK} catalogs track subhalos from one snapshot to the next, and thus enable us to track subhalos both backwards and forwards in time from any given snapshot.
 
-Mergers:  Pairs chosen at $z=2$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between 0.5-8Gyr.
+We utilize the \texttt{RootDescendantID} and \texttt{DescendantID} fields of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$), and to determine when the merger occurs.
+The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
+The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshot, if it has one. 
+% By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.  
+The \texttt{RootDescendantID} field provides the SubhaloID of the root subhalo, which is the latest subhalo in the simulation that is a descendent of the subhalo.
 
-Flybys:
+We define the set of merging pairs as the collection of all pairs where the primary and secondary halo have the same \texttt{RootDescendantID}, meaning that the branches of their merger trees have the same `root', or common descendent.
+If a given pair has two distinct \texttt{RootDescendantID} values, the pair is defined as a `non-merger.'
 
-\subsubsection{Interpolation}
-\begin{figure}[htb]
-    \begin{center}
-    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
-    \caption{The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
-At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
-The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.}
-    \label{fig:fmerge}
-    \end{center}
-\end{figure}
+For each merging pair, we define the `merger snapshot' as the snapshot that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
+For example, if the primary and secondary have different \texttt{DescendantIDs} for snapshots 40-45, but have the same \texttt{DescendantID} at snapshot 46, then the merger must take place some time between snapshot 46 and 47.
+We take snapshot 47 to be the merger snapshot.
+
+\subsubsection{Orbits} 
+We calculate orbits for all mergers and non-mergers in our base sample. 
+An orbit for a single pair is defined to be the physical separation between the primary and secondary halo as a function of redshift or Lookback Time.
+
+A given pair from the base sample at snapshot $S_n$, which by definition passes all selection criteria at $S_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.
+We track the positions of both the primary and secondary subhalo at each snapshot and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
+In cases where the primary or secondary does not have a defined position at a given snapshot, we set the separation at that snapshot to NaN.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs.
+The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be computed contiguously for a pair on .
+See Sec.~3 in ~\citet{RG2015} for more details.}
+
+In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot. We use the definition of scaled separation from~\cite{Chamberlain2024}, given as
+\begin{equation}
+    \rm \Rsc = \frac{\rsep}{\Rvir}
+\end{equation}
+where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the primary's FoF group, which reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
+The scaled radius $\Rsc$ is by construction a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs, or for halo growth over time.
+
+We collect the following data for each pair at each snapshot where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group \texttt{Group\_R\_TopHat200}, the physical separation (in $\kpc$), and the scaled separation (dimensionless).
+
+\subsubsection{"First Infall"}
+The orbit of a pair may be technically defined at very early times, well before the primary and secondary are physically associated. 
+However, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group.
+Specifically, we will consider only the "Post-Infall" part of each pair's orbit.
+
+We define the snapshot of "First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo, and "Post-Infall" as the set of snapshots including the snapshot of "First Infall" and all snapshots thereafter.
+The post-infall orbit definition is slightly more robust for considering the full interaction timescales of merging pairs than only considering the orbit when the FoF group of the primary and secondary are the same.
+This is because there are cases of pair interactions where the secondary will enter the FoF group of the primary, have a pericentric passage, then leave the FoF group for a few gigayears before re-entering the primary FoF group and eventually merging with the primary.
+
+An example of the orbit of a merging pair where a "common group" criteria may differ from a "post-infall" can be seen in Figure~\ref{fig:example-orbits}, which shows the variety of orbits that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$.
+The top panel shows the orbits of 5 pairs that merge, where solid lines show the parts of the orbit that occur while the primary and secondary share a common FoF group, and dashed lines show the parts of the merger where the FoF group of the secondary is different than that of the primary.
+The "post-infall" part of the orbit occurs after the first snapshot where the solid lines begins.
+Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between $\sim1-10\Gyr$.
+
+The middle panel shows the orbits for 3 pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a couple Gyr past the end of the simulation, and are thus "Potential Future Mergers".
+These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbit has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
+The bottom panel shows the orbits of "Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
+Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.
 
@@
 \end{figure}
+
+\subsubsection{Uniqueness of pairs}
+Since our entire pair sample consists of pairs chosen at separate snapshots, and since a single pair could pass our selection criteria at many snapshots, we keep only a single orbital instance of each pair.
+
+To distinguish pairs from one snapshot to another, each pair is assigned a pairkey while constructing their orbit, which is created by combining the primary's and secondary's earliest SubhaloID from the Sublink catalogs, which will be unique for each pair of halos.
+After each pair is assigned a unique pairkey, we save the orbit as a function of time only once for a pair that otherwise would have been double/multi-counted in our orbit catalog.
+
+The total number of low-mass pairs selected at each snapshot (given by $\Sigma_{S_n} n_p$) the sum of the number of pairs that pass the selection criteria at each snapshot is XXX, while the total number of \textit{unique} pairs is XXX.
+The total number of high-mass pairs is XXX, while the total number of unique pairs is XXX.
+
+Note that a single subhalo may be a member of many different pairs, but will only have one unique orbit per unique pair.
+For example, the primary of a low-mass pair selected at $z=3$ and merges before $z=2$ may be selected at a later snapshot with a new secondary companion. 
+In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are kept, since having the primary occur twice in the catalog does not mean that the orbit is being double counted ugh.
+
+\subsection{Calculating the Merger Fraction}
+\subsection{Calculating the Time Until Merger}
+
+\begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
+    \caption{The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
+At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
+The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.}
+    \label{fig:fmerge}
+    \end{center}
+\end{figure}
+
+\section{Pair Sample Properties}
 
@@
     \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger_full_fit.png}
-    \caption{}
-    % \label{fig:low-vs-high}
+    \caption{\kc{TODO: add LookbackTime vs.
+Z line, add Hubble time line} The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs.
+Shaded regions represent the 1st and 3rd quartile spread on the median.
+Pairs at each redshift are selected if they have experienced first infall.
+    % 
+    The black dashed line shows the Hubble time, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
+    The dotted grey line shows the time to $z=0$ (the Lookback Time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
+    The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
+    \label{fig:timescales}
 \end{figure*}
@@
     \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger.png}
-    \caption{}
-    % \label{fig:low-vs-high}
+    \caption{(Top) The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $(50,70,100,150,200,\mbox{and }300)\,\kpc$ from left to right.
+    (Bottom) The median time until merger as a function of redshift for pairs with physical separations greater than $10\,\kpc$ and less than $(0.25, 0.5, 1, 1.5, 2,\mbox{and }2.5)\,\Rvir$ from left to right.
+    % 
+    The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$.
+    }
+    \label{fig:timescales-sep}
 \end{figure*}

### 2024-03-18 paper/main.tex
Updates from Overleaf

@@
 
-The total number of low-mass pairs selected at each snapshot (given by $\Sigma_{S_n} n_p$) the sum of the number of pairs that pass the selection criteria at each snapshot is XXX, while the total number of \textit{unique} pairs is XXX.
-The total number of high-mass pairs is XXX, while the total number of unique pairs is XXX.
+For $n_k$ the number of pairs selected at snapshot $S_k$ via the selection criteria/from the base sample, the total number of pairs is given by
+\begin{equation*}
+    \sum_{k=0}^{99}\, n_k
+\end{equation*}
+The total number of low-mass pairs is 71,429, and the total number of high-mass pairs is 20,824.
+After selecting only unique pairs, there remain 22,213 unique low-mass pairs, and 3,029 unique high-mass pairs.
 
 Note that a single subhalo may be a member of many different pairs, but will only have one unique orbit per unique pair.
-For example, the primary of a low-mass pair selected at $z=3$ and merges before $z=2$ may be selected at a later snapshot with a new secondary companion. 
-In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are kept, since having the primary occur twice in the catalog does not mean that the orbit is being double counted ugh.
+For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
+In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are kept, since having the primary occur twice in the catalog does not mean that the orbit is being double counted.
+ugh.
 
-\subsection{Calculating the Merger Fraction}
-\subsection{Calculating the Time Until Merger}
+Maybe need to define the "unique pair sample" and the "snapshot selected sample" to explain the plots?
+I think all plots EXCEPT the merger fraction are with unique pair sample....
+DO I need to change that plot?
+% \subsection{Calculating the Number of Unique Pairs}
+
+% \subsection{Calculating the Merger Fraction}
+
+% \subsection{Calculating the Time Until Merger}
+
+\section{Pair Sample Properties}
+\subsection{Number of Unique Pairs}
+We calculate the number of unique pairs at a given snapshot by selecting all unique pairs at that snapshot that have already experienced first infall, and then sum the number of pairs that have yet to merge.
+Thus, the total number of unique pairs at $z=2$ includes pairs that were added to the catalog at \textit{any} snapshot, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
+A single pair (for example, a pair that was first selected in the snapshot-selected catalogs at $z=1$) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of unique pairs at all redshifts from $z=0-3$.
+
+In Fig.~\ref{fig:num-pairs}, we show the number of low-mass and high-mass unique pairs as a function of redshift from $z=0-6$. 
+Unique low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
+The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
+The number of pairs that merge decreases to 0 at $z=0$ since pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and low redshift selected pairs will on average have merger timescales longer than the remaining time to the end of the simulation.
+
+The light green and light pink shaded regions show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met.
+These lines are identical to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}.
+
+\begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/3_number_pairs.png}
+    \caption{\kc{change redshift range to 0-6}, add \# pairs from pears, change dash to dots}
+    \label{fig:num-pairs}
+    \end{center}
+\end{figure}
+
+\subsection{Merger Fraction}
 
@@
 \end{figure}
-
-\section{Pair Sample Properties}
 
@@
 % \end{figure}
-\begin{figure}[htb]
-    \centering
-    \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_number_pairs.png}
-\end{figure}
+
 % \begin{figure}[htb]

### 2024-03-20 paper/main.tex

@@
-\documentclass[twocolumn]{aastex631}
+\documentclass[twocolumn,linenumbers]{aastex631}
 
@@
   }
-\newcommand{\gb}[1]{\shadetext[left color=blue, right color=red, middle color=lime, shading angle=45]{\textbf{g: #1}} }
+\newcommand{\ep}[1]{\shadetext[left color=blue, right color=red, middle color=lime, shading angle=45]{\textbf{e: #1}} }
 % \newcommand{\ecite}[1]{\textcolor{pink}{\textbf{: #1}} }
@@
 
-\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxies Across Cosmic Time
+\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time
 }
@@
 % Problem 
-However, many studies continue to use equivalent separation criteria to select galaxy pairs for pair fraction, merger fraction, and merger rate studies, though it is unclear that such a selection permits an equitable comparison between galaxies of different mass, nor as a function of redshift.
+Pair fraction, merger fraction, and merger rate studies of high-mass pairs have been studied extensively, both observationally and theoretically. 
+However, it is unclear if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift.
 % Our solution
-We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass major pairs as a function of cosmic time. 
+Here, we present a physically-motivated framework by which merger timescales for high-mass and low-mass pairs can be compared equivalently 
+%and converted from one system to another 
+by scaling by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$). 
+We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass major pairs as a function of cosmic time, and compare to the merger timescales of similarly-selected high-mass major pairs.
+Using our framework to select pairs via a scaled separation criterion leads to equivalent merger timescales for low- and high-mass systems at redshifts $z=0-6$. 
 % What are the timescales? 
-We show that separation limits that vary with the mass and redshift of the system, such as scaling by the virial radius of the host halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$), lead to equivalent merger timescales for low- and high-mass systems at all redshifts.
-Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$) leads to timescales that differ between low- and high-mass pairs.
+Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$), leads to timescales that differ by up to $\sim1\Gyr$ between low- and high-mass pairs.
 % Why it matters 
-Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates must be done with self-consistent separation criteria that scales as a function of mass and redshift of the system.
+Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system.
 As a result, applying the same merger timescales to different mass systems will lead to a bias that systematically (over/under) predicts low-mass galaxy merger rates. 
@@
 
-Our sample consists of major low-mass and high-mass pairs that are isolated but physically associated.
+Our sample consists of major (stellar mass ratio $\M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
 From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.
@@
 \subsection{Pair orbit sample}
-We begin with a subset of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
-A brief version of the selection routine is described here.
+We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
+Rather than only considering pairs for $z<4$, we will collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
+A brief version of the selection routine is transcribed here for completeness, and note that more detail and discussion of the choices for selection criteria are in Sec.~2 of~\citet{Chamberlain2024}.
 
@@
         \mbox{\textbf{low mass:}}&\,\rm M_{G} = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
-        \mbox{\textbf{high mass:}}&\, \rm M_{G}=10^{12}- 6.5\times10^{12}\,\Msun.
+        \mbox{\textbf{high mass:}}&\, \rm M_{G}=1\times 10^{12}- 6.5\times10^{12}\,\Msun.
 \end{align*}
-Utilizing the most massive subhalos from the same FoF group ensures that the pairs are isolated from massive nearby systems that may perturb the dynamical state of the pair.
+Utilizing the most massive subhalos from the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.
 
-We require that subhalos that constitute a pair meet a minimum subhalo mass criteria of 
+We require that subhalos constituting a pair meet a minimum subhalo mass criteria of 
 \begin{equation*}
@@
 
-Stellar masses are then assigned to each subhalo in the FoF group using the Abundance Matching prescription of \citet{Moster2013}. 
-The peak halo mass and current snapshot are used to calculate the stellar mass of each subhalo, so that the abundance matching prescription is of the form $\ms{}=f(\Mpeak,z_{\mathrm{snap}})$.
+Stellar masses are then assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
+The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo, so that the abundance matching prescription is of the form $\ms{}=f(\Mpeak,z)$, with $z$ the redshift of consideration.
 
-In \citet{Chamberlain2024}, the abundance matching prescription is sampled 1000 times for each subhalo to account for the spread in the abundance matching prescription. 
+In \citet{Chamberlain2024}, the abundance matching prescription was sampled 1000 times for each subhalo to account for the spread in the Stellar Mass -- Halo Mass relation. 
 However, for the purposes of the present study, we only use the stellar masses given by the median of the abundance matching relationship. 
-This choice allows us to more thoroughly consider the population under study, and also permits a roughly 1:1 monotonic relationship between stellar mass and peak halo mass, so that taking subsamples of the lowest stellar mass objects will result in a subsample with lower peak halo mass as well.
+This choice permits a roughly 1:1 monotonic relationship between stellar mass and peak halo mass, so that taking subsamples of the lowest stellar mass objects will result in a subsample with lower peak halo mass as well.
 % is this a good enough reason?
-% else, maybe spread of number of pairs is so small, that we expect the noise of the timescale analysis to be dominated by the snapshot spacing anyways?
+Additionally, ~\cite{Chamberlain2024} found that the spread in the number of pairs selected at a given redshift is small, such that we expect the spread on the timescale analysis to be dominated by the redshift spacing of the simulation.
 
@@
 
-The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} catalogs, which contain information about every FoF group and subhalo at each snapshot.
+The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each snapshot.
 The \texttt{SUBLINK} catalogs track subhalos from one snapshot to the next, and thus enable us to track subhalos both backwards and forwards in time from any given snapshot.
@@
 \subsubsection{Orbits} 
-We calculate orbits for all mergers and non-mergers in our base sample. 
-An orbit for a single pair is defined to be the physical separation between the primary and secondary halo as a function of redshift or Lookback Time.
+\begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=\columnwidth, trim={0.4cm 1.75cm 1cm 3.5cm}, clip]{plots/bet-on-it/5_exampleorbits.png}
+    \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time.
+The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$.
+    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group, while the dashed lines show where they are not within the same FoF group. 
+    (Top) Orbits of pairs that merge before $z=0$.
+    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
+    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation). \kc{update fig}
+    }
+    \label{fig:example-orbits}
+    \end{center}
+\end{figure}
+
+We extract orbits for all mergers and non-mergers in our base sample. 
+An orbit for a single pair is defined to be the physical separation between the primary and secondary halo as a function of redshift or lookback time.
 
@@
 However, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group.
-Specifically, we will consider only the "Post-Infall" part of each pair's orbit.
+Specifically, we will consider only the ``Post-Infall" part of each pair's orbit.
 
-We define the snapshot of "First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo, and "Post-Infall" as the set of snapshots including the snapshot of "First Infall" and all snapshots thereafter.
+We define the snapshot of``First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo, and "Post-Infall" as the set of snapshots including the snapshot of "First Infall" and all snapshots thereafter.
 The post-infall orbit definition is slightly more robust for considering the full interaction timescales of merging pairs than only considering the orbit when the FoF group of the primary and secondary are the same.
@@
 The "post-infall" part of the orbit occurs after the first snapshot where the solid lines begins.
-Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs, though the spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between $\sim1-10\Gyr$.
+The spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between $\sim1-10\Gyr$.
+% a variety of different orbits
 
@@
 
-\begin{figure}[htb]
-    \begin{center}
-    \includegraphics[width=\columnwidth, trim={0.4cm 1.75cm 1cm 3.5cm}, clip]{plots/bet-on-it/5_exampleorbits.png}
-    \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of Lookback Time.
-The vertical grey line marks $z=1.5$ at a Lookback Time of $9.5\,\Gyr$.
-    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group, while the dashed lines show where they are not within the same FoF group. 
-    (Top) Orbits of pairs that merge before $z=0$. Pairs chosen at $z=1.5$ on average merge at \kc{$z=XX$/within XX Gyrs}.
-    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
-    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation). 
-    }
-    \label{fig:example-orbits}
-    \end{center}
-\end{figure}
-
-\subsubsection{Uniqueness of pairs}
+\subsubsection{Uniqueness of orbits}
 Since our entire pair sample consists of pairs chosen at separate snapshots, and since a single pair could pass our selection criteria at many snapshots, we keep only a single orbital instance of each pair.
@@
 \section{Pair Sample Properties}
+\begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
+    \caption{\kc{need caption}\kc{Change redshift limits?} The number of low-mass (green) and high-mass (pink) pairs as a function of redshift.
+The solid lines show the number of unique pairs at a given redshift, while dashed lines show the number of unique pairs at a given redshift that will merge before $z=0$.
+    For these two lines, we have only counted pairs at a given redshift if they have not merged, but have experienced first infall at a previous time. 
+    \kc{I believe} The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample due to mergers.
+    %
+    The thin dotted lines show the number of pairs that were selected via the pair criteria at that specific redshift. 
+    }
+    \label{fig:num-pairs}
+    \end{center}
+\end{figure}
+
+\begin{figure}[htb]
+    \begin{center}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
+    \caption{\kc{redshift:  0-6} The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
+At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
+The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.\kc{might need to add on second set of lines, so i have both merger fraction from snapshot-selected and from orbit-selected.}}
+    \label{fig:fmerge}
+    \end{center}
+\end{figure}
+
 \subsection{Number of Unique Pairs}
@@
 
-In Fig.~\ref{fig:num-pairs}, we show the number of low-mass and high-mass unique pairs as a function of redshift from $z=0-6$. 
+Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass unique pairs as a function of redshift from $z=0-6$. 
 Unique low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
@@
 
-The light green and light pink shaded regions show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met.
-These lines are identical to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}.
-
-\begin{figure}[htb]
-    \begin{center}
-    \includegraphics[width=\columnwidth]{plots/bet-on-it/3_number_pairs.png}
-    \caption{\kc{change redshift range to 0-6}, add \# pairs from pears, change dash to dots}
-    \label{fig:num-pairs}
-    \end{center}
-\end{figure}
+The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met.
+These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.
 
 \subsection{Merger Fraction}
+We calculate the merger fraction, $\rm F_{merge}$, at a given snapshot by dividing the number of unique merging pairs (dashed lines in Fig.~\ref{fig:num-pairs}) by the number of total unique pairs (solid lines).
+As before, a pair whose first infall occurs at $z=2$ and that merges at $z=1$ will be included in the merger fraction calculation for redshift $z=1-2$.
 
-\begin{figure}[htb]
-    \begin{center}
-    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
-    \caption{The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
-At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
-The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.}
-    \label{fig:fmerge}
-    \end{center}
-\end{figure}
+Note that the merger fraction is a measure of the fraction of isolated pairs, whose primary and secondary subhalos have shared a common FoF group, that merge before $z=0$. 
+It is not, however, a measure of the fraction of \textit{all} low-mass and high-mass pairs that will merge.
+
+Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation. 
+At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$, but declines to $0$ at $z=0$. 
+This is due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
 
 % ###########################################################################################
-% Results II 
+% Results I
 % ###########################################################################################
 
-\section{Results - The Mass and Redshift Dependence of Merger Timescales}
+\section{Results: The Mass and Redshift Dependence of Merger Timescales}
 % The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled with mass and redshift.
+\begin{figure*}[htb]
+    \begin{center}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
+    \caption{\kc{Consider what redshifts to show.
+Either add colorbar normalized for all panels (in counts) or add \# of halos and a percentage.
+Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
+    There is a positive correlation between separation at the selected redshift and the time until merger. 
+    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$. 
+    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
+    The lowest separation pairs have the shortest merger timescales.
+    This plot can be used to get estimates for the merger timescale of an isolated low-mass pair at a given redshift.
+For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
+    \kc{add z=5, z=6}
+    }
+    \label{fig:timevssep-low}
+    \end{center}
+\end{figure*}
+
+\begin{figure*}[htb]
+    \begin{center}
+    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
+    \caption{\kc{Consider what redshifts to show.
+Add colorbar.
+Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
+    Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
+    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$. 
+    This plot can be used to get estimates for the merger timescale of an isolated high-mass pair at a given redshift. 
+    For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
+    \kc{add z=5, z=6}
+    }
+    \label{fig:timevssep-high}
+    \end{center}
+\end{figure*}
+
+\subsection{Time Until Merger vs. Separation}
+% how we calculate for this section
+We calculate the time until merger as a function of separation for low-mass and high-mass pairs at 6 different redshifts ($z=0.5,1,1.5,2,3,4$). 
+At a given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each different pair.
+
+% low mass example
+For example, at $z=1.5$, there are XXX unique low-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$.
+Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.
+
+% high mass example
+Likewise, at $z=1.5$, there are XXX unique high-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$.
+Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.
+% might put the above in a table with more info?
+
+% description of figures 
+Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs. 
+In Fig.~\ref{fig:timevssep-low}, the time until merger for low-mass pairs is positively correlated to the separation of the pair at each redshift.
+
+The figures can be used to get approximate `time until merger' ranges for a given separation at a given redshift.
+
+% 
+\input{table-septime.tex}
+
+\subsection{Time Until Merger vs. Redshift}
+
+Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs.
+
+\subsubsection{Separation-selected samples}
+
 \subsection{"time til merger" for selected snapshots for LvH.}
 Maybe also (distribution plot 2D hist? as function of selected separation)
-\subsection{Cumulative distributions}
-\subsubsection{Physical}
-\subsubsection{Scaled}
 
@@
 \section{Discussion}
+
+\subsection{About the number of pairs}
+Can we say anything about the REASON the unique pairs decreases at low z?
+Does this plot mean that it's only because things merge and so we lose them?
+And not that it's driven by mass-accretion thing?
+I think so.
+
+Need to be careful about phrasing since we didn't actually quantify the number of pairs that leave the sample due to having too large of masses at later snapshots.
+
+\subsection{About the merger fraction}
+Were low-mass pairs expected to have such a high merger rate in isolated environs? 
+How do these merger fractions compare to other literature results? 
+If we can make comparisons, discuss the differences. 
+If we cannot make comparisons, explain why our study is done in a way that makes for comparisons unfair/undoable .
+
 % \subsection{Merger Rates for Dwarf Pairs}
@@
 \appendix
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
-    \caption{}
-    % \label{fig:low-vs-high}
-\end{figure*}
-
-\begin{figure*}[htb]
-    \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
-    \caption{}
-    % \label{fig:low-vs-high}
-\end{figure*}
+adfadfa
 % \begin{figure}[htb]

### 2024-03-20 paper/main.tex
Updates from Overleaf

@@
     \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time. 
-The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$.
-    Solid lines represent the parts of the orbit where the primary and secondary halo have the same FoF group, while the dashed lines show where they are not within the same FoF group. 
     (Top) Orbits of pairs that merge before $z=0$.
     (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
-    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation). \kc{update fig}
+    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
+    Solid lines represent the parts of the orbit post-infall, while the dashed lines show the pre-infall portion of the orbit before the pairs first share a common FoF group. 
+    Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
+    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$.
     }

### 2024-03-20 paper/main.tex
Updates from Overleaf

@@
 \subsubsection{Orbits} 
-\begin{figure}[htb]
+\begin{figure}[tb]
     \begin{center}
-    \includegraphics[width=\columnwidth, trim={0.4cm 1.75cm 1cm 3.5cm}, clip]{plots/bet-on-it/5_exampleorbits.png}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
     \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time. 
@@
     Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
-    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$.
+    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$. \kc{state in words that the FoF group need not be common for the remainder of the orbit, but will definitely be by merger.}
     }
@@
     For these two lines, we have only counted pairs at a given redshift if they have not merged, but have experienced first infall at a previous time. 
-    \kc{I believe} The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample due to mergers.
+    \kc{I believe... think more} The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample due to mergers.
     %

### 2024-03-20 paper/main.tex
finished first round of edits on methods, snapshot -> redshift focus, and updated plots/captions/description of the sample

@@
 %recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
-We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-XX$.
+We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
+The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).
 
@@
 
-\subsection{Pair orbit sample}
+\subsection{Pair sample}
 We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
@@
 
-The base sample used for this analysis then consists of the set of all low-mass and high-mass major pairs from each snapshot of the TNG100 simulation.
+The base sample of pairs used for this analysis then consists of the set of all low-mass and high-mass major pairs from each redshift of the TNG100 simulation.
 
-\subsubsection{Mergers}
+\subsection{Mergers}
 % Plot \#1: Merger fraction either before or after orbits
@@
 
-The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each snapshot.
-The \texttt{SUBLINK} catalogs track subhalos from one snapshot to the next, and thus enable us to track subhalos both backwards and forwards in time from any given snapshot.
+The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift.
+The \texttt{SUBLINK} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.
 
-We utilize the \texttt{RootDescendantID} and \texttt{DescendantID} fields of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$), and to determine when the merger occurs.
+We utilize the \texttt{DescendantID} field of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
 The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
@@
 % By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.
-The \texttt{RootDescendantID} field provides the SubhaloID of the root subhalo, which is the latest subhalo in the simulation that is a descendent of the subhalo.
 
-We define the set of merging pairs as the collection of all pairs where the primary and secondary halo have the same \texttt{RootDescendantID}, meaning that the branches of their merger trees have the same `root', or common descendent.
-If a given pair has two distinct \texttt{RootDescendantID} values, the pair is defined as a `non-merger.'
+A pair will be classified as a merger if the primary and secondary subhalo share the same \texttt{DescendantID} at the same redshift. 
+This means that the primary and secondary subhalo will merge, producing a single descendant subhalo, by the following redshift from the simulation.
+If a pair consists of a primary and secondary that never share a \texttt{DescendantID} at a given redshift, the pair is defined as a `non-merger.'
 
-For each merging pair, we define the `merger snapshot' as the snapshot that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
-For example, if the primary and secondary have different \texttt{DescendantIDs} for snapshots 40-45, but have the same \texttt{DescendantID} at snapshot 46, then the merger must take place some time between snapshot 46 and 47.
-We take snapshot 47 to be the merger snapshot.
+For each merging pair, we define the `merger redshift' as the redshift that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
+For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place some time between $z=2$ and $z=1.9$, which is the next redshift in the simulation.
+We take $z=1.9$ to be the merger redshift.
 
-\subsubsection{Orbits} 
+\subsection{Orbits} 
 \begin{figure}[tb]
@@
     \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
-    \caption{Selection of example orbits of low-mass major pairs selected at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time. 
+    \caption{Selection of example orbits of low-mass major pairs that pass the pair selection criteria at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time.
     (Top) Orbits of pairs that merge before $z=0$.
-    (Middle) Orbits of pairs that did not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
-    (Bottom) Orbits of pairs that did not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
-    Solid lines represent the parts of the orbit post-infall, while the dashed lines show the pre-infall portion of the orbit before the pairs first share a common FoF group. 
+    (Middle) Orbits of pairs that do not merge before $z=0$ (non-mergers), but are likely to merge if the simulation continued. 
+    (Bottom) Orbits of pairs that do not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
+    Solid lines represent the post-infall parts of the orbit, while the dashed lines show the pre-infall portion of the orbit before the pairs first share a common FoF group. 
     Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
-    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$. \kc{state in words that the FoF group need not be common for the remainder of the orbit, but will definitely be by merger.}
+    The vertical grey line marks $z=1.5$ at a lookback time of $9.5\,\Gyr$. 
     }
@@
 
-We extract orbits for all mergers and non-mergers in our base sample. 
-An orbit for a single pair is defined to be the physical separation between the primary and secondary halo as a function of redshift or lookback time.
+We extract orbits for all mergers and non-mergers in our base pair sample. 
+An orbit for a single pair is defined to be the physical separation between the primary and secondary subhalo as a function of redshift (or lookback time).
 
-A given pair from the base sample at snapshot $S_n$, which by definition passes all selection criteria at $S_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.
-We track the positions of both the primary and secondary subhalo at each snapshot and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
-In cases where the primary or secondary does not have a defined position at a given snapshot, we set the separation at that snapshot to NaN.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs.
-The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be computed contiguously for a pair on .
+A given pair from the base sample at redshift $z_n$, which by definition passes all selection criteria at $z_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.
+We track the positions of both the primary and secondary subhalo at each redshift and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
+In cases where the primary or secondary does not have a defined position at a given redshift, we set the separation at that snapshot to NaN.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs.
+The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be evaluated before and after the skip occurs.
 See Sec.~3 in ~\citet{RG2015} for more details.}
 
-In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot. We use the definition of scaled separation from~\cite{Chamberlain2024}, given as
-\begin{equation}
-    \rm \Rsc = \frac{\rsep}{\Rvir}
-\end{equation}
+In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot.
+We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction of the virial radius $\rsep=X\Rvir$
+% \begin{equation}
+%     \rm \Rsc = \frac{\rsep}{\Rvir}
+% \end{equation}
 where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the primary's FoF group, which reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
-The scaled radius $\Rsc$ is by construction a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs, or for halo growth over time.
+This ``scaled radius" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.
 
-We collect the following data for each pair at each snapshot where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group \texttt{Group\_R\_TopHat200}, the physical separation (in $\kpc$), and the scaled separation (dimensionless).
+We collect the following data for each pair at each redshift where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group \texttt{Group\_R\_TopHat200}, the physical separation (in $\kpc$), and the scaled separation (dimensionless).
 
-\subsubsection{"First Infall"}
-The orbit of a pair may be technically defined at very early times, well before the primary and secondary are physically associated. 
-However, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group.
-Specifically, we will consider only the ``Post-Infall" part of each pair's orbit.
+\subsubsection{Defining Post-Infall}
+While the orbit of a pair may be calculated at very early times, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group.
+Specifically, we will consider only the ``post-infall" part of each pair's orbit.
 
-We define the snapshot of``First Infall" as the first snapshot where the primary and secondary have the same parent FoF halo, and "Post-Infall" as the set of snapshots including the snapshot of "First Infall" and all snapshots thereafter.
-The post-infall orbit definition is slightly more robust for considering the full interaction timescales of merging pairs than only considering the orbit when the FoF group of the primary and secondary are the same.
-This is because there are cases of pair interactions where the secondary will enter the FoF group of the primary, have a pericentric passage, then leave the FoF group for a few gigayears before re-entering the primary FoF group and eventually merging with the primary.
+We define the redshift of ``first infall" as the redshift of the first snapshot where the primary and secondary have the same parent FoF halo, and post-infall is all following redshifts starting with the redshift of first infall.
 
-An example of the orbit of a merging pair where a "common group" criteria may differ from a "post-infall" can be seen in Figure~\ref{fig:example-orbits}, which shows the variety of orbits that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$.
-The top panel shows the orbits of 5 pairs that merge, where solid lines show the parts of the orbit that occur while the primary and secondary share a common FoF group, and dashed lines show the parts of the merger where the FoF group of the secondary is different than that of the primary.
-The "post-infall" part of the orbit occurs after the first snapshot where the solid lines begins.
-The spread of merger timescales is large, leading to a single redshift-selected sample having merger timescales between $\sim1-10\Gyr$.
+Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$.
+The top panel shows the orbits of 5 pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo.
+Even amongst galaxies of the same approximate mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\Gyr$ from first infall to merger.
+
+Some secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo. 
+Because all of our mergers occur when the subhalos are in the same FoF group, the
+post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same.
 % a variety of different orbits
 
-The middle panel shows the orbits for 3 pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a couple Gyr past the end of the simulation, and are thus "Potential Future Mergers".
+The middle panel of Fig.~\ref{fig:example-orbits} (``Potential Future Mergers") shows the orbits for 3 pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a few Gyr past the end of the simulation.
 These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
-The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past 10Gyr, while the darker blue orbit has a much shorter period with many close encounters throughout, and 3 close passages in just the past 2 Gyr.
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with many close encounters throughout, and 3 close passages in just the past $2\,\Gyr$.
+
 The bottom panel shows the orbits of "Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
@@
 \subsubsection{Uniqueness of orbits}
-Since our entire pair sample consists of pairs chosen at separate snapshots, and since a single pair could pass our selection criteria at many snapshots, we keep only a single orbital instance of each pair.
+A given pair from the \paircat{} may pass our selection criteria at many redshifts. 
+Since we collect the full orbit (the separation at every redshift where the primary and secondary both exist) for each pair and each redshift, the orbits that are collected for pairs that pass our selection criteria at more than one redshift will all result in identical orbits.
+% However, we only want to keep a single instance of any given orbit in our catalog.
 
-To distinguish pairs from one snapshot to another, each pair is assigned a pairkey while constructing their orbit, which is created by combining the primary's and secondary's earliest SubhaloID from the Sublink catalogs, which will be unique for each pair of halos.
+To distinguish the collection of unique orbits, each pair is assigned a `pairkey' while constructing their orbit. 
+The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the Sublink catalogs, and is unique for each pair of halos. 
 After each pair is assigned a unique pairkey, we save the orbit as a function of time only once for a pair that otherwise would have been double/multi-counted in our orbit catalog.
-
-For $n_k$ the number of pairs selected at snapshot $S_k$ via the selection criteria/from the base sample, the total number of pairs is given by
-\begin{equation*}
-    \sum_{k=0}^{99}\, n_k
-\end{equation*}
-The total number of low-mass pairs is 71,429, and the total number of high-mass pairs is 20,824.
-After selecting only unique pairs, there remain 22,213 unique low-mass pairs, and 3,029 unique high-mass pairs.
 
@@
 For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
-In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are kept, since having the primary occur twice in the catalog does not mean that the orbit is being double counted.
-ugh.
+In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are retained in the orbit catalog, since each orbit is unique.
 
-Maybe need to define the "unique pair sample" and the "snapshot selected sample" to explain the plots?
-I think all plots EXCEPT the merger fraction are with unique pair sample....
-DO I need to change that plot?
-% \subsection{Calculating the Number of Unique Pairs}
+Removing multiplicity decreases the raw number of orbits collected by collecting an orbit for each pair from the base sample. 
+The total number of pairs (including double counting the same pair at multiple redshifts) is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
+However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits.
 
-% \subsection{Calculating the Merger Fraction}
+% For $n_k$ the number of pairs selected at snapshot $S_k$ via the selection criteria/from the base sample, the total number of pairs is given by
+% \begin{equation*}
+%     \sum_{k=0}^{99}\, n_k
+% \end{equation*}
+% The total number, and the total number of high-mass pairs is .
+% After selecting only unique pairs, there remain
 
-% \subsection{Calculating the Time Until Merger}
+The collection of all unique orbits constitutes our orbit sample, and is the dataset that will be used for the remainder of the paper.
 
@@
     \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
-    \caption{\kc{need caption}\kc{Change redshift limits?} The number of low-mass (green) and high-mass (pink) pairs as a function of redshift.
-The solid lines show the number of unique pairs at a given redshift, while dashed lines show the number of unique pairs at a given redshift that will merge before $z=0$.
-    For these two lines, we have only counted pairs at a given redshift if they have not merged, but have experienced first infall at a previous time. 
-    \kc{I believe... think more} The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample due to mergers.
-    %
-    The thin dotted lines show the number of pairs that were selected via the pair criteria at that specific redshift. 
+    \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
+    The solid lines show the total number of orbits the are post-infall at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
+    We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
+    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
     }
-    \label{fig:num-pairs}
+    \label{fig:numorbits}
     \end{center}
@@
     \begin{center}
-    \includegraphics[width=\columnwidth]{plots/bet-on-it/1_fmerge_comp.png}
-    \caption{\kc{redshift:  0-6} The fraction of pairs that merge before z=0 as a function of the redshift where they were selected via the selection criteria for the \paircat{} from \citet{Chamberlain2024}.
-At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs.
-The sharp decline to $F_{\rm merge}$ at $z=0$ is a non-physical feature of the simulation ending at $z=0$, thus unable to follow the pairs after the simulation ends to see what fraction of pairs may have merged.\kc{might need to add on second set of lines, so i have both merger fraction from snapshot-selected and from orbit-selected.}}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
+    \caption{The fraction of low-mass (green) and high-mass (pink) post-infall orbits that result in a merger before z=0 as a function of redshift. 
+    The horizontal line shows a merger fraction of 0.9, or 90\%. 
+    High-mass pairs have merger fractions greater than 0.9 for $z>1.1$, and low-mass pairs for $z>1.6$.
+    %At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. 
+    The sharp decline of the merger fraction to 0 at low redshift is a non-physical feature of the simulation ending at $z=0$.
+    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
     \label{fig:fmerge}
@@
 
-\subsection{Number of Unique Pairs}
-We calculate the number of unique pairs at a given snapshot by selecting all unique pairs at that snapshot that have already experienced first infall, and then sum the number of pairs that have yet to merge.
-Thus, the total number of unique pairs at $z=2$ includes pairs that were added to the catalog at \textit{any} snapshot, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
-A single pair (for example, a pair that was first selected in the snapshot-selected catalogs at $z=1$) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of unique pairs at all redshifts from $z=0-3$.
+\subsection{Number of Orbits}
+We calculate the number of orbits (or orbits in progress) by selecting all orbits at a given redshift that are post-infall and either pre-merger or non-mergers.
+The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
+For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
+A single pair (or equivalently, the pair's orbit) orbit may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of orbits at all redshifts from $z=0-3$.
 
-Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass unique pairs as a function of redshift from $z=0-6$. 
-Unique low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
+Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
+Low-mass orbits (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
 The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
-The number of pairs that merge decreases to 0 at $z=0$ since pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and low redshift selected pairs will on average have merger timescales longer than the remaining time to the end of the simulation.
+The number of pairs that merge decreases to 0 at $z=0$ since 
+%pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and 
+many pairs that exist at low redshift will have merger timescales longer than the remaining time to the end of the simulation.
 
-The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met.
-These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.
+% removed
+% The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met.
+% These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.
 
 \subsection{Merger Fraction}
-We calculate the merger fraction, $\rm F_{merge}$, at a given snapshot by dividing the number of unique merging pairs (dashed lines in Fig.~\ref{fig:num-pairs}) by the number of total unique pairs (solid lines).
-As before, a pair whose first infall occurs at $z=2$ and that merges at $z=1$ will be included in the merger fraction calculation for redshift $z=1-2$.
+We calculate the merger fraction by dividing the number of merging orbits (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift. 
+As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
 
-Note that the merger fraction is a measure of the fraction of isolated pairs, whose primary and secondary subhalos have shared a common FoF group, that merge before $z=0$. 
+Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation as a function of redshift. 
+At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
+The merger fraction for both mass ranges declines to $0$ at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
+At $z=0$, any remaining pairs are classified as non-mergers since they did not merge \textit{$z=0$}.
+
+Note that the merger fraction is a measure of the fraction of isolated pairs that merge before $z=0$. 
 It is not, however, a measure of the fraction of \textit{all} low-mass and high-mass pairs that will merge.
-
-Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation. 
-At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$, but declines to $0$ at $z=0$. 
-This is due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
 
@@
 
-** 
-\appendix
-adfadfa
+% ** 
+% \appendix
+% adfadfa
 % \begin{figure}[htb]

### 2024-03-21 paper/main.tex
almost finished with results section, only full sample to write. next: discussion, conclusions, plot fixes, and intro

@@
 \newcommand{\rsep}{\ensuremath{\rm r_{sep}}}
+
+\newcommand{\chambe}{\citet{Chamberlain2024}}
 
@@
 %%% OUTLINE
-\section{Methodology}
+\section{Methodology} \label{sec:methods}
 %recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
@@
 
-Our sample consists of major (stellar mass ratio $\M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
+Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
 From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.
@@
 In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot.
-We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction of the virial radius $\rsep=X\Rvir$
-% \begin{equation}
-%     \rm \Rsc = \frac{\rsep}{\Rvir}
-% \end{equation}
-where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the primary's FoF group, which reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
-This ``scaled radius" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.
+We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
+The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
+This ``scaled separation" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.
 
@@
 
-Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
+Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
 Low-mass orbits (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
@@
 \section{Results: The Mass and Redshift Dependence of Merger Timescales}
-% The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled with mass and redshift.
+We have created catalogs of unique orbits for a large sample of isolated low-mass and high-mass pairs in the TNG100 simulation.
+Next, we will calculate the merger timescales, or time until merger, for all of the merging orbits in our catalog. 
+In Sec.~\ref{sec:results-timevsep}, we explore how the time until merger changes as a function of the separation of the pair at a variety of redshifts, and study how the time until merger changes between the low-mass sample and high-mass sample.
+In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all post-infall and pre-merger isolated pairs as a function of redshift. 
+We will additionally examine the merger timescale's dependence on a variety of separation criteria.
+
+In the following analysis, we will utilize the orbit catalog, which consists of the unique orbits for all low-mass and high-mass major isolated pairs from the~\citet{Chamberlain2024} study. 
+At a given redshift, we will include all merging pairs that are post-infall and pre-merger at that redshift in our analysis. 
+However, separation criteria will also be applied to each redshift independently, so a pair that is in the $10-50\,\kpc$ bin at one redshift will not be part of the sample used at redshifts where it's separation is $>50\,\kpc$.
+
 \begin{figure*}[htb]
@@
     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
-    \caption{\kc{Consider what redshifts to show.
-Either add colorbar normalized for all panels (in counts) or add \# of halos and a percentage.
-Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
+    \caption{\kc{Add z=5, z=6 panels.
+Add colorbar for percentage per panel (and add number of pairs at that redshift) Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
     There is a positive correlation between separation at the selected redshift and the time until merger. 
@@
     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
-    \caption{\kc{Consider what redshifts to show.
+    \caption{\kc{Add z=5, z=6 panels.
 Add colorbar.
-Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
+Add low/high label somewhere, check the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
     Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
@@
     For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
-    \kc{add z=5, z=6}
     }
@@
 
-\subsection{Time Until Merger vs. Separation}
+\subsection{Separation Dependence of Merger Timescales}\label{sec:results-timevsep}
+\kc{note: this might go in the appendix as additional info}
 % how we calculate for this section
-We calculate the time until merger as a function of separation for low-mass and high-mass pairs at 6 different redshifts ($z=0.5,1,1.5,2,3,4$). 
-At a given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each different pair.
+We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
+At the given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each pair.
 
 % low mass example
-For example, at $z=1.5$, there are XXX unique low-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$.
-Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.
+For example, at $z=1.5$, there are 5,787 unique orbits that are post-infall and will merge prior to $z=0$.
+Of those, and 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 of those low-separation pairs will merge in the next $2\,\Gyr$.
+A compilation of this data is available in Table~\ref{tab:timevssepvsz}, where we list the fraction of post-infall and pre-merger pairs within 3 separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction within each of those separation bins that merge in 4 different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$, for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$..
 
-% high mass example
-Likewise, at $z=1.5$, there are XXX unique high-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$.
-Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.
-% might put the above in a table with more info?
+As an accompanying visual aid, figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs.
+separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$.
+Fig.~\ref{fig:timevssep-low} presents the distribution for low-mass subhalos, and shows that the time until merger for low-mass pairs is positively correlated with the separation of the pair at each redshift, but that the positive correlation is smaller at larger redshifts.
+In addition, the spread of times until merger and separations is smaller at higher redshift  than the spread at lower redshift, likely due to the size growth of halos over time. 
+% could make these same plots with scaled separation, then it would presumably show roughly the same spread for each redshift.
 
-% description of figures 
-Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs. 
-In Fig.~\ref{fig:timevssep-low}, the time until merger for low-mass pairs is positively correlated to the separation of the pair at each redshift.
+Likewise, fig.~\ref{fig:timevssep-high} presents the distribution for high-mass subhalos, where the trends between time until merger and separation are roughly the same. 
+At each redshift, there is a positive correlation between time until merger and separation, but the spread on either of those decreases with increasing redshift.
 
-The figures can be used to get approximate `time until merger' ranges for a given separation at a given redshift.
-
+These figures can be used to get approximate ranges of time until merger for a pair at a given separation and for a given redshift, and are useful for seeing the difference between the relationship between time until merger and separation for low-mass and high-mass pairs.
+\kc{what differences?}
 % 
@@
 
-\subsection{Time Until Merger vs. Redshift}
+\subsection{Redshift Dependence of Merger Timescales}\label{sec:results-timevredshift}
+% The behavior of time until merger as a function of redshift for low-mass and high-mass pairs:
+% 1 - the full sample 
+% 2 - the physical and static kpc separation-selected sample
+% 3 - the mass and redshift evolving scaled separation-selected sample
+We study the merger timescale of low-mass and high-mass pairs as a function of redshift. 
+In particular, we consider 3 different samples of the low- and high-mass systems. 
+First, we will consider the orbit sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that are post-infall and pre-merger for $z=0-6$. 
+We will then consider two different sets of separation criteria to create separation-selected subsamples, the first of which are physical separation criteria that are identically applied for both mass scales and all redshifts, and the second of which defines separation criteria that vary with mass and redshift.
+% We ascertain the impact of the selection criteria on these subsamples of the orbit catalog .....
+based on separation-selected pairs, using a set of separation criteria that we apply independently at each redshift.
 
-Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs.
+Both separation-selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \texttt{SUBFIND} catalogs.
+This lower separation criteria is also commonly applied to observationally-selected pairs in studies of merger fractions and merger rates (see Sec.~\ref{sec:discussion} for further detail).
 
-\subsubsection{Separation-selected samples}
+Throughout, medians at each redshift are shown by solid pink and green lines, while the 1st and 3rd quartile spread on the median is shown by the shaded regions.
 
-\subsection{"time til merger" for selected snapshots for LvH.}
-Maybe also (distribution plot 2D hist? as function of selected separation)
+\subsubsection{Full sample}
+We calculate the time until merger for all post-infall and pre-merger pairs at each redshift from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
+We make no selection cuts based on pair separation, and thus include the full catalog of orbits at each redshift to calculate the merger timescale (once again, provided that they are post-infall and pre-merger at the given redshift).
+Essentially, we find the median time that any physically associated isolated pair will take to merge, provided it's a merging pair to begin with.
 
-% \begin{figure*}[htb]
-%     \centering
-%     \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi.png}
-%     \caption{\kc{reverse the direction of z} 
-%     The median (solid) and 1st-3rd quartile spread (shaded regions) of the cumulative time that isolated low-mass pairs (left) and high-mass pairs (right) spend with separations \rsep{} between $10\kpc$ and \Rphys{}.
-%     The Merging Pairs sample (dark green and dark pink) selected at $z=(1,1.5,2)$ merge before $z=0$, while the All Pairs sample (light green and light pink) include all the pairs from the \paircat{} (mergers and non-mergers).
-%     The low-mass (high-mass) merging pairs represent XX\%(XX\%) of the low-mass (high-mass) pair sample at $z=1,1.5, \mbox{and } 2$ respectively. 
-%     The sample of All Pairs spend more time at higher separations than the merging pairs, as expected since the non-merging pairs are not likely to have very low separations that would result in mergers prior to $z=0$.
-%     For a more direct comparison between the low-mass and high-mass pairs, see Fig.~\ref{fig:phys-vs-scaled}. 
-%     }
-%     \label{fig:low-vs-high}
-% \end{figure*}
+Figure~\ref{fig:timescales} presents the median, and 1st and 3rd quartile spread, of the time until merger for merging pairs from $z=0-6$. 
+Low-mass merger timescales (green) and high-mass merger timescales (pink) are roughly equivalent at all redshifts. 
+The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
 
-% \begin{figure*}[htb]
-%     \centering
-%     \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT.png}
-%     \caption{\kc{Add legend to plot, and I think reverse the direction of z} 
-%     The median and 1st-3rd quartile spread of the cumulative time that isolated merging pairs, selected at $z=(1,1.5,2)$, spend with separations \rsep{} greater than $10\kpc$ and less than either \Rphys{} or \Rsc{}.
-%     %
-%     (Top) The median total time spent by isolated merging pairs with separations between $10\kpc$ and \Rphys{}.
-%     Low-mass pairs spend a longer amount of time within $\sim200\kpc$ than high-mass pairs at each of these redshifts. 
-%     However, high-mass galaxies spend more time at higher separations of $\rsep>200-300\kpc$.
-This is because the elapsed time is measured starting at the infall snapshot of the pair, and high-mass halos have larger radii, the high-mass halo orbits have a broader distribution of separations that extend to larger separations than those of the low-mass pairs.
-%     (Bottom) The median total time spent by pairs with separations greater than $\rsep>10\kpc$ and less than a given fraction of the virial radius ($\rsep/\Rvir<\Rsc$) of the pair, which scales with mass and redshift.
-%     High-mass pairs spend more time within the same scaled separation than low-mass pairs.
-For example, a high-mass pair selected from the $z=1.5$ sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}, while a low-mass pair from the same sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}.
-%     Low-mass and high-mass pairs have similar cumulative time profiles as a function of scaled separation, with a roughly constant offset for all scaled separations. 
-%     The high-mass pairs spend a median of \kc{$XX\Gyr$} more than low-mass pairs at every scaled separation at $z=1$ and $z=1.5$, and a median of \kc{$XX\Gyr$} more $z=2$.
-%     }
-%     \label{fig:phys-vs-scaled}
-% \end{figure*}
+\kc{left off here}
+
+% Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs.
+
+\subsubsection{Separation Criteria: Static \& Physical }
+The first set of separation criteria will select only pairs at a given redshift that have separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$. 
+These separation criteria do not vary as a function of time or mass, and are thus applied equivalently to the low-mass and high-mass samples. 
+This set of static physical separation criteria produces 6 subsamples of merging orbits that are post-infall.
+
+Note that an orbit in the separation bin $10-50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts. 
+Rather, the separation criteria select pairs with separations in the given separation bin independently at each redshift.
+
+% figure description 
+The top panel of Fig.~\ref{fig:timescales-sep} shows the time until merger versus redshift for low-mass (green) and high-mass (pink) orbits that have separations less than the separation listed above each plot.
+% redshift evolution of each subsample
+We first note that the evolution of the time until merger is approximately the same for each separation-selected subsample, with a peak time at redshift $z\sim0.5$ present for all subsamples, and a rapid decline of the median time until merger at $z=0-0.5$.
+These traits were also present in Fig.~\ref{fig:timescales} where the median time until merger was computed for the entire sample, without a selection-criterion.
+As the selection-criteria gets larger and larger, the subsample will contain more and more of the full sample, eventually reconstructing the full sample plot with a large enough separation selection.
+
+% low mass vs high mass
+There is an offset between the median times for low-mass and high-mass pairs for each of the subsamples with maximum separations of $[50, 70, \mbox{and } 100]\,\kpc$. 
+The difference in the median time until merger is up to $0.8\,\Gyr$ longer for low-mass pairs than high-mass pairs at the same redshift. 
+The offset between the low-mass and high-mass merger times decreases for large enough physical separation criteria. 
+In the rightmost top panel, the median merger times overlap almost identically for $z=0-0.5$ and $z>1.5$.
+
+% impact of the separation criteria
+The median time until merger for pairs selected in smaller separation bins (such as $10-50\,\kpc$ and $10-70\,\kpc$) is lower at all redshifts for both low-mass and high mass pairs. 
+This follows directly from Sec.~\ref{sec:results-timevsep}, where we found that the time until merger is positively correlated with increasing separation for all pairs. 
+Thus, including more and more larger-separation systems in our merger time calculation will tend to increase the median time until merger.
+
+\subsubsection{Separation Criteria: Scaled Separation}
+The second set of selection criteria will select pairs at a given redshift in 6 different scaled separation bins, where the scaled separation is written as a fraction of the virial radius of the pair's FoF group, $\Rvir$ (see Sec.~\ref{sec:methods} for more details).
+The scaled separation-selected pairs have separations greater than $10\,\kpc$ and less than $<[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$.
+
+A scaled separation criterion scales with both redshift and mass (since the virial radius scales with redshift and mass), and produces 6 subsamples of merging orbits that are post-infall.
+As mentioned in \chambe{}, the median virial radius for low-mass FoF groups at $z=[0,1,2,3,4]$ is approximately $[134, 85, 59, 43, 33]\,\kpc$, and for high-mass FoF groups is approximately $[348, 206, 134, 97, 76]\,\kpc$.
+Thus, choosing a scaled separation cut of $<1\,\Rvir$ at $z=1$ will select high-mass pairs with separations $\lesssim205\,\kpc$ and low-mass pairs with separations $\lesssim85\,\kpc$.
+
+As in the previous section, the selection criteria are applied at each redshift independently. 
+This means that the time until merger of a pair at a given redshift will only be included in the calculation of the median merger time if its separation is within the given separation bin at that redshift.
+
+% figure description 
+The bottom panel of Fig.~\ref{fig:timescales-sep} shows the median time until merger versus redshift for low-mass (green) and high-mass (pink) systems with scaled separations less than those listed above each plot.
+In the panel on the left, only pairs with a separation between $10\,\kpc-0.25\,\Rvir$ are utilized for the calculation. 
+We find that the median time until merger at $z=1$ is XX..XX for each scaled separation cut respectively, for both low-mass and high-mass samples.
+
+% redshift evolution of each subsample
+Similar to the static physical separation-selected subsamples in the top panel, the median time until merger starts off small at $z>4$, then increases to a peak around $z\sim0.5$, before quickly decreasing to 0 at $z=0$.
+This behavior also mimics that of the full sample.
+
+% impact of the separation criteria
+Additionally, the median time until merger at a given redshift increases for separation criteria that include a larger fraction of the virial radius, and thus larger separation pairs. 
+This again matches the behaviour of the static physical separation criteria shown in the top panel.
+
+% low mass vs high mass
+However, quite unlike the subsamples in the top panels, the scaled separation-selected subsamples yield nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts, regardless of what fraction of the virial radius is selected.
 
@@
 
-% \begin{figure*}[htb]
-%     \centering
-%     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsz.png}
-%     \caption{}
-%     % \label{fig:low-vs-high}
-% \end{figure*}
-% \begin{figure*}[htb]
-%     \centering
-%     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsz-2d.png}
-%     \caption{}
-%     % \label{fig:low-vs-high}
-% \end{figure*}
 \begin{figure*}[htb]
@@
 
-\section{Discussion}
+\section{Discussion} \label{sec:discussion}
 
@@
 
+\subsection{Results}
+
+Why did we pick the separation cuts that we did? (motivated already in Chambe+ 2024, restate here about close pair definitions from lit.)
+
+For a given static separation selection (particularly <50, <70, and <100), the merger timescales of low-mass and high-mass systems disagree by ~0.5\Gyr at almost all redshifts.
+
+Scaled separation criteria are akin to picking the same fraction of the total volume of a FoF group, regardless of what redshift it is at, not what mass it has.
+
+\subsection{What does it mean that the Time until Merge plot has the function form that it does?}
+Upper bounds of the hubble time and the lookback time as functions of z.
+
 % \subsection{Merger Rates for Dwarf Pairs}
@@
 \section{Summary and Conclusions}
+% what we did and how
 
-% To do:
-% -Sample: starting with FoF group cut and always show the total sample (merged and unmerged) since the unmerged sample is so small
-% -Keep Fig 1: (two panels) add in a few more orbits to show the diversity and then have a panel to show one example where you shade everything below 100 kpc to show how the total time at S is computed
-% -Keep Figure 2 as is 
-% -Remove Figure 3
-% -Figure 4: remove normalization of both samples so it's clear that the unmerged ones spend a lot of time at high separation and that they are a small percent of the fraction (put this in discussion or an appendix)
-% -Remove Fig 5 and state conclusions in words along with the FoF group selection (most pairs have been in the same group for 1.5 Gyr) (can quote percentages from a cumulative version of the plot, or sum)
-% -Figure 6: only show the *full* sample with the FoF group cut 
-% check the half mass radius of the most massive dwarfs in your sample and justify 10 kpc by saying you want to be able to resolve twice that
+% set of conclusions from results
+Our main findings are as follows: 
+\begin{itemize}
+    \item for a given separation (such as 100kpc), high-mass pairs have shorter merger timescales than low-mass pairs. 
+    \item the median time until merger for the full sample of low and high-mass pairs increases as $\sim(1+z)^{-1}$ from $z=6$ to $z\sim0.75$.
+it then decreases to 0 at low redshifts, with an upper bound set by the time remaining in the simulation before $z=0$.
+The decrease in the merger time at low redshifts is artificial and not representative of the true merger timescales that pairs at $z=0$ will have.
+    \item separation criteria that include larger separations will increase the median time until merger.
+    \item static physical separation-selected subsamples of low- and high-mass pairs have different median merger timescales.
+    \item scaled separation-selected subsamples yield nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts. 
+\end{itemize}
 
-% \section{Discussion}
+% set of conclusions for discussion
+- We find that the merger timescale of low-mass or high-mass merging pairs is sensitive to the selection criteria that are used to define the pair sample.
+- We recommend that future studies that seek to compare or compute pair fractions, merger fractions, merger timescales, or merger rates, define selected pair samples using separation criteria that evolve with redshift and mass.
 
@@
 % adfadfa
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
-% \end{figure}
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
-%     \caption{}
-% \end{figure}
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
-% \end{figure}
-
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_beforemerger.png}
-% \end{figure}
-% \begin{figure}[htb]
-%     \centering
-%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_untilmerger_phys.png}
-% \end{figure}
-
-% \begin{figure*}[htb]
-%     \centering
-%     \includegraphics[width=\columnwidth]{plots/4_timescales/timevssep_phys+co.pdf}
-%     \caption{}
-%     % \label{fig:phys-comoving}
-% \end{figure*}
-
-% \begin{figure*}[htb]
-%     \centering
-%     \includegraphics[width=\columnwidth]{lotz.png}
-%     \caption{}
-%     % \label{fig:phys-comoving}
-% \end{figure*}
 

### 2024-03-22 paper/main.tex
Updates from Overleaf

@@
     \centering
-    \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger_full_fit.png}
+    \includegraphics[width=0.9\textwidth]{plots/bet-on-it/3_time_til_merger_full_fit.png}
     \caption{\kc{TODO: add LookbackTime vs.
@@
 
+maybe a point: 
+For things that don't merge, what's... going on?
+
 % \subsection{Merger Rates for Dwarf Pairs}
@@
 % set of conclusions for discussion
-- We find that the merger timescale of low-mass or high-mass merging pairs is sensitive to the selection criteria that are used to define the pair sample.
+- We find that the merger timescale of low-mass or high-mass merging pairs is sensitive to the selection criteria that are used to define the pair sample.\\
+
 - We recommend that future studies that seek to compare or compute pair fractions, merger fractions, merger timescales, or merger rates, define selected pair samples using separation criteria that evolve with redshift and mass.

### 2024-03-22 paper/main.tex
Updates from Overleaf

@@
 Figure~\ref{fig:timescales} presents the median, and 1st and 3rd quartile spread, of the time until merger for merging pairs from $z=0-6$. 
+% low vs high
 Low-mass merger timescales (green) and high-mass merger timescales (pink) are roughly equivalent at all redshifts.
+
+% redshift evolution
 The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
+The abrupt decrease to 0 from $z=0.6$ to $z=0$ is an artificial rather than a true physical feature, and is caused by the lookback time of the simulation becoming shorter than the merger timescales of more and more pairs.
+The lookback time of the simulation as a function of redshift is given by the dotted black line on the left of Fig.~\ref{fig:timescales}.
+At $z=0$, the number of pairs that will merge is 0, since we cannot track the orbits of the subhalos into the future.
 
-\kc{left off here}
-
-% Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs.
+In addition, we show 
+\begin{equation}
+\frac{2}{5}*H(z)^{-1}
+\end{equation}
+given the Hubble parameter $H(z)$ as the dashed black line that sits above the median time until merger.
+We also show the age of the simulation as a function of redshift (dashed grey).
+% rest of stuff to talk about with this plot is discussion (i think) like what does this imply about the merger timescales of low vs high mass pairs?
 
@@
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
     \centering
-    \includegraphics[width=0.9\textwidth]{plots/bet-on-it/3_time_til_merger_full_fit.png}
-    \caption{\kc{TODO: add LookbackTime vs.
-Z line, add Hubble time line} The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs.
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale.png}
+    \caption{The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs.
 Shaded regions represent the 1st and 3rd quartile spread on the median.
-Pairs at each redshift are selected if they have experienced first infall.
+Pairs at each redshift are selected if they have experienced first infall and are pre-merger.
     % 
-    The black dashed line shows the Hubble time, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
-    The dotted grey line shows the time to $z=0$ (the Lookback Time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
+    The black dashed line shows a fraction of the Hubble time as a function of redshift, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
+    The dotted black line shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
     The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
     \label{fig:timescales}
-\end{figure*}
+\end{figure}
 

### 2024-03-23 paper/main.tex
edited sections 2-4 based on ekta comments, added outline to discussion section

@@
 
-Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
-From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.
+Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated as in \citet{Chamberlain2024}.
+From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.
 
@@
 We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
-Rather than only considering pairs for $z<4$, we will collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
-A brief version of the selection routine is transcribed here for completeness, and note that more detail and discussion of the choices for selection criteria are in Sec.~2 of~\citet{Chamberlain2024}.
+We collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
+A brief version of the selection routine is transcribed here for completeness, and we refer readers to Sec.~2 of~\citet{Chamberlain2024} for more detail and a discussion of the selection criteria choices.
 
@@
 at the snapshot of consideration. 
-For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to trace their mass history~\citep{RG2015}, and find the peak halo mass of each subhalo.
+For TNG100, this ensures that subhalos are resolved into over 130 particles which should be robustly identified in the SUBFIND and SUBLINK catalogs.
+For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.
 
-Stellar masses are then assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
-The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo, so that the abundance matching prescription is of the form $\ms{}=f(\Mpeak,z)$, with $z$ the redshift of consideration.
+Stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
+The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo via the abundance matching prescription  $\ms{}=f(\Mpeak,z)$, where $z$ is the redshift of consideration.
 
@@
 
+A subhalo can belong to different pairs at different redshifts. 
+For example, the primary of a pair that merges at $z=2$ can be selected with a different secondary at $z=1$, constituting a new pair. 
+More detail regarding the uniqueness of pairs and orbits is discussed in~\ref{sec:methods-unique}.
+
 % i am removing the 10kpc component of this because I think it'd just be easier to handle those cases in the 
@@
 % Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).
-
+\label{subsec:mergers}
 The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift.
@@
 The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
-The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshot, if it has one. 
+The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one. 
 % By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.
 
-A pair will be classified as a merger if the primary and secondary subhalo share the same \texttt{DescendantID} at the same redshift. 
-This means that the primary and secondary subhalo will merge, producing a single descendant subhalo, by the following redshift from the simulation.
+In this analysis, a pair is classified as a merger if the primary and secondary subhalo share the same \texttt{DescendantID} at the same redshift. 
+This means that the primary and secondary subhalo have merged such that SUBFIND can no longer distinguish them as two separate halos, therefore yielding a single descendant subhalo.
 If a pair consists of a primary and secondary that never share a \texttt{DescendantID} at a given redshift, the pair is defined as a `non-merger.'
@@
 For each merging pair, we define the `merger redshift' as the redshift that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
-For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place some time between $z=2$ and $z=1.9$, which is the next redshift in the simulation.
-We take $z=1.9$ to be the merger redshift.
+For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place some time between $z=2$ and $z=1.9$, which is the next redshift corresponding to a snapshot in the simulation.
+In this example, we take $z=1.9$ to be the merger redshift.
 
@@
 We track the positions of both the primary and secondary subhalo at each redshift and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
-In cases where the primary or secondary does not have a defined position at a given redshift, we set the separation at that snapshot to NaN.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs.
+In cases where the primary or secondary does not have a defined position at a given redshift, we do not compute a value for the separation at that redshift.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs.
 The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be evaluated before and after the skip occurs.
@@
 
-In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot.
+In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift.
 We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
@@
 \subsubsection{Defining Post-Infall}
-While the orbit of a pair may be calculated at very early times, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group.
+While the orbit of a pair may be calculated at very early times, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they belong to the same FoF group.
 Specifically, we will consider only the ``post-infall" part of each pair's orbit.
@@
 Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$.
-The top panel shows the orbits of 5 pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo.
-Even amongst galaxies of the same approximate mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\Gyr$ from first infall to merger.
+The top panel shows the orbits of five pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo.
+Even amongst galaxies of the same approximate mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\,\Gyr$ from first infall to merger,  indicated by the triangles.
 
-Some secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo. 
-Because all of our mergers occur when the subhalos are in the same FoF group, the
-post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same.
+For some pairs, the secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo than their primary.
+Because all of our mergers occur when the subhalos are in the same FoF group by definition, in these cases, we use the full orbit starting when the subhalos are first identified as members of the same FoF group through when they merge, as described in Section \ref{subsec:mergers}.
+This definition is more robust for considering the full interaction timescales of merging pairs.
+%the post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same.
+}
 % a variety of different orbits
 
-The middle panel of Fig.~\ref{fig:example-orbits} (``Potential Future Mergers") shows the orbits for 3 pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a few Gyr past the end of the simulation.
+The middle panel of Fig.~\ref{fig:example-orbits} (``Potential Future Mergers") shows the orbits for three pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a few Gyr past the end of the simulation.
 These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
-The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with many close encounters throughout, and 3 close passages in just the past $2\,\Gyr$.
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with many close encounters throughout, and three close passages in just the past $2\,\Gyr$.
 
-The bottom panel shows the orbits of "Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
+The bottom panel shows the orbits of ``Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
 Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.
@@
 \subsubsection{Uniqueness of orbits}
-A given pair from the \paircat{} may pass our selection criteria at many redshifts. 
-Since we collect the full orbit (the separation at every redshift where the primary and secondary both exist) for each pair and each redshift, the orbits that are collected for pairs that pass our selection criteria at more than one redshift will all result in identical orbits.
-% However, we only want to keep a single instance of any given orbit in our catalog.
+\label{sec:methods-unique}
+Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists.
+However, we only want to keep a single instance of any given orbit in our catalog so we can track the merger timescale of the pair.
 
@@
 The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the Sublink catalogs, and is unique for each pair of halos. 
-After each pair is assigned a unique pairkey, we save the orbit as a function of time only once for a pair that otherwise would have been double/multi-counted in our orbit catalog.
+After each pair is assigned a unique pairkey, we only keep one instance of an orbit per pair to avoid double/multi-counting in our orbit catalog.
 
@@
 For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
-In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are retained in the orbit catalog, since each orbit is unique.
+In this case, both orbits (of the original low-mass pair and the new pair which includes the primary subhalo from the previous merger) are retained in the orbit catalog, since each orbit is unique.
 
-Removing multiplicity decreases the raw number of orbits collected by collecting an orbit for each pair from the base sample. 
 The total number of pairs (including double counting the same pair at multiple redshifts) is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
-However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits.
-
-% For $n_k$ the number of pairs selected at snapshot $S_k$ via the selection criteria/from the base sample, the total number of pairs is given by
-% \begin{equation*}
-%     \sum_{k=0}^{99}\, n_k
-% \end{equation*}
-% The total number, and the total number of high-mass pairs is .
-% After selecting only unique pairs, there remain
+However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits that each correspond to a unique pair of galaxies.
 
@@
 \section{Pair Sample Properties}
+
 \begin{figure}[htb]
@@
     \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
-    The solid lines show the total number of orbits the are post-infall at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
+    The solid lines show the total number of pairs the are post-infall at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
     We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
-    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
+    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.\kc{change ylabel to "pairs"}
     }
@@
 
-\subsection{Number of Orbits}
-We calculate the number of orbits (or orbits in progress) by selecting all orbits at a given redshift that are post-infall and either pre-merger or non-mergers.
-The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
-For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
-A single pair (or equivalently, the pair's orbit) orbit may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of orbits at all redshifts from $z=0-3$.
+\subsection{Number of Pairs}
+We calculate the number of pairs (i.e. orbits in progress) by selecting all pairs at a given redshift that are post-infall and have not merged.
+% The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
+%For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
+A single pair (or equivalently, the pair's orbit) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of pairs at all redshifts from $z=0-3$.
 
-Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
-Low-mass orbits (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
+Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass pairs as a function of redshift from $z=0-6$. 
+Low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $z=0-1$.
 The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
-The number of pairs that merge decreases to 0 at $z=0$ since 
-%pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and 
-many pairs that exist at low redshift will have merger timescales longer than the remaining time to the end of the simulation.
+The number of pairs that merge decreases to 0 at $z=0$ since many pairs that exist at low redshift will have merger timescales that span beyond the end of the simulation (i.e., $z=0$).
 
-% removed
-% The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met.
-% These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.
+When a pair experiences first infall and thus begins to contribute to the number of pairs at later redshifts, we refer to such as pairs being ``added to the sample".
+Likewise, mergers at a given redshift will decrease the number of pairs counted at that redshift, since the pair is then post-merger and thus not included in the count.
+
+% The difference between high and low mass
+The number of low-mass pairs continues to increase from $z=6$ until around $z=2$, at which point the number of added pairs at a given redshift slows. 
+The decrease in the number of pairs after $z=2$ means that the number of low-mass pairs that are merging at each redshift is larger than the rate at which pairs are being added to the sample.
+
+The number of high-mass pairs continues increasing until $z=1$, at which point it remains approximately constant from $z=0-1$ at $\sim1000$ pairs. 
+There is a slight decrease in the number of high mass pairs at the very lowest redshifts $z<0.1$, where the mergers begin to outnumber the added pairs at each redshift.
 
 \subsection{Merger Fraction}
-We calculate the merger fraction by dividing the number of merging orbits (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift. 
+We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift, including all pairs that do and do not merge.
 As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
@@
 The merger fraction for both mass ranges declines to $0$ at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
-At $z=0$, any remaining pairs are classified as non-mergers since they did not merge \textit{$z=0$}.
+
+% difference/similiarity between low and high
+The merger fraction of low-mass and high-mass pairs is the same from $z=0-0.5$ and $z>2.5$. 
+Between $z=0.5-2.5$, the high-mass merger fraction is higher than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift.
+This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$.
+When the number of pairs stops increasing and beings to decrease, more pairs that will not have time to merge before $z=0$ are being added, and many merging pairs are being removed from the sample, thus decreasing the merger fraction.
+Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.
 
@@
 % ###########################################################################################
-
-\section{Results: The Mass and Redshift Dependence of Merger Timescales}
-We have created catalogs of unique orbits for a large sample of isolated low-mass and high-mass pairs in the TNG100 simulation.
-Next, we will calculate the merger timescales, or time until merger, for all of the merging orbits in our catalog. 
-In Sec.~\ref{sec:results-timevsep}, we explore how the time until merger changes as a function of the separation of the pair at a variety of redshifts, and study how the time until merger changes between the low-mass sample and high-mass sample.
-In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all post-infall and pre-merger isolated pairs as a function of redshift. 
-We will additionally examine the merger timescale's dependence on a variety of separation criteria.
-
-In the following analysis, we will utilize the orbit catalog, which consists of the unique orbits for all low-mass and high-mass major isolated pairs from the~\citet{Chamberlain2024} study. 
-At a given redshift, we will include all merging pairs that are post-infall and pre-merger at that redshift in our analysis. 
-However, separation criteria will also be applied to each redshift independently, so a pair that is in the $10-50\,\kpc$ bin at one redshift will not be part of the sample used at redshifts where it's separation is $>50\,\kpc$.
-
 \begin{figure*}[htb]
     \begin{center}
-    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
+    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
     \caption{\kc{Add z=5, z=6 panels.
@@
 For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
-    \kc{add z=5, z=6}
+    \kc{add z=5, z=6, add horizontal line for the time remaining in the sim in panel 1}
     }
@@
     \begin{center}
-    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
+    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
     \caption{\kc{Add z=5, z=6 panels.
@@
 
+\section{Results: The Mass and Redshift Dependence of Merger Timescales}
+Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
+In Sec.~\ref{sec:results-timevsep}, we explore how the merger timescale changes as a function of the separation of the pair across redshifts, and how the merger timescales differ between the low-mass and high-mass pairs.
+In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all post-infall and pre-merger isolated pairs as a function of redshift. 
+We will additionally examine the merger timescale's dependence on a variety of separation criteria. 
+Separation criteria are applied to pairs at each redshift independently, such that a pair in the $10-50\,\kpc$ bin at one redshift will not be part of the sample used at redshifts where its separation is $>50\,\kpc$.
+
 \subsection{Separation Dependence of Merger Timescales}\label{sec:results-timevsep}
-\kc{note: this might go in the appendix as additional info}
 % how we calculate for this section
 We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
-At the given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each pair.
+Pairs with small separations may be undergoing a pericentric passage, or very close to merging if on a radial orbit.
+Binning the pairs by separation, we can study how the merger timescale changes for pairs at different points in their orbits.
 
 % low mass example
-For example, at $z=1.5$, there are 5,787 unique orbits that are post-infall and will merge prior to $z=0$.
-Of those, and 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 of those low-separation pairs will merge in the next $2\,\Gyr$.
-A compilation of this data is available in Table~\ref{tab:timevssepvsz}, where we list the fraction of post-infall and pre-merger pairs within 3 separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction within each of those separation bins that merge in 4 different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$, for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$..
+We bin our pair sample by merger timescale, in bins of $0.5\Gyr$, and separation in bins of width $10\,\kpc$.
+Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$. 
+The horizontal line in the first panel shows the time remaining in the simulation until $z=0$, above which no merging pairs exist.
 
-As an accompanying visual aid, figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs.
-separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$.
-Fig.~\ref{fig:timevssep-low} presents the distribution for low-mass subhalos, and shows that the time until merger for low-mass pairs is positively correlated with the separation of the pair at each redshift, but that the positive correlation is smaller at larger redshifts.
-In addition, the spread of times until merger and separations is smaller at higher redshift  than the spread at lower redshift, likely due to the size growth of halos over time. 
-% could make these same plots with scaled separation, then it would presumably show roughly the same spread for each redshift.
+Fig.~\ref{fig:timevssep-low} presents a heatmap distribution of the density of low-mass pairs in merger timescale-separation space. 
+We find that the time until merger for low-mass pairs is positively correlated with the separation of the pair at each redshift. 
+However, the slope of the correlation decreases at higher redshifts.
+Pairs at $z=4$ and separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.
 
-Likewise, fig.~\ref{fig:timevssep-high} presents the distribution for high-mass subhalos, where the trends between time until merger and separation are roughly the same. 
-At each redshift, there is a positive correlation between time until merger and separation, but the spread on either of those decreases with increasing redshift.
+In addition, the spread of times until merger and separations is smaller at higher redshift. 
+The spread of merger timescales at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$, and likewise for the separations spanned by pairs at both redshifts. 
+This is due to the growth of halo size over time, since more pairs will be post-infall at lower redshifts where the virial radius of the FoF groups is larger.
 
-These figures can be used to get approximate ranges of time until merger for a pair at a given separation and for a given redshift, and are useful for seeing the difference between the relationship between time until merger and separation for low-mass and high-mass pairs.
-\kc{what differences?}
+Likewise, Fig.~\ref{fig:timevssep-high} presents the distribution for high-mass subhalos.
+The trends between merger timescale and separation are roughly the same as low-mass pairs, with a positive correlation between the two parameters that decreases in slope from low redshift to high redshift.
+The high-mass pairs also have a larger spread in the distribution of merger timescales and separations at low redshift than at higher redshift.
+This is once again caused by the change in the FoF halo size over time.
+
 % 
 \input{table-septime.tex}
+A compilation of more specific data is available in Table~\ref{tab:timevssepvsz}.
+Here, we list the fraction of post-infall and pre-merger pairs within three separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction of pairs within each of those separation bins that merge in four different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$.
+We present these data for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$.
 
-\subsection{Redshift Dependence of Merger Timescales}\label{sec:results-timevredshift}
-% The behavior of time until merger as a function of redshift for low-mass and high-mass pairs:
-% 1 - the full sample 
-% 2 - the physical and static kpc separation-selected sample
-% 3 - the mass and redshift evolving scaled separation-selected sample
-We study the merger timescale of low-mass and high-mass pairs as a function of redshift. 
-In particular, we consider 3 different samples of the low- and high-mass systems. 
-First, we will consider the orbit sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that are post-infall and pre-merger for $z=0-6$. 
-We will then consider two different sets of separation criteria to create separation-selected subsamples, the first of which are physical separation criteria that are identically applied for both mass scales and all redshifts, and the second of which defines separation criteria that vary with mass and redshift.
-% We ascertain the impact of the selection criteria on these subsamples of the orbit catalog .....
-based on separation-selected pairs, using a set of separation criteria that we apply independently at each redshift.
-
-Both separation-selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \texttt{SUBFIND} catalogs.
-This lower separation criteria is also commonly applied to observationally-selected pairs in studies of merger fractions and merger rates (see Sec.~\ref{sec:discussion} for further detail).
-
-Throughout, medians at each redshift are shown by solid pink and green lines, while the 1st and 3rd quartile spread on the median is shown by the shaded regions.
-
-\subsubsection{Full sample}
-We calculate the time until merger for all post-infall and pre-merger pairs at each redshift from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
-We make no selection cuts based on pair separation, and thus include the full catalog of orbits at each redshift to calculate the merger timescale (once again, provided that they are post-infall and pre-merger at the given redshift).
-Essentially, we find the median time that any physically associated isolated pair will take to merge, provided it's a merging pair to begin with.
-
-Figure~\ref{fig:timescales} presents the median, and 1st and 3rd quartile spread, of the time until merger for merging pairs from $z=0-6$. 
-% low vs high
-Low-mass merger timescales (green) and high-mass merger timescales (pink) are roughly equivalent at all redshifts.
-
-% redshift evolution
-The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
-The abrupt decrease to 0 from $z=0.6$ to $z=0$ is an artificial rather than a true physical feature, and is caused by the lookback time of the simulation becoming shorter than the merger timescales of more and more pairs.
-The lookback time of the simulation as a function of redshift is given by the dotted black line on the left of Fig.~\ref{fig:timescales}.
-At $z=0$, the number of pairs that will merge is 0, since we cannot track the orbits of the subhalos into the future.
-
-In addition, we show 
-\begin{equation}
-\frac{2}{5}*H(z)^{-1}
-\end{equation}
-given the Hubble parameter $H(z)$ as the dashed black line that sits above the median time until merger.
-We also show the age of the simulation as a function of redshift (dashed grey).
-% rest of stuff to talk about with this plot is discussion (i think) like what does this imply about the merger timescales of low vs high mass pairs?
-
-\subsubsection{Separation Criteria: Static \& Physical }
-The first set of separation criteria will select only pairs at a given redshift that have separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$. 
-These separation criteria do not vary as a function of time or mass, and are thus applied equivalently to the low-mass and high-mass samples. 
-This set of static physical separation criteria produces 6 subsamples of merging orbits that are post-infall.
-
-Note that an orbit in the separation bin $10-50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts. 
-Rather, the separation criteria select pairs with separations in the given separation bin independently at each redshift.
-
-% figure description 
-The top panel of Fig.~\ref{fig:timescales-sep} shows the time until merger versus redshift for low-mass (green) and high-mass (pink) orbits that have separations less than the separation listed above each plot.
-% redshift evolution of each subsample
-We first note that the evolution of the time until merger is approximately the same for each separation-selected subsample, with a peak time at redshift $z\sim0.5$ present for all subsamples, and a rapid decline of the median time until merger at $z=0-0.5$.
-These traits were also present in Fig.~\ref{fig:timescales} where the median time until merger was computed for the entire sample, without a selection-criterion.
-As the selection-criteria gets larger and larger, the subsample will contain more and more of the full sample, eventually reconstructing the full sample plot with a large enough separation selection.
-
-% low mass vs high mass
-There is an offset between the median times for low-mass and high-mass pairs for each of the subsamples with maximum separations of $[50, 70, \mbox{and } 100]\,\kpc$. 
-The difference in the median time until merger is up to $0.8\,\Gyr$ longer for low-mass pairs than high-mass pairs at the same redshift. 
-The offset between the low-mass and high-mass merger times decreases for large enough physical separation criteria. 
-In the rightmost top panel, the median merger times overlap almost identically for $z=0-0.5$ and $z>1.5$.
-
-% impact of the separation criteria
-The median time until merger for pairs selected in smaller separation bins (such as $10-50\,\kpc$ and $10-70\,\kpc$) is lower at all redshifts for both low-mass and high mass pairs. 
-This follows directly from Sec.~\ref{sec:results-timevsep}, where we found that the time until merger is positively correlated with increasing separation for all pairs. 
-Thus, including more and more larger-separation systems in our merger time calculation will tend to increase the median time until merger.
-
-\subsubsection{Separation Criteria: Scaled Separation}
-The second set of selection criteria will select pairs at a given redshift in 6 different scaled separation bins, where the scaled separation is written as a fraction of the virial radius of the pair's FoF group, $\Rvir$ (see Sec.~\ref{sec:methods} for more details).
-The scaled separation-selected pairs have separations greater than $10\,\kpc$ and less than $<[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$.
-
-A scaled separation criterion scales with both redshift and mass (since the virial radius scales with redshift and mass), and produces 6 subsamples of merging orbits that are post-infall.
-As mentioned in \chambe{}, the median virial radius for low-mass FoF groups at $z=[0,1,2,3,4]$ is approximately $[134, 85, 59, 43, 33]\,\kpc$, and for high-mass FoF groups is approximately $[348, 206, 134, 97, 76]\,\kpc$.
-Thus, choosing a scaled separation cut of $<1\,\Rvir$ at $z=1$ will select high-mass pairs with separations $\lesssim205\,\kpc$ and low-mass pairs with separations $\lesssim85\,\kpc$.
-
-As in the previous section, the selection criteria are applied at each redshift independently. 
-This means that the time until merger of a pair at a given redshift will only be included in the calculation of the median merger time if its separation is within the given separation bin at that redshift.
-
-% figure description 
-The bottom panel of Fig.~\ref{fig:timescales-sep} shows the median time until merger versus redshift for low-mass (green) and high-mass (pink) systems with scaled separations less than those listed above each plot.
-In the panel on the left, only pairs with a separation between $10\,\kpc-0.25\,\Rvir$ are utilized for the calculation. 
-We find that the median time until merger at $z=1$ is XX..XX for each scaled separation cut respectively, for both low-mass and high-mass samples.
-
-% redshift evolution of each subsample
-Similar to the static physical separation-selected subsamples in the top panel, the median time until merger starts off small at $z>4$, then increases to a peak around $z\sim0.5$, before quickly decreasing to 0 at $z=0$.
-This behavior also mimics that of the full sample.
-
-% impact of the separation criteria
-Additionally, the median time until merger at a given redshift increases for separation criteria that include a larger fraction of the virial radius, and thus larger separation pairs. 
-This again matches the behaviour of the static physical separation criteria shown in the top panel.
-
-% low mass vs high mass
-However, quite unlike the subsamples in the top panels, the scaled separation-selected subsamples yield nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts, regardless of what fraction of the virial radius is selected.
-
-% ###########################################################################################
-% Results II 
-% ###########################################################################################
+Data from the table can be interpreted in the following way: at $z=1.5$, there are 5,787 pairs that are post-infall and merge prior to $z=0$. 
+Of those, 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 of those $10-50\,\kpc$ pairs will merge in the next $2\,\Gyr$.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale.png}
+    \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale_mod.png}
     \caption{The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs.
@@
 \end{figure*} 
+\subsection{Redshift Dependence of Merger Timescales}\label{sec:results-timevredshift}
+% The behavior of time until merger as a function of redshift for low-mass and high-mass pairs:
+% 1 - the full sample 
+% 2 - the physical and static kpc separation-selected sample
+% 3 - the mass and redshift evolving scaled separation-selected sample
+In this subsection, we study the merger timescale of low-mass and high-mass pairs as a function of redshift. 
+First, we consider the pairs sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that are post-infall and pre-merger for $z=0-6$. 
+We will then consider two different sets of separation criteria to create separation-selected subsamples.
+Since pair samples are typically picked via separation criteria, in both simulations and observations, we study the impact of different separation-based selection criteria on inferred merger timescales.
 
+Throughout, medians at each redshift are shown by solid pink and green lines, while the 1st and 3rd quartile spread on the median is shown by the shaded regions.
+
+\subsubsection{Full sample}
+    We calculate the time until merger for all post-infall and pre-merger pairs at each redshift in the simulation from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift.
+    We include the full catalog of orbits at each redshift to calculate the merger timescale. 
+    % why do we want to consider the entire population of halos? why interesting?
+
+    Figure~\ref{fig:timescales} presents the median of the time until merger for post-infall merging pairs from $z=0-6$. 
+    The 1st and 3rd quartiles are indicated with shaded regions. 
+    % low vs high
+    Low-mass merger timescales (green) and high-mass merger timescales (pink) are roughly equivalent at all redshifts. 
+    Since we have used the same selection procedure for both high-mass and low-mass pairs, assuming that the pair has experienced first infall, which implies that mergers don't proceed in fundamentally different ways at different mass scales.
+    We explore this point further in Sections~\ref{sec:results-phys} and~\ref{sec:results-scal}.
+
+    % redshift evolution
+    The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
+    The abrupt decrease to 0 from $z=0.6$ to $z=0$ is artificial rather than a true physical feature.
+    The majority of low redshift pairs will have merger timescales that span past the end of the simulation, since the time until the end of the simulation gets increasing small for smaller redshift. 
+    However, as $z\to0$, there is a also an increasingly small fraction of total pairs that still merge before $z=0$, as seen in Fig~\ref{fig:numorbits}. 
+    These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$. 
+    This is shown by the dotted black line on the left of Fig.~\ref{fig:timescales}, which shows the lookback time of the simulation as a function of redshift.\footnote{The lookback time at a given redshift $z_n$ is equivalent to the time elapsed between $z_n$ and $z=0$.
+Thus, merging pairs cannot have merger timescales larger than the lookback time of the simulation at a given redshift.} % this could go in the methods section, if we think it's more necessary there.
+I don't mind it here for more detail
+
+    In addition, we show 
+    \begin{equation}
+    \frac{2}{5}*H(z)^{-1}
+    \end{equation}
+    given the Hubble parameter $H(z)$ as the dashed black line that sits above the median time until merger.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation.
+Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
+    We also show the age of the simulation as a function of redshift (dashed grey). 
+    These lines show XXX...
+    % so far missing the WHY we are comparing to the hubble time or age vs z
+    % look at Hopkins dynamical friction/merger timescale plot.
+
+\subsubsection{Physical Separation Selected Pairs }
+\label{sec:results-phys}
+    Both separation selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \texttt{SUBFIND} catalogs.
+    This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates (see Sec.~\ref{sec:discussion} for further detail).
+
+    The first set of separation criteria will select only pairs at a given redshift that have separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, to create six pair subsamples.
+    These separation criteria do not vary as a function of time or mass, and are applied equivalently to the low-mass and high-mass samples at all redshifts. 
+    We note that an orbit in the separation bin $10-50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts. 
+    Rather, the separation criteria are applied independently at each redshift.
+
+    % figure description 
+    The top panel of Fig.~\ref{fig:timescales-sep} shows the time until merger versus redshift for each of the low-mass (green) and high-mass (pink) pair subsamples with separations less than the physical separation listed above each column.
+    % redshift evolution of each subsample
+    We first note that the merger timescale peaks around redshift $z\sim0.5$ for all low-mass and high-mass subsamples. 
+    In addition, all subsamples show the same decline of mean merger time as $z\to0$, as discussed in the previous section. 
+    These traits were also present in Fig.~\ref{fig:timescales}, so they are features that are roughly independent of a separation criteria.
+    As the selection criteria gets larger, the subsample contains more of the full sample, eventually including the full sample with a large enough separation selection.
+
+    % low mass vs high mass
+    The subsamples with maximum separations of $[50, 70, \mbox{and } 100]\,\kpc$ result in median merger timescales that are higher for low-mass pairs than high-mass pairs at all redshifts. 
+    The difference in the median time until merger is up to $0.8\,\Gyr$ longer for low-mass pairs than high-mass pairs at the same redshift for the same separation criteria. 
+    The offset between the low-mass and high-mass merger timescale decreases for large enough physical separation criteria. 
+    In the rightmost top panel, the median merger times overlap almost identically for $z=0-0.5$ and $z>1.5$, almost mimicking our findings from the full sample analysis.
+
+    % impact of the separation criteria
+    When the maximum separation is increased (from left to right), more of the full sample is being used in the calculation for the mean merger timescale. 
+    Indeed, the merger timescale for both low-mass and high-mass pairs increases with an increasing maximum separation cut.  
+    This follows directly from Sec.~\ref{sec:results-timevsep}, where we found that the time until merger is positively correlated with increasing separation for all pairs. 
+    Thus, including a higher fraction of larger-separation systems in our merger time calculation tends to increase the median time until merger.
+
+    The median merger timescale for low-mass pairs does not change significantly for any separation criteria $>150\,\kpc$. 
+    On the otherhand, high-mass pairs tend to see an increase in the merger timescale for each larger separation cut. 
+    This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
+    High-mass pairs tend to have higher separations than low-mass pairs, as show in~\ref{sec:results-timevsep}.
+
+\subsubsection{Scaled Separation Selected Pairs}
+\label{sec:results-scal}
+    We calculate the merger timescales for six additional subsamples of pairs, this time applying separation criteria that scale with both redshift and mass.
+    The scaled separation, which we define in~\ref{sec:methods},  is the separation of a pair divided by the virial radius of the pair's FoF group, $\Rvir$. 
+    The scaled separation selected pairs have separations greater than $10\,\kpc$ and less than $<[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$. 
+    As in the previous section, the selection criteria are applied at each redshift independently.
+
+    As detailed in \chambe{}, the median virial radius for low-mass FoF groups at $z=[0,1,2,3,4]$ is approximately $[134, 85, 59, 43, 33]\,\kpc$, and for high-mass FoF groups is approximately $[348, 206, 134, 97, 76]\,\kpc$.
+    Thus, choosing a scaled separation cut of $<1\,\Rvir$ at $z=1$ will select high-mass pairs with separations $\lesssim205\,\kpc$ and low-mass pairs with separations $\lesssim85\,\kpc$.
+
+    % figure description 
+    The bottom panel of Fig.~\ref{fig:timescales-sep} shows the median time until merger versus redshift for low-mass (green) and high-mass (pink) pair subsamples with scaled separations less than those listed at the top of each plot.
+    We find that the median time until merger at $z=1$ is $[0.64, 0.94,1.25,1.51-1.58,1.76-1.91,1.91-2.05]\,\Gyr$ for each scaled separation cut respectively, for both low-mass and high-mass samples.
+%\kc{gonna have to triple check this cause I get the SAME median time til merger for low and high mass for Rvir 0.25, 0.5, and 1, down to like 14 decimals...}
+
+    % redshift evolution of each subsample
+    Similar to the static physical separation-selected subsamples in the top panel, the median time until merger starts off small at $z>4$, then increases to a peak around $z\sim0.5$, before quickly decreasing to 0 at $z=0$.
+    Additionally, the median time until merger at a given redshift increases for separation criteria that include a larger fraction of the virial radius, and thus larger separation pairs.
+
+    % low mass vs high mass
+    However, quite unlike the physical separation selected subsamples in the top panels, using a scaled separation criteria results in nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts.
+    The low-mass and high-mass merger timescales increase with each larger scaled separation cut, but recover the behavior of the full sample by $<2\,\Rvir$.
+
+    In addition, the slope of the time until merger as a function of redshift for $z>0.5$ is steeper for high-mass pairs in the scaled separation subsamples than the physical separation subsamples.
+    This is especially noticeable in the first panel of both the top and bottom rows, where the merger timescale of high-mass pairs is approximately flat for $<50\,\kpc$.
+
+% ##################
+% Discussion section
 \section{Discussion} \label{sec:discussion}
+We have calculated the merger timescales, or time until merger, for a sample of low-mass and high-mass pairs that have experienced first infall but not yet merged, to ensure that they are physically collocated.
+In this section, we will explore the implications and broader impacts of our findings, and put our study in context with previous work. 
+%In Sec.~\ref{}, etc...
 
-\subsection{About the number of pairs}
-Can we say anything about the REASON the unique pairs decreases at low z?
-Does this plot mean that it's only because things merge and so we lose them?
-And not that it's driven by mass-accretion thing?
-I think so.
+\begin{itemize}
+    % \item Implications for 
+    \item Sec 1: Comparison to previous works
 
-Need to be careful about phrasing since we didn't actually quantify the number of pairs that leave the sample due to having too large of masses at later snapshots.
+    Subsec 1: merger fraction studies
+    %Why did we pick the separation cuts that we did? 
+    \begin{itemize} 
+        \item What other studies have looked at merger fractions in hydro simulations, specifically? 
+        \item Have they studied merger fractions as a function of mass and redshift? What did they find?
+        \item What is unique about our study?
+        \item Were low-mass pairs expected to have such high merger rates in isolated environments? 
+        \item Are comparisons with previous works fair, or are they not really comparable?
+        \item specific papers: Ventou 2017
+    \end{itemize}
 
-\subsection{About the merger fraction}
-Were low-mass pairs expected to have such a high merger rate in isolated environs? 
-How do these merger fractions compare to other literature results? 
-If we can make comparisons, discuss the differences. 
-If we cannot make comparisons, explain why our study is done in a way that makes for comparisons unfair/undoable .
+    Subsec 2: existing merger timescale studies
+    \begin{itemize} 
+        \item What other studies have looked at merger timescales in hydro simulations? 
+        \item Have they studied merger timescales as a function of mass and redshift? 
+        \item What is unique about our study?
+        \item Do our findings agree? 
+        \item specific papers: Hopkins 2007,
+    \end{itemize}
+    \item Sec. 2: Implications for observational merger fractions and merger rates
+    \begin{itemize}
+        \item How did we motivate separation criteria?(motivated already in Chambe+ 2024, restate here about close pair definitions from lit.)
+        \item Pair fractions can vary wildly depending on separation criteria
+        \item This is important too for merger rates, since they are a function of pair fractions and merger timescales, and you should really be selecting pairs at each stage as self-consistently as possible
+        \item What is our suggestion moving forwards?
+        \item specific papers: Lotz 2011, 
+    \end{itemize}
+    \item Sec 3: The functional form of the merger timescale
+    \begin{itemize}
+        \item What can we say about the fact that the merger timescales follow a $(1+z)^{-1}$ relation at $z>0.5-1$? Physical explanation for this behaviour? 
+        \item Scaled separation criteria are akin to picking the same fraction of the total volume of a FoF group, regardless of what redshift it is at, not what mass it has, so it may not be surprising that we got what we did in re: the scaled separation being better, but finding the SAME merger timescales for hi and low is interesting!
+        \item Maybe something to consider in this section: For things that don't merge, what's... going on? 
+        \item relevant papers: 
+    \end{itemize}
+\end{itemize}
 
-\subsection{Results}
-
-Why did we pick the separation cuts that we did? (motivated already in Chambe+ 2024, restate here about close pair definitions from lit.)
-
-For a given static separation selection (particularly <50, <70, and <100), the merger timescales of low-mass and high-mass systems disagree by ~0.5\Gyr at almost all redshifts.
-
-Scaled separation criteria are akin to picking the same fraction of the total volume of a FoF group, regardless of what redshift it is at, not what mass it has.
-
-\subsection{What does it mean that the Time until Merge plot has the function form that it does?}
-Upper bounds of the hubble time and the lookback time as functions of z.
-
-maybe a point: 
-For things that don't merge, what's... going on?
-
-% \subsection{Merger Rates for Dwarf Pairs}
-% % - What combination of values best reproduces the merger fraction acquired from the sims?
-% % Do you need the scale of merger probability as in Ventou to get the match?
-% % Vicente & Casteels 2014 comparison (pair fractions plot)
-% \subsection{Pairs of Dwarfs That Do Not Merge}
-% % can talk about the unmerged fraction merger rates here
-% \subsection{Choosing Samples at Lower and Higher Redshift}
-% % can also compare to samples to merger fractions and merger rates chosen at z=1 and z=2 (plots you have already)
-% \subsection{Comparing with Merger Rates for Massive Galaxy Pairs}
-% % compare to massive pairs in both obs window and merger rate
-% % compare back to the assumption of isolation
+Bring up assumption of isolation criteria and it's drawbacks and strengths.
+Reiterate that our findings are for isolated pairs.
 
@@
 % what we did and how
+$<$Longer summary of what we did here$>$
 
@@
 \begin{itemize}
-    \item for a given separation (such as 100kpc), high-mass pairs have shorter merger timescales than low-mass pairs. 
+    \item for a given physical separation (such as 100kpc), high-mass pairs have shorter merger timescales than low-mass pairs. 
     \item the median time until merger for the full sample of low and high-mass pairs increases as $\sim(1+z)^{-1}$ from $z=6$ to $z\sim0.75$.
@@
 
+- Why is this important?
+Cause future studies are gonna get so many pairs, especially low mass, and now we have a framework to compare low and high mass pairs in a variety of ways, from pair fractions to merger fractions to merger timescales!
+Woo!
+
 % ** 

### 2024-03-25 paper/main.tex
added some notes for intro based on readings

@@
 
+% defining colors for personalized comments
 \definecolor{pink}{RGB}{232,132,161}
 \definecolor{yellow}{RGB}{255,213,0}
+\newcommand{\kc}[1]{\textcolor{yellow}{\textbf{kc: #1}} }
+\newcommand{\ep}[1]{\textcolor{pink}{\textbf{e: #1}} }
 
-\newcommand{\kc}[1]{\textcolor{yellow}{\textbf{kc: #1}} }
-\newcommand\shadetext[2][]{%
-  \setbox0=\hbox{{\special{pdf:literal 7 Tr }#2}}%
-  \tikz[baseline=0]\path [#1] \pgfextra{\rlap{\copy0}} (0,-\dp0) rectangle (\wd0,\ht0);% 
-  }
-\newcommand{\ep}[1]{\shadetext[left color=blue, right color=red, middle color=lime, shading angle=45]{\textbf{e: #1}} }
-% \newcommand{\ecite}[1]{\textcolor{pink}{\textbf{: #1}} }
-% \newcommand{\e}[1]{\textcolor{yellow}{\textbf{: #1}} }
+% defining affiliations
+\newcommand{\affuofa}{University of Arizona, 933 N. Cherry Ave,
+    Tucson, AZ 85721, USA}
+\newcommand{\affuofu}{Department of Astronomy, University of Utah, Salt Lake City, UT 84112, USA}
 
-\newcommand{\remove}[1]{\textcolor{red}{#1}}
-\newcommand{\add}[1]{\textcolor{green}{#1}}
-
-\newcommand{\mlg}{\ensuremath{M_{\rm LG}}}
-\newcommand{\mmto}{\ensuremath{M_{\rm M31}}}
-\newcommand{\mmw}{\ensuremath{M_{\rm MW}}}
-\newcommand{\vtan}{\ensuremath{v_\textrm{tan}}}
-\newcommand{\vrad}{\ensuremath{v_\textrm{rad}}}
-\newcommand{\ms}[1]{\ensuremath{M_{*{#1}}}}
-\newcommand{\reflabel}[1]{\ensuremath{^{\mbox{\scriptsize{#1}}}}}
-\newcommand{\scsep}{\ensuremath{\rm r_{sep}/r_{vir}}}
-\newcommand{\scvel}{\ensuremath{\rm v_{rel}/v_{vir}}}
-
-\newcommand{\paircat}{\textit{Full Pair Catalog}}
-\newcommand{\Rvir}{\ensuremath{\rm R_{vir}}}
-\newcommand{\Rphys}{\ensuremath{\rm R_{phys}}}
-\newcommand{\Rsc}{\ensuremath{\rm R_{sc}}}
-\newcommand{\rsep}{\ensuremath{\rm r_{sep}}}
-
+% reference to previous paper
 \newcommand{\chambe}{\citet{Chamberlain2024}}
@@
 % Style tweaks
-% \renewcommand{\twocolumngrid}{\onecolumngrid}
-% \setlength{\parindent}{1.1\baselineskip}
-% \sloppy\sloppypar\raggedbottom\frenchspacing
+\sloppy\sloppypar\raggedbottom\frenchspacing
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\shorttitle{Merger Timescales in TNG100}
-\shortauthors{Chamberlain et al.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Inputs for path to plots and math defs
 \graphicspath{{./}{../plots/}}
@@
 
-% Affiliations
-\newcommand{\affuofa}{University of Arizona, 933 N. Cherry Ave,
-    Tucson, AZ 85721, USA}
-
-\newcommand{\affuofu}{Department of Astronomy, University of Utah, Salt Lake City, UT 84112, USA}
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
@@
 }
+\shorttitle{Merger Timescales in TNG100}
 
@@
 
+\shortauthors{Chamberlain et al.}
+
 \begin{abstract}
     % Statement
-Evolutionary differences between low-mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high-mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) galaxy pairs have been studied in detail for the last few decades, with studies finding differences in star formation rates, gas fractions, and dark matter halo profiles.
+    % Evolutionary differences between low-mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high-mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) galaxy pairs have been studied in detail for the last few decades, with studies finding differences in star formation rates, gas fractions, and dark matter halo profiles.
+% could remove this line
     % Problem 
-Pair fraction, merger fraction, and merger rate studies of high-mass pairs have been studied extensively, both observationally and theoretically. 
+    The merger rates of high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively, both observationally and theoretically. 
+    Since low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) are comparatively under studied, it is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems.
     However, it is unclear if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift.
@@
 \end{abstract}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-% $z=1.5$ corresponds to snapshot \#40 in the TNG sim
 % Introduce RG15, Lotz, etc.
+% \input{paper/notes_for_intro}
 
@@
 %recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
-We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
+We utilize the group catalogs, produced by the \subfind{} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \sublink{} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
 The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).
@@
 \subsection{Pair sample}
-We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
-We collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
-A brief version of the selection routine is transcribed here for completeness, and we refer readers to Sec.~2 of~\citet{Chamberlain2024} for more detail and a discussion of the selection criteria choices.
+We begin with an extension of the \paircat{} described in \chambe{}, which consists of a collection of isolated galaxy pairs at each snapshot in TNG100. 
+We collect our base sample at each redshift in the simulation. 
+A brief version of the selection routine is transcribed here for completeness, and we refer readers to Sec.~2 of \chambe{} for more detail and a discussion of the selection criteria choices.
 
@@
 at the snapshot of consideration. 
-For TNG100, this ensures that subhalos are resolved into over 130 particles which should be robustly identified in the SUBFIND and SUBLINK catalogs.
-For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.
+For TNG100, this ensures that subhalos are resolved into over 
+$\sim$100 particles, enough to robustly identify gravitationally bound subhalos in the \subfind{} and \sublink{} catalogs.
+For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \sublink{} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.
 
@@
 
-In \citet{Chamberlain2024}, the abundance matching prescription was sampled 1000 times for each subhalo to account for the spread in the Stellar Mass -- Halo Mass relation. 
-However, for the purposes of the present study, we only use the stellar masses given by the median of the abundance matching relationship. 
-This choice permits a roughly 1:1 monotonic relationship between stellar mass and peak halo mass, so that taking subsamples of the lowest stellar mass objects will result in a subsample with lower peak halo mass as well.
-% is this a good enough reason?
-Additionally, ~\cite{Chamberlain2024} found that the spread in the number of pairs selected at a given redshift is small, such that we expect the spread on the timescale analysis to be dominated by the redshift spacing of the simulation.
+In \chambe{}, the abundance matching prescription was sampled 1000 times for each subhalo to account for the spread in the Stellar Mass -- Halo Mass (SMHM) relation. 
+For the present study, we only use the stellar masses given by the median of the abundance matching relationship. 
+We expect that the spread of merger timescales from different orbital configurations, as well as the redshift spacing of the TNG100 snapshots, will dominate over uncertainties from the number of pairs in the catalog, which varied by $\sim3\%$ (as shown in Fig.~1 in \chambe{}).
+% Since we will be following the orbits of selected pairs both backwards and forwards in time, the
 
@@
 
-% i am removing the 10kpc component of this because I think it'd just be easier to handle those cases in the 
+% I removed the 10kpc component of this because I think it'd just be easier to handle those cases in the results sections or orbits section
 % Finally, we will only consider pairs that have physical separations of $\rsep>10\,\kpc$ at the snapshot of selection. 
@@
 
-\subsection{Mergers}
-% Plot \#1: Merger fraction either before or after orbits
-% Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).
-\label{subsec:mergers}
-The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift.
-The \texttt{SUBLINK} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.
+\subsection{Mergers} \label{subsec:mergers}
+Here we describe how subhalo mergers are identified in the TNG simulations. 
+Note that this definition is specific to subhalo mergers and therefore results are not guaranteed to hold for galaxy mergers, especially in cases where subhalos and galaxies have different centers of mass.
+% might  note that this is discussed in more deets in discussion - I think Patton 2024 addresses this a bit?
+The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \subfind{} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift.
+The \sublink{} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.
 
-We utilize the \texttt{DescendantID} field of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
+We utilize the \texttt{DescendantID} field of the \sublink{} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
 The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
-The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one. 
-% By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.
+The \sublink{} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one.
 
 In this analysis, a pair is classified as a merger if the primary and secondary subhalo share the same \texttt{DescendantID} at the same redshift. 
-This means that the primary and secondary subhalo have merged such that SUBFIND can no longer distinguish them as two separate halos, therefore yielding a single descendant subhalo.
+This means that the primary and secondary subhalo have merged such that \subfind{} can no longer distinguish them as two separate halos, therefore yielding a single descendant subhalo.
 If a pair consists of a primary and secondary that never share a \texttt{DescendantID} at a given redshift, the pair is defined as a `non-merger.'
@@
 
-A given pair from the base sample at redshift $z_n$, which by definition passes all selection criteria at $z_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.
+A given pair from the base sample at redshift $z_n$, which by definition passes all selection criteria at $z_n$, can be followed backwards and forwards in time using the \sublink{} merger trees.  
 We track the positions of both the primary and secondary subhalo at each redshift and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
-In cases where the primary or secondary does not have a defined position at a given redshift, we do not compute a value for the separation at that redshift.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs.
-The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be evaluated before and after the skip occurs.
+In cases where the primary or secondary does not have a defined position at a given redshift, we do not compute a value for the separation at that redshift.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the \subfind{} algorithm, it will not have a defined position in the \sublink{} catalogs.
+The \sublink{} algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be evaluated before and after the skip occurs.
 See Sec.~3 in ~\citet{RG2015} for more details.}
@@
 In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift.
-We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
+We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs.
+Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
 The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
@@
 
-We collect the following data for each pair at each redshift where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group \texttt{Group\_R\_TopHat200}, the physical separation (in $\kpc$), and the scaled separation (dimensionless).
+% We collect the following data for each pair at each redshift where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group , the physical separation (in $\kpc$), and the scaled separation (dimensionless).
 
@@
 To distinguish the collection of unique orbits, each pair is assigned a `pairkey' while constructing their orbit. 
-The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the Sublink catalogs, and is unique for each pair of halos. 
+The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the \sublink{} catalogs, and is unique for each pair of halos. 
 After each pair is assigned a unique pairkey, we only keep one instance of an orbit per pair to avoid double/multi-counting in our orbit catalog.
@@
     The sharp decline of the merger fraction to 0 at low redshift is a non-physical feature of the simulation ending at $z=0$.
-    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
+    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.\kc{remove line at 0.9}}
     \label{fig:fmerge}
@@
 \subsection{Number of Pairs}
-We calculate the number of pairs (i.e. orbits in progress) by selecting all pairs at a given redshift that are post-infall and have not merged.
+We calculate the number of pairs (i.e. or equivalently the pairs' orbits) by selecting all pairs at a given redshift that have not merged.
 % The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
 %For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
-A single pair (or equivalently, the pair's orbit) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of pairs at all redshifts from $z=0-3$.
+A single pair may meet the conditions of being a "post-infall and pre-merger" pair for $z=0-3$, and will thus contribute to the number of pairs at all redshifts from $z=0-3$.
 
@@
 The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
-The number of pairs that merge decreases to 0 at $z=0$ since many pairs that exist at low redshift will have merger timescales that span beyond the end of the simulation (i.e., $z=0$).
+The number of pairs that merge decreases to zero at $z=0$ since many pairs that exist at low redshift will have merger timescales that span beyond the end of the simulation (i.e., $z=0$).
 
-When a pair experiences first infall and thus begins to contribute to the number of pairs at later redshifts, we refer to such as pairs being ``added to the sample".
-Likewise, mergers at a given redshift will decrease the number of pairs counted at that redshift, since the pair is then post-merger and thus not included in the count.
-
+As more pairs experience first infall, the number of pairs increases with decreasing redshift. 
+On the other hand, as mergers occur, the number of pairs decreases with decreasing redshift, as pairs that have merged are not longer members of the pairs sample.
 % The difference between high and low mass
@@
 The number of high-mass pairs continues increasing until $z=1$, at which point it remains approximately constant from $z=0-1$ at $\sim1000$ pairs. 
-There is a slight decrease in the number of high mass pairs at the very lowest redshifts $z<0.1$, where the mergers begin to outnumber the added pairs at each redshift.
+There is a slight decrease in the number of high mass pairs at the very lowest redshifts $z<0.1$, where the mergers begin to outnumber the new pairs being added at each redshift.
+
+In Fig.~1 of \chambe{}, the number of low-mass pairs peaks (with $\sim3,000$ pairs) and begins to decrease at $z\sim2$, and the number of high-mass pairs is approximately constant between $z=0-1$, peaking at $\sim700$ pairs.
+We find the same behavior with our pair sample, with low-mass pairs peaking at z=2 and high mass pairs leveling off between $z=0-1$. 
+However, in this study, the number of pairs at a given redshift is higher than in our previous work, since the orbit catalog includes a unique orbit for every pair from the previous work.
+For example, a pair that passes the \chambe{} selection criteria only at $z=1$ will only be counted as a pair at $z=1$ in the previous work, while in this study may be counted at many more redshifts since we can follow the orbit forwards and backwards in time.
 
@@
 At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
-The merger fraction for both mass ranges declines to $0$ at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
+The merger fraction for both mass ranges declines to zero at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
 
@@
 Between $z=0.5-2.5$, the high-mass merger fraction is higher than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift.
-This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$.
-When the number of pairs stops increasing and beings to decrease, more pairs that will not have time to merge before $z=0$ are being added, and many merging pairs are being removed from the sample, thus decreasing the merger fraction.
+This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$, while the number of high-mass pairs continues increasing to $z\sim1$.
+At these redshifts, the number of merging pairs begins declining rapidly. 
+% When the number of pairs stops increasing and beings to decrease, pairs that will not have time to merge before $z=0$ are being added, while many merging pairs are being removed from the sample, thus decreasing the merger fraction.
 Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.
@@
     Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
+    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the selected redshift and the merger redshift.\footnote{Since our orbits are only defined for a pair post-infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
+
     In Sec.~\ref{sec:results-timevsep}, we explore how the merger timescale changes as a function of the separation of the pair across redshifts, and how the merger timescales differ between the low-mass and high-mass pairs.
@@
     We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
-Pairs with small separations may be undergoing a pericentric passage, or very close to merging if on a radial orbit.
+    % Pairs with small separations may be undergoing a pericentric passage, or very close to merging if on a radial orbit.
     Binning the pairs by separation, we can study how the merger timescale changes for pairs at different points in their orbits.
@@
     % low mass example
-We bin our pair sample by merger timescale, in bins of $0.5\Gyr$, and separation in bins of width $10\,\kpc$.
-Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$. 
-The horizontal line in the first panel shows the time remaining in the simulation until $z=0$, above which no merging pairs exist.
+    We bin our pair sample by merger timescale, in bins of $0.5\Gyr$, and separation bins of $10\,\kpc$.
+    Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of merger timescale vs. separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$. 
+    The horizontal line in the first panel shows the time remaining in the simulation until $z=0$, above which no merging pairs exist. \kc{add to fig.}
 
@@
     In addition, the spread of times until merger and separations is smaller at higher redshift. 
-The spread of merger timescales at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$, and likewise for the separations spanned by pairs at both redshifts. 
-This is due to the growth of halo size over time, since more pairs will be post-infall at lower redshifts where the virial radius of the FoF groups is larger.
+    The spread of merger timescales at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
+    Likewise, the spread of separations at $z=1$ is $0-200\kpc$, but at $z=4$ the spread is smaller, between $10-125\kpc$. 
+    This is due to the growth of halos over time, such that more pairs are identified at low redshifts where the virial radius of FoF groups is larger.
 
-Likewise, Fig.~\ref{fig:timevssep-high} presents the distribution for high-mass subhalos.
+    Likewise, Fig.~\ref{fig:timevssep-high} presents the distribution for high-mass pairs.
     The trends between merger timescale and separation are roughly the same as low-mass pairs, with a positive correlation between the two parameters that decreases in slope from low redshift to high redshift.
@@
     A compilation of more specific data is available in Table~\ref{tab:timevssepvsz}.
-Here, we list the fraction of post-infall and pre-merger pairs within three separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction of pairs within each of those separation bins that merge in four different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$.
+    Here, we list the fraction of pairs within three separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction of pairs within each of those separation bins that merge in four different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$.
     We present these data for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$.
@@
     \end{figure*}
+
 \subsection{Redshift Dependence of Merger Timescales}\label{sec:results-timevredshift}
@@
     In this subsection, we study the merger timescale of low-mass and high-mass pairs as a function of redshift.
+
     First, we consider the pairs sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that are post-infall and pre-merger for $z=0-6$. 
@@
 
-Throughout, medians at each redshift are shown by solid pink and green lines, while the 1st and 3rd quartile spread on the median is shown by the shaded regions.
+    Each merging orbit is defined to start at first infall, and
 
     \subsubsection{Full sample}
-    We calculate the time until merger for all post-infall and pre-merger pairs at each redshift in the simulation from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift.
-    We include the full catalog of orbits at each redshift to calculate the merger timescale. 
+        We calculate the merger timescales for pairs at each redshift in the simulation from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
+        We include the full catalog of merging orbits at each redshift to calculate the merger timescale. 
         % why do we want to consider the entire population of halos? why interesting?
@@
         % low vs high
-    Low-mass merger timescales (green) and high-mass merger timescales (pink) are roughly equivalent at all redshifts. 
-    Since we have used the same selection procedure for both high-mass and low-mass pairs, assuming that the pair has experienced first infall, which implies that mergers don't proceed in fundamentally different ways at different mass scales.
+        Low-mass merger timescales (green) and high-mass merger timescales (pink) are roughly equivalent at all redshifts, which implies that mergers do not proceed in fundamentally different ways at different mass scales.
         We explore this point further in Sections~\ref{sec:results-phys} and~\ref{sec:results-scal}.
@@
         The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
-    The abrupt decrease to 0 from $z=0.6$ to $z=0$ is artificial rather than a true physical feature.
-    The majority of low redshift pairs will have merger timescales that span past the end of the simulation, since the time until the end of the simulation gets increasing small for smaller redshift. 
-    However, as $z\to0$, there is a also an increasingly small fraction of total pairs that still merge before $z=0$, as seen in Fig~\ref{fig:numorbits}. 
+        The abrupt decrease to 0 from $z\sim0.6$ to $z=0$ is artificial rather than a true physical feature. 
+        As $z\to0$, there is an increasingly small fraction of pairs that merge before $z=0$, as seen by the dashed lines in Fig~\ref{fig:numorbits}. 
         These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$. 
@@
     \label{sec:results-phys}
-    Both separation selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \texttt{SUBFIND} catalogs.
+        Both separation selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \subfind{} catalogs.
         This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates (see Sec.~\ref{sec:discussion} for further detail).
 
-    The first set of separation criteria will select only pairs at a given redshift that have separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, to create six pair subsamples.
+        The first set of separation criteria will select only pairs at a given redshift that have separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, dividing the sample into six pair subsamples.
         These separation criteria do not vary as a function of time or mass, and are applied equivalently to the low-mass and high-mass samples at all redshifts. 
@@
         These traits were also present in Fig.~\ref{fig:timescales}, so they are features that are roughly independent of a separation criteria.
-    As the selection criteria gets larger, the subsample contains more of the full sample, eventually including the full sample with a large enough separation selection.
+        As the selection criteria gets larger, the subsample contains more of the full sample, eventually including the full sample at large enough separations.
 
@@
         The difference in the median time until merger is up to $0.8\,\Gyr$ longer for low-mass pairs than high-mass pairs at the same redshift for the same separation criteria. 
-    The offset between the low-mass and high-mass merger timescale decreases for large enough physical separation criteria. 
+        The offset between the low-mass and high-mass merger timescale decreases for each successive separation cut, which will each include pairs with larger separation. 
         In the rightmost top panel, the median merger times overlap almost identically for $z=0-0.5$ and $z>1.5$, almost mimicking our findings from the full sample analysis.
@@
         % impact of the separation criteria
-    When the maximum separation is increased (from left to right), more of the full sample is being used in the calculation for the mean merger timescale. 
+        When the maximum separation is increased (from left to right), more of the full sample is used in the calculation for the median merger timescale. 
         Indeed, the merger timescale for both low-mass and high-mass pairs increases with an increasing maximum separation cut.  
@@
         The median merger timescale for low-mass pairs does not change significantly for any separation criteria $>150\,\kpc$. 
-    On the otherhand, high-mass pairs tend to see an increase in the merger timescale for each larger separation cut. 
+        On the other hand, high-mass pairs tend to see an increase in the merger timescale for each larger separation cut. 
         This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
@@
         We calculate the merger timescales for six additional subsamples of pairs, this time applying separation criteria that scale with both redshift and mass.
-    The scaled separation, which we define in~\ref{sec:methods},  is the separation of a pair divided by the virial radius of the pair's FoF group, $\Rvir$. 
+        The scaled separation, which we define in Section~\ref{sec:methods},  is the separation of a pair divided by the virial radius of the pair's FoF group, $\Rvir$. 
         The scaled separation selected pairs have separations greater than $10\,\kpc$ and less than $<[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$. 
@@
         % \item Implications for 
-    \item Sec 1: Comparison to previous works
-
-    Subsec 1: merger fraction studies
+        \item Sec 1: Comparison to previous works: merger fraction/timescale studies
         %Why did we pick the separation cuts that we did? 
         \begin{itemize} 
-        \item What other studies have looked at merger fractions in hydro simulations, specifically? 
+            \item What other studies have looked at merger fractions/timescales in hydro simulations, specifically? 
             \item Have they studied merger fractions as a function of mass and redshift? What did they find?
+            \item Were low-mass pairs expected to have such high merger rates in isolated environments? 
+            \item Have they studied merger timescales as a function of mass and redshift? 
             \item What is unique about our study?
-        \item Were low-mass pairs expected to have such high merger rates in isolated environments? 
             \item Are comparisons with previous works fair, or are they not really comparable?
-        \item specific papers: Ventou 2017
+            \item specific papers: Ventou 2017, Hopkins 2007
         \end{itemize}
 
-    Subsec 2: existing merger timescale studies
-    \begin{itemize} 
-        \item What other studies have looked at merger timescales in hydro simulations? 
-        \item Have they studied merger timescales as a function of mass and redshift? 
-        \item What is unique about our study?
-        \item Do our findings agree? 
-        \item specific papers: Hopkins 2007,
-    \end{itemize}
         \item Sec. 2: Implications for observational merger fractions and merger rates

### 2024-03-25 paper/notes_for_intro.tex
added some notes for intro based on readings

General notes:

RG 2019 looked at the simulated morphology of illustristng galaxies and found that for M*<1e9.5, the morphologies are "less reliable" -> motivates using pair fractions for merger rate studies

Further work that would be nice to do would be to see if all the timescale stuff stays the same for minor mergers, and for smaller bins from 1-1:4

O'Leary 2021 merging systems can be missed if you don't use wide enough sep criteria, thus leading to under predict merger rate
Jiang 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq. 4\&5). Also quantified the circularity of orbits as a function of mass ratio.
Jiang 2014

Paper:
Author:
What's the connection:
Other notes:

PRIORITY: 
    Author: O'Leary 2021
    Paper: Emerge: Constraining merging probabilities and timescales of close galaxy pairs 
    What's the connection:
    Other notes: 
    Empirical models (EMERGE) to probe the dependencies of pair merging probabilities and merging timescales. Also show that some commonly used pair selection criteria may not represent a suitable sample of galaxies to reproduce underlying merger rates. Seem to talk about both observation timescales and merger timescales, and I'm not sure which one they focus on.
    Quantified probability of a "close pair" merging before z=0 and dependence on rsep-proj, LOS vel, and z. 
    On what timescale will an observed pair merge, and for how long is that pair observable given the selection criteria. 
    What determines pair observation timescales (might be a followup from their 2020 paper) 
    Notes that papers that attempt to come up with fitting functions for timescales etc. will all be wrong for pairs w different selection criteria.
    -> Maybe motivates that we should say that future extensions of our work may lead to the development of fitting functions that can be applied to any system? (is this true?)

    Author: Ventou2017
    Paper: 
    What's the connection: 
    Other notes:

    Author: Patton2024
    Paper: 
    What's the connection:
    Other notes:

    Author: Guzman-Ortega 2022
    Paper: Morphological signatures of mergers in the TNG50 simulation and the Kilo-Degree Survey: the merger fraction from dwarfs to Milky Way-like galaxies
    What's the connection:
    Other notes:

OTHER:
    Author: Wang2020 (+Pearson, Rodriguez-Gomez)
    Paper: Towards a consistent framework of comparing galaxy mergers in observations and simulations
    What's the connection: Looked at the major merger fraction of stuff at low-z. Their stellar mass criteria did not cover our range of interest for low-mass systems
    Other notes: Mass dependence of the galaxy major merger fraction at low z. Make synthetic training images from TNG and apply to KiDS images from GAMA. Find that major merger fraction decreases slightly with increasing stellar mass (1e9.5-1e11.5) -> 
    major mergers defined as happening in the next 1Gyr or in the previous 500Myr.

    Paper: 
    Author: RG2017
    What's the connection: 
    Other notes:

    Paper: 
    Author: Patton2020
    What's the connection:
    Other notes:

    Paper: 
    Author: Patton2000
    What's the connection:
    Other notes:

    Paper: Emerge: Empirical predictions of galaxy merger rates since $z \sim 6$ 
    Author: O'Leary 2020
    What's the connection:
    Other notes:

    Paper: SATELLITES IN MILKY-WAY-LIKE HOSTS: ENVIRONMENT DEPENDENCE AND CLOSE PAIRS
    Author: González2013
    What's the connection:
    Other notes:

    Paper: Observational Constraints on the Merger History of Galaxies since $z\sim6$: Probabilistic Galaxy Pair Counts in the CANDELS Fields
    Author: Duncan 2019
    What's the connection:
    Other notes:

    Paper: THE MILLENNIUM GALAXY CATALOGUE: THE CONNECTION BETWEEN CLOSE PAIRS AND ASYMMETRY; IMPLICATIONS FOR THE GALAXY MERGER RATE
    Author: De Propris 2007
    What's the connection:
    Other notes:

    Paper: Galaxy interactions in IllustrisTNG-100, I: The power and limitations of visual identification
    Author: Blumenthal 2020
    What's the connection:
    Other notes:

    Paper: Small-Scale Challenges to the ΛCDM Paradigm
    Author: Bullock 2017
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes:

    Paper:
    Author:
    What's the connection:
    Other notes:

### 2024-03-26 paper/main.tex
Updates from Overleaf

@@
     % Our solution
-    Here, we present a physically-motivated framework by which merger timescales for high-mass and low-mass pairs can be compared equivalently 
+    Here, we present a physically-motivated framework by which merger timescales for high-mass and low-mass pairs can be compared equivalently, 
     %and converted from one system to another 
-    by scaling by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$). 
+    %by scaling 
+    % GB added below
+    wherein separation criteria are scaled
+    by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$). 
     We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass major pairs as a function of cosmic time, and compare to the merger timescales of similarly-selected high-mass major pairs.
@@
 \section{Introduction} \label{sec:intro}
+Outline draft - 
+1 - Galaxy mergers are an importance feature of LCDM theory and are critical for the formation and evolution of galaxies across time. 
+    - An important feature of galaxy mergers is the merger timescale...
 
-% Introduce RG15, Lotz, etc.
-% \input{paper/notes_for_intro}
+2 - Merger timescales are important to calibrate 
+    - They are crucial for understanding the observations of galaxy morphologies and understanding heirarchical evolution, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
+    - Also important for understanding the few well modelled orbits that we do have, particularly in the LG (might come way later)
+
+3 - Merger timescales are also used to get merger rates, both observationally and theoretically, of close pairs.
+    - How are close pairs defined? 
+    - How do merger timescales enter merger rate calculations?
+    - What is important about the merger rates that are used? 
+    - Maybe need to make distinction here that there are also observability timescales, but they are going to be basically useless for TNG100 studies (see below), but also are low mass pairs have a lot of gas and look super weird.
+    - RG 2019 looked at the simulated morphology of illustristng galaxies and found that for $M*<1e9.5$, the morphologies are "less reliable" $\to$ motivates using pair fractions for merger rate studies
+
+4 - Comparing the merger rates of high mass and low mass pairs is important why?
+    - Is important for our understanding of the build up of higher mass things in LCDM, and thus a test of LCDM
+    - Different merger mechanics would be really interesting, i.e. if the DM halo profile (cuspy/core etc) impacted the merger processes
+
+5 - Merger timescales have been studied a lot (mostly in obs and theory of high mass pairs)
+    - Lotz 2011
+    - RG2015
+    - Snyder (maybe?) 
+    - Jiang 2008 \& BK 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq.
+4\&5).
+Also quantified the circularity of orbits as a function of mass ratio $\leftarrow$ might be good to mention
+    - Jiang 2014
+
+6 - Problem is, we don't have a lot of studies on low mass pair merger rates? 
+    - This is not ideal because it means that we typically extend merger rates calibrated for high mass down to low mass, but this means that we aren't able to test LCDM and draw conclusions about the impact
+
+7 - In addition, our previous work showed that even the evolution of the fractions of high and low mass pairs can't be robustly studied and compared if even the selection criteria for the pairs are off.
+    - This means that the selection of pairs for pair fractions, merger fractions, and for merger timescales, are ALL going to be impacted by the selection criteria that's used. 
+    - This is also a problem in observations already, since different samples will have different completeness down to different projected separations, and with different mass ratios, etc. 
+    - O'Leary 2021 merging systems can be missed if you don't use wide enough sep criteria, thus leading to under predict merger rate
+
+8 - Fortunately, our previous work showed that using scaled criteria for pair selection as a function of mass and redshift allows you to fairly compare samples. 
+    - This motivated us to study the merger fraction and merger timescales of our galaxy pair sample from the first study, to see how merger timescales might ALSO depend on the selection criteria that are used to pick the pair sample.
+
+9 - Here, we aim to build a framework to understand the pair selection criteria that are used to select samples, and the merger timescales that result from various selections.
+we look at the orbits of the pairs from our previous work, and determine the merger timescale for low and high mass pairs as a function of pair separation and redshift.
+In sec.
+1...
+etc.
 
@@
 %recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
-We utilize the group catalogs, produced by the \subfind{} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \sublink{} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
+We utilize the group catalogs, produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009}, and the merger tree catalogs, generated by the \sublink{} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
 The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).
 
-Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated as in \citet{Chamberlain2024}.
+Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated, but physically associated, as in \citet{Chamberlain2024}.
 From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.
@@
 
-At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive halos (by stellar mass) from FoF groups with virial mass\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass.
-This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Brynorman1998} and $\rho_c$ is the critical density of the universe at the time calculated.
+At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive subhalos (by assigned stellar mass, using abundance matching; see below) from FoF groups with virial mass
+% GB added M_G
+($\rm M_{G}$)\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass.
+This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Bryan1998} and $\rho_c$ is the critical density of the universe at the time calculated.
 The corresponding virial radius in TNG100 is given by \texttt{Group\_R\_TopHat200}.}
@@
 
-We require that subhalos constituting a pair meet a minimum subhalo mass criteria of 
+We require that subhalos constituting a pair meet a minimum subhalo mass 
+% GB added Mhalo
+($\Mhalo$)
+criteria of 
 \begin{equation*}
@@
 $\sim$100 particles, enough to robustly identify gravitationally bound subhalos in the \subfind{} and \sublink{} catalogs.
-For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \sublink{} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.
+For each subhalo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \sublink{} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.
 
-Stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
+% GB added 
+As in \chambe{}, 
+stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
 The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo via the abundance matching prescription  $\ms{}=f(\Mpeak,z)$, where $z$ is the redshift of consideration.
 
-In \chambe{}, the abundance matching prescription was sampled 1000 times for each subhalo to account for the spread in the Stellar Mass -- Halo Mass (SMHM) relation. 
+In \chambe{}, 
+abundance matching prescription was sampled 1000 times for each subhalo to account for the spread in the Stellar Mass -- Halo Mass (SMHM) relation. 
 For the present study, we only use the stellar masses given by the median of the abundance matching relationship. 
@@
 
-Primary halos are defined as the subhalo with the highest stellar mass ($M_{*1}$) in their FoF group, and secondaries are defined as the second most massive subhalo with stellar mass $M_{*2}$. 
-Our sample of major pairs then consists of all pairs of primary and secondary halos with 
+Primary subhalos are defined as the subhalo with the highest assigned stellar mass ($M_{*1}$) in their FoF group, and secondaries are defined as the second most massive subhalo with stellar mass $M_{*2}$.
+Our sample of major pairs then consists of all pairs of primary and secondary subhalos with 
 \begin{align*} 
@@
 \end{align*}
-A primary or secondary halo can only be a member of one single pair at a given snapshot, such that a collection of $N$ pairs consists of $N$ unique primaries and $N$ unique secondaries.
+A primary or secondary subhalo can only be a member of one single pair at a given snapshot, such that a collection of $N$ pairs consists of $N$ unique primaries and $N$ unique secondaries.
 
@@
 
-The base sample of pairs used for this analysis then consists of the set of all low-mass and high-mass major pairs from each redshift of the TNG100 simulation.
+The base sample of pairs used for this analysis then consists of the set of all
+% GB added
+isolated 
+low-mass and high-mass major pairs from each redshift of the TNG100 simulation.
 
 \subsection{Mergers} \label{subsec:mergers}
-Here we describe how subhalo mergers are identified in the TNG simulations. 
-Note that this definition is specific to subhalo mergers and therefore results are not guaranteed to hold for galaxy mergers, especially in cases where subhalos and galaxies have different centers of mass.
+Here we describe how subhalo mergers are identified. 
+Note that this definition is specific to subhalo mergers and therefore results are not guaranteed to hold for galaxy mergers, especially in cases where subhalos and galaxies have different centers of mass~\citep[see e.g.,]{RG2015}.
 % might  note that this is discussed in more deets in discussion - I think Patton 2024 addresses this a bit?
@@
 This means that the primary and secondary subhalo have merged such that \subfind{} can no longer distinguish them as two separate halos, therefore yielding a single descendant subhalo.
-If a pair consists of a primary and secondary that never share a \texttt{DescendantID} at a given redshift, the pair is defined as a `non-merger.'
+If a primary and secondary never share the same \texttt{DescendantID} at the same redshift, the pair is defined as a `non-merger.'
 
@@
     \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
-    \caption{Selection of example orbits of low-mass major pairs that pass the pair selection criteria at $z=1.5$, showing the separation between the primary and secondary halo of the pair as a function of lookback time.
+    \caption{Selection of example orbits of low-mass major pairs that pass the pair selection criteria at $z=1.5$, showing the separation between the primary and secondary as a function of lookback time.
     (Top) Orbits of pairs that merge before $z=0$.
@@
     (Bottom) Orbits of pairs that do not merge before $z=0$ (non-mergers) and are unlikely to do so in the future $\sim2\Gyr$ (past the end of the simulation).
-    Solid lines represent the post-infall parts of the orbit, while the dashed lines show the pre-infall portion of the orbit before the pairs first share a common FoF group. 
+    Solid lines represent the post-infall orbits, i.e. after the pair share a common FoF group.
+    Dashed lines show the pre-infall portion of the orbit. 
     Triangle points in the first panel show the first redshift after merger where the orbit has a separation of $0\,\kpc$.
@@
 
-A given pair from the base sample at redshift $z_n$, which by definition passes all selection criteria at $z_n$, can be followed backwards and forwards in time using the \sublink{} merger trees.  
+A given pair from the base sample at redshift $z_n$ passes all of the selection criteria from \chambe{} at $z_n$, and can additionally be followed backwards and forwards in time using the \sublink{} merger trees.
 We track the positions of both the primary and secondary subhalo at each redshift and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
@@
 In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift.
-We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs.
-Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
+We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group (also in $\kpc$).\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs.
+If the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
 The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
@@
 
-We define the redshift of ``first infall" as the redshift of the first snapshot where the primary and secondary have the same parent FoF halo, and post-infall is all following redshifts starting with the redshift of first infall.
+We define the redshift of ``first infall" as the redshift of the first snapshot where the primary and secondary have the same parent FoF halo, and post-infall as all following redshifts starting with the redshift of first infall.
 
@@
 The top panel shows the orbits of five pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo.
-Even amongst galaxies of the same approximate mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\,\Gyr$ from first infall to merger,  indicated by the triangles.
+Even amongst galaxies of the same approximate stellar mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\,\Gyr$ from first infall to merger (triangles).
 
-For some pairs, the secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo than their primary.
+For some pairs, the secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo than that of their primary.
 Because all of our mergers occur when the subhalos are in the same FoF group by definition, in these cases, we use the full orbit starting when the subhalos are first identified as members of the same FoF group through when they merge, as described in Section \ref{subsec:mergers}.
@@
 \label{sec:methods-unique}
-Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists.
+Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts after infall, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists.
 However, we only want to keep a single instance of any given orbit in our catalog so we can track the merger timescale of the pair.
@@
 The total number of pairs (including double counting the same pair at multiple redshifts) is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
-However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits that each correspond to a unique pair of galaxies.
+However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits that each correspond to a unique pair of subhalos.
 
@@
 \section{Pair Sample Properties}
-
 \begin{figure}[htb]
@@
     \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
-    The solid lines show the total number of pairs the are post-infall at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
+    The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
     We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
@@
     The sharp decline of the merger fraction to 0 at low redshift is a non-physical feature of the simulation ending at $z=0$.
-    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.\kc{remove line at 0.9}}
+    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
     \label{fig:fmerge}
@@
 \subsection{Number of Pairs}
-We calculate the number of pairs (i.e. or equivalently the pairs' orbits) by selecting all pairs at a given redshift that have not merged.
+We calculate the number of pairs by selecting all pairs at a given redshift that have not merged.
 % The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
 %For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
-A single pair may meet the conditions of being a "post-infall and pre-merger" pair for $z=0-3$, and will thus contribute to the number of pairs at all redshifts from $z=0-3$.
+A single, non-merging pair with first infall at $z=3$ will thus contribute to the number of pairs at all redshifts from $z=0-3$.
 
@@
 
-As more pairs experience first infall, the number of pairs increases with decreasing redshift. 
-On the other hand, as mergers occur, the number of pairs decreases with decreasing redshift, as pairs that have merged are not longer members of the pairs sample.
+As more subhalos experience first infall into a FoF, the number of pairs increases with decreasing redshift. 
+On the other hand, as mergers occur, the number of pairs decreases with decreasing redshift, as pairs that have merged are no longer members of the pairs sample.
 % The difference between high and low mass
@@
 However, in this study, the number of pairs at a given redshift is higher than in our previous work, since the orbit catalog includes a unique orbit for every pair from the previous work.
-For example, a pair that passes the \chambe{} selection criteria only at $z=1$ will only be counted as a pair at $z=1$ in the previous work, while in this study may be counted at many more redshifts since we can follow the orbit forwards and backwards in time.
+For example, a pair that passes the \chambe{} selection criteria only at $z=1$ will only be counted as a pair at $z=1$ in the previous work, while, in this study, it may be counted at many more redshifts since we can follow the orbit forwards and backwards in time.
 
@@
 
-Note that the merger fraction is a measure of the fraction of isolated pairs that merge before $z=0$.
-It is not, however, a measure of the fraction of \textit{all} low-mass and high-mass pairs that will merge.
+Note that the merger fraction defined here is a measure of the fraction of isolated pairs that merge before $z=0$.
+It is not, however, a measure of the fraction of \textit{all} low-mass and high-mass pairs in all environments that will merge.
 
@@
         The dotted black line shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
-        The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
+        The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.\kc{only one of the dashed lines will remain in this plot once I discuss w Paul}}
         \label{fig:timescales}
@@
 
-        In addition, we show 
+    % this is all really discussion, so I can either move this to discussion and reference that we talk about the H(z) line more explicitly in the discussion, or not
+        In principle, merger timescales\footnote{Specifically, the elapsed time between crossing into the virial radius of a subhalo and coalescence with the central galaxy.} for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}.
+        Departures from the idealized case introduce perturbations to that solution, the validity of which has been tested in cosmological hydrodynamic and N-body simulations~\citep{Jiang2008, Boylan_kolchin2008}.
+
+        Such studies have found that the merger timescale in N-body hydrodynamic simulations is of the form
         \begin{equation}
-        \frac{2}{5}*H(z)^{-1}
+            \rm \tau_{\rm merge} = \frac{A(\Theta)}{\ln \Lambda}\frac{\mprim}{\msec}\tau_{\rm dyn}
         \end{equation}
-        given the Hubble parameter $H(z)$ as the dashed black line that sits above the median time until merger.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation.
-Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
-        We also show the age of the simulation as a function of redshift (dashed grey). 
-        These lines show XXX...
-        % so far missing the WHY we are comparing to the hubble time or age vs z
-        % look at Hopkins dynamical friction/merger timescale plot.
+        where $\mprim$ and $\msec$ are the primary and secondary subhalo masses, $A(\Theta)$ is a constant for a given orbital configuration, $\ln\Lambda$ is the Coulomb logarithm taken to be $\ln\Lambda = \ln(1+\mprim/\msec)$, and $\tau_{\rm dyn}$ is the dynamical timescale at the virial radius of the primary subhalo.
+        The dynamical timescale is related to the crossing time at the virial radius, and is given by 
+        \begin{equation}
+            \rm \tau_{\rm dyn} = \frac{\Rvir}{V_{\rm circ}(\Rvir)},
+        \end{equation}
+        where $V_{\rm circ}(\Rvir)$ is the circular velocity at the virial radius of the primary subhalo. 
+        Note that the dynamical time can be rewritten as 
+        \begin{equation}
+            \rm \tau_{\rm dyn} = (G\rho_{crit})^{-1/2} = \bigg(\frac{3\,H^2(z)}{8\pi}\bigg)^{-1/2}.
+        \end{equation}
+
+        In our study, we keep the stellar and FoF group mass criteria fixed as a function of redshift, such that for a given pair, the merger timescale scales with redshift as
+        \begin{equation}
+            \rm \tau_{merge} \propto \tau_{dyn} \propto H(z)^{-1},
+        \end{equation}
+        assuming that there is no (or weak) redshift dependence of the distribution of orbital parameters.
+        Thus, the merger timescale is expected to scale with the Hubble Time at a given redshift.
+
+        In Fig.~\ref{fig:timescales}, the black dashed line represents a merger timescale that scales with $1/H(z)$,
+        \footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation. Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
+        and which is consistent with our findings for redshifts between $z=1.5-6$.
 
@@
         This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
-        High-mass pairs tend to have higher separations than low-mass pairs, as show in~\ref{sec:results-timevsep}.
+        High-mass pairs tend to have higher separations than low-mass pairs, as shown in~\ref{sec:results-timevsep}.
 
@@
             \item Are comparisons with previous works fair, or are they not really comparable?
-            \item specific papers: Ventou 2017, Hopkins 2007
+            \item specific papers: Ventou 2017, Hopkins 2007, O'Leary 20
         \end{itemize}
@@
 
+    Further work that would be nice to do would be to see if all the timescale stuff stays the same for minor mergers, and for smaller bins in range 1-1:4
+
 % ** 

### 2024-03-26 paper/notes_for_intro.tex

@@
 
-RG 2019 looked at the simulated morphology of illustristng galaxies and found that for M*<1e9.5, the morphologies are "less reliable" -> motivates using pair fractions for merger rate studies
+Outline - 
+1 - Galaxy mergers are an importance feature of LCDM theory and are critical for the formation and evolution of galaxies across time. 
+    - An important feature of galaxy mergers is the merger timescale...
 
-Further work that would be nice to do would be to see if all the timescale stuff stays the same for minor mergers, and for smaller bins from 1-1:4
+2 - Merger timescales are important to calibrate 
+    - They are crucial for understanding the observations of galaxy morphologies and understanding heirarchical evolution, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
+    - Also important for understanding the few well modelled orbits that we do have, particularly in the LG (might come way later)
 
-O'Leary 2021 merging systems can be missed if you don't use wide enough sep criteria, thus leading to under predict merger rate
-Jiang 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq.
+3 - Merger timescales are also used to get merger rates, both observationally and theoretically, of close pairs.
+    - How are close pairs defined? 
+    - How do merger timescales enter merger rate calculations?
+    - What is important about the merger rates that are used? 
+    - Maybe need to make distinction here that there are also observability timescales, but they are going to be basically useless for TNG100 studies (RG2019 mocks i think?), but also are low mass pairs have a lot of gas and look super weird.
+    - RG 2019 looked at the simulated morphology of illustristng galaxies and found that for M*<1e9.5, the morphologies are "less reliable" -> motivates using pair fractions for merger rate studies
+
+4 - Comparing the merger rates of high mass and low mass pairs is important why?
+    - Is important for our understanding of the build up of higher mass things in LCDM, and thus a test of LCDM
+    - Different merger mechanics would be really interesting, i.e. if the DM halo profile (cuspy/core etc) impacted the merger processes
+
+5 - Merger timescales have been studied a lot (mostly in obs and theory of high mass pairs)
+    - Lotz 2011
+    - RG2015
+    - Snyder (maybe?) 
+    - Jiang 2008 & BK 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq.
 4\&5).
-Also quantified the circularity of orbits as a function of mass ratio.
-Jiang 2014
+Also quantified the circularity of orbits as a function of mass ratio <- might be good to mention
+    - Jiang 2014
+
+6 - Problem is, we don't have a lot of studies on low mass pair merger rates? 
+    - This is not ideal because it means that we typically extend merger rates calibrated for high mass down to low mass, but this means that we aren't able to test LCDM and draw conclusions about the impact
+
+7 - In addition, our previous work showed that even the evolution of the fractions of high and low mass pairs can't be robustly studied and compared if even the selection criteria for the pairs are off.
+    - This means that the selection of pairs for pair fractions, merger fractions, and for merger timescales, are ALL going to be impacted by the selection criteria that's used. 
+    - This is also a problem in observations already, since different samples will have different completeness down to different projected separations, and with different mass ratios, etc. 
+    - O'Leary 2021 merging systems can be missed if you don't use wide enough sep criteria, thus leading to under predict merger rate
+
+8 - Fortunately, our previous work showed that using scaled criteria for pair selection as a function of mass and redshift allows you to fairly compare samples. 
+    - This motivated us to study the merger fraction and merger timescales of our galaxy pair sample from the first study, to see how merger timescales might ALSO depend on the selection criteria that are used to pick the pair sample.
+
+9 - Here, we aim to build a framework to understand the pair selection criteria that are used to select samples, and the merger timescales that result from various selections.
+we look at the orbits of the pairs from our previous work, and determine the merger timescale for low and high mass pairs as a function of pair separation and redshift.
+In sec.
+1...
+etc.
+
+Conclusions -
 

### 2024-03-26 paper/main.tex
Updates from Overleaf

@@
     We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
-    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.\kc{change ylabel to "pairs"}
+    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
     }
@@
     \begin{center}
-    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
-    \caption{\kc{Add z=5, z=6 panels.
-Add colorbar for percentage per panel (and add number of pairs at that redshift) Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
+    \includegraphics[width=0.8\textwidth]{plots/bet-on-it/8_2Dhist.png}
+    \caption{\kc{(and add number of pairs at that redshift?) Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
     There is a positive correlation between separation at the selected redshift and the time until merger. 

### 2024-03-26 paper/main.tex
combined the 2dhist figures and text

@@
 % ###########################################################################################
-\begin{figure*}[htb]
-    \begin{center}
-    \includegraphics[width=0.8\textwidth]{plots/bet-on-it/8_2Dhist.png}
-    \caption{\kc{(and add number of pairs at that redshift?) Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
-    There is a positive correlation between separation at the selected redshift and the time until merger. 
-    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$. 
-    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
-    The lowest separation pairs have the shortest merger timescales.
-    This plot can be used to get estimates for the merger timescale of an isolated low-mass pair at a given redshift.
-For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
-    \kc{add z=5, z=6, add horizontal line for the time remaining in the sim in panel 1}
-    }
-    \label{fig:timevssep-low}
-    \end{center}
-\end{figure*}
 
@@
     \begin{center}
-    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
-    \caption{\kc{Add z=5, z=6 panels.
-Add colorbar.
-Add low/high label somewhere, check the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
-    Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
-    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$. 
-    This plot can be used to get estimates for the merger timescale of an isolated high-mass pair at a given redshift. 
-    For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
+    \includegraphics[width=0.75\textwidth]{plots/bet-on-it/8_2Dhist.png}
+    \caption{The distribution of times until merger for low-mass (top) and high-mass (bottom) pairs as a function of physical separation at $z=0.5,1,1.5,2,3,4,5\mbox{and }6$. 
+    There is a positive correlation between separation at the selected redshift and the time until merger. 
+    The lowest separation pairs tend to have the shortest merger timescales.
+    Of low-mass pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$, while high-mass pairs have separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$.
+    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
+    This plot can be used to get estimates for the merger timescale of an isolated pair at a given redshift. 
+    For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
+    On the other hand, a high-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-2\,\Gyr$, with a most likely time to merger of around $0-1\Gyr$.
     }
-    \label{fig:timevssep-high}
+    \label{fig:2dhist}
     \end{center}
@@
     Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
-    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the selected redshift and the merger redshift.\footnote{Since our orbits are only defined for a pair post-infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
+    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the redshift at which pairs are selected and the merger redshift.\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
+    Only the merging pairs will be considered for the remainder of the analysis.
 
     In Sec.~\ref{sec:results-timevsep}, we explore how the merger timescale changes as a function of the separation of the pair across redshifts, and how the merger timescales differ between the low-mass and high-mass pairs.
-    In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all post-infall and pre-merger isolated pairs as a function of redshift. 
+    In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all isolated pairs as a function of redshift. 
     We will additionally examine the merger timescale's dependence on a variety of separation criteria. 
@@
     % Pairs with small separations may be undergoing a pericentric passage, or very close to merging if on a radial orbit.
-    Binning the pairs by separation, we can study how the merger timescale changes for pairs at different points in their orbits.
+    Binning the pairs by separation, we can study how the merger timescale changes for pairs selected at different points in their orbits.
 
     % low mass example
-    We bin our pair sample by merger timescale, in bins of $0.5\Gyr$, and separation bins of $10\,\kpc$.
-    Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of merger timescale vs. separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$. 
-    The horizontal line in the first panel shows the time remaining in the simulation until $z=0$, above which no merging pairs exist. \kc{add to fig.}
+    We bin our pair sample by merger timescale, in bins of $0.5\Gyr$, and separation bins of $20\,\kpc$, and only consider the pairs with separations $\rsep>10\kpc$.
+    Figure~\ref{fig:2dhist} shows a heatmap distribution of pairs as a function of merger timescale vs.
+separation for low-mass (top) and high-mass (bottom) pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$.
+    The colorbars to the right of each figure show the percentage of the population of pairs at that redshift that are in each bin, and the number of pairs at each redshift is shown by the number in the bottom right corner of each panel.
+    The horizontal line in the first panel shows the time remaining in the simulation until $z=0$, above which no merging pairs exist.
 
-    Fig.~\ref{fig:timevssep-low} presents a heatmap distribution of the density of low-mass pairs in merger timescale-separation space. 
-    We find that the time until merger for low-mass pairs is positively correlated with the separation of the pair at each redshift. 
-    However, the slope of the correlation decreases at higher redshifts.
-    Pairs at $z=4$ and separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.
+    We find that the time until merger for low-mass and high-mass pairs is positively correlated with the separation of the pair at each redshift. 
+    However, the slope of this correlation decreases at higher redshifts.
+    Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.
 
-    In addition, the spread of times until merger and separations is smaller at higher redshift. 
-    The spread of merger timescales at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
+    In addition, the spread of times until merger and separations is smaller at higher redshift, for both low- and high-mass pairs. 
+    The spread of merger timescales for low-mass pairs at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
     Likewise, the spread of separations at $z=1$ is $0-200\kpc$, but at $z=4$ the spread is smaller, between $10-125\kpc$. 
+    The high-mass pairs also have a larger spread in the distribution of merger timescales and separations at low redshift than at higher redshift.
     This is due to the growth of halos over time, such that more pairs are identified at low redshifts where the virial radius of FoF groups is larger.
-
-    Likewise, Fig.~\ref{fig:timevssep-high} presents the distribution for high-mass pairs.
-    The trends between merger timescale and separation are roughly the same as low-mass pairs, with a positive correlation between the two parameters that decreases in slope from low redshift to high redshift.
-    The high-mass pairs also have a larger spread in the distribution of merger timescales and separations at low redshift than at higher redshift.
-    This is once again caused by the change in the FoF halo size over time.
 
@@
     Since pair samples are typically picked via separation criteria, in both simulations and observations, we study the impact of different separation-based selection criteria on inferred merger timescales.
-
-    Each merging orbit is defined to start at first infall, and
 
@@
 \section{Discussion} \label{sec:discussion}
-    We have calculated the merger timescales, or time until merger, for a sample of low-mass and high-mass pairs that have experienced first infall but not yet merged, to ensure that they are physically collocated.
+    We have calculated the merger timescales, or time until merger, for a sample of low-mass and high-mass pairs that have experienced first infall and merge before $z=0$ to ensure that they are physically collocated.
     In this section, we will explore the implications and broader impacts of our findings, and put our study in context with previous work. 

### 2024-03-28 paper/main.tex
added text to discussion and conclusions

@@
 \section{Introduction} \label{sec:intro}
-Outline draft - 
-1 - Galaxy mergers are an importance feature of LCDM theory and are critical for the formation and evolution of galaxies across time. 
-    - An important feature of galaxy mergers is the merger timescale...
+% Outline draft - 
+% 1 - Galaxy mergers are an importance feature of LCDM theory and are critical for the formation and evolution of galaxies across time. 
+%     - An important feature of galaxy mergers is the merger timescale...
 
-2 - Merger timescales are important to calibrate 
-    - They are crucial for understanding the observations of galaxy morphologies and understanding heirarchical evolution, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
-    - Also important for understanding the few well modelled orbits that we do have, particularly in the LG (might come way later)
+% 2 - Merger timescales are important to calibrate 
+%     - They are crucial for understanding the observations of galaxy morphologies and understanding heirarchical evolution, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
+    %- Also important for understanding the few well modelled orbits that we do have, particularly in the LG (might come way later)
 
-3 - Merger timescales are also used to get merger rates, both observationally and theoretically, of close pairs.
-    - How are close pairs defined? 
-    - How do merger timescales enter merger rate calculations?
-    - What is important about the merger rates that are used? 
-    - Maybe need to make distinction here that there are also observability timescales, but they are going to be basically useless for TNG100 studies (see below), but also are low mass pairs have a lot of gas and look super weird.
-    - RG 2019 looked at the simulated morphology of illustristng galaxies and found that for $M*<1e9.5$, the morphologies are "less reliable" $\to$ motivates using pair fractions for merger rate studies
+% 3 - Merger timescales are also used to get merger rates, both observationally and theoretically, of close pairs.
+%     - How are close pairs defined? 
+%     - How do merger timescales enter merger rate calculations?
+%     - What is important about the merger rates that are used? 
+%     - Maybe need to make distinction here that there are also observability timescales, but they are going to be basically useless for TNG100 studies (see below), but also are low mass pairs have a lot of gas and look super weird.
 
-4 - Comparing the merger rates of high mass and low mass pairs is important why?
-    - Is important for our understanding of the build up of higher mass things in LCDM, and thus a test of LCDM
-    - Different merger mechanics would be really interesting, i.e. if the DM halo profile (cuspy/core etc) impacted the merger processes
+Galaxy merger rates are a test of LCDM theory.
+In the era of JWST, these tests can push to higher redshifts and lower mass scales.
+At high redshift and at lower masses, even isolated galaxies are highly disturbed, making pair fractions critical for merger rate studies \cite{Martin2018, RG2019}.
+However, there is no framework for comparing merger rates from pair fractions for low mass versus high mass galaxies and tensions exist in the literature at low masses (e.g.
+Vicente paper figure).
+i.e.
+There is a tension at low mass% vicente's work is done consistently, but we're basically showing that if you want to put the observational stuff on there too, you can't just willy nilly use any old pair criteria
 
-5 - Merger timescales have been studied a lot (mostly in obs and theory of high mass pairs)
-    - Lotz 2011
-    - RG2015
-    - Snyder (maybe?) 
-    - Jiang 2008 \& BK 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq.
+This study seeks to establish a framework to enable comparisons of merger rates across mass scales and redshift so that merger rates of low mass galaxies can be used as tests of LCDM theory.
+These results can also be used to interpret observations of galaxy morphologies and their hierarchical evolution.
+, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
+
+% 4 - Comparing the merger rates of high mass and low mass pairs is important why?
+%     - Is important for our understanding of the build up of higher mass things in LCDM, and thus a test of LCDM
+%     - Different merger mechanics would be really interesting, i.e. if the DM halo profile (cuspy/core etc) impacted the merger processes
+
+% 5 - Merger timescales have been studied a lot (mostly in obs and theory of high mass pairs)
+%     - Lotz 2011
+%     - RG2015
+%     - Snyder (maybe?) 
+%     - Jiang 2008 \& BK 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq.
 4\&5).
 Also quantified the circularity of orbits as a function of mass ratio $\leftarrow$ might be good to mention
-    - Jiang 2014
+%     - Jiang 2014
 
-6 - Problem is, we don't have a lot of studies on low mass pair merger rates? 
-    - This is not ideal because it means that we typically extend merger rates calibrated for high mass down to low mass, but this means that we aren't able to test LCDM and draw conclusions about the impact
+% 6 - Problem is, we don't have a lot of studies on low mass pair merger rates? 
+%     - This is not ideal because it means that we typically extend merger rates calibrated for high mass down to low mass, but this means that we aren't able to test LCDM and draw conclusions about the impact
 
@@
 \end{align*}
-Utilizing the most massive subhalos from the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.
+Requiring pairs to belong to the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.
 
@@
 If a primary and secondary never share the same \texttt{DescendantID} at the same redshift, the pair is defined as a `non-merger.'
+% in this study, we will look at the merger timescales of the merger sample.
 
@@
 
-The collection of all unique orbits constitutes our orbit sample, and is the dataset that will be used for the remainder of the paper.
+The collection of all unique orbits constitutes our orbit sample and is the dataset that will be used for the remainder of the paper.
 
-\section{Pair Sample Properties}
+\section{Pair Sample Properties}\label{sec:pairprops}
 \begin{figure}[htb]
@@
     The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
-    We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
+    We only count the number of orbits ``at" a given redshift if the pairs have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
     The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
@@
 
-\subsection{Number of Pairs}
+\subsection{Number of Pairs}\label{sec:pairprops-num}
 We calculate the number of pairs by selecting all pairs at a given redshift that have not merged.
@@
 
-\subsection{Merger Fraction}
+\subsection{Merger Fraction}\label{sec:pairprops-frac}
 We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift, including all pairs that do and do not merge.
@@
 Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.
+% move the above to discussion and end with something shorter: 
+% This COULD be that the timescales. This will answer this with this study and discuss this in detail in Disc.~\ref.
 
@@
     However, the slope of this correlation decreases at higher redshifts. 
+    % For the same physical separation, the merger timescale is lower than for low-mass 
+    % This does that the overall merger timescales are lower for high-mass systems. 
+    % This is discussed in detail in the next section.
     Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.
@@
 
-    Data from the table can be interpreted in the following way: at $z=1.5$, there are 5,787 pairs that are post-infall and merge prior to $z=0$. 
-    Of those, 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 of those $10-50\,\kpc$ pairs will merge in the next $2\,\Gyr$.
+    Data from the table can be interpreted in the following way: at $z=1.5$, there are 5,787 pairs that merge prior to $z=0$. 
+    Of those, 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 (38.1\%) of those $10-50\,\kpc$ pairs will merge in the next $2\,\Gyr$.
 
@@
         \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale_mod.png}
-        \caption{The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs.
+        \caption{The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs that will merge by $z=0$.
 Shaded regions represent the 1st and 3rd quartile spread on the median.
-Pairs at each redshift are selected if they have experienced first infall and are pre-merger.
+%Pairs at each redshift are selected if they have experienced first infall and are pre-merger.
         % 
-        The black dashed line shows a fraction of the Hubble time as a function of redshift, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
+        The black dashed line shows the fraction of the Hubble time as a function of redshift, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
         The dotted black line shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
-        The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.\kc{only one of the dashed lines will remain in this plot once I discuss w Paul}}
+        The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.\kc{only one of the dashed lines will remain in this plot once I discuss w Paul; likely remove age of sim}}
         \label{fig:timescales}
@@
 
-    First, we consider the pairs sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that are post-infall and pre-merger for $z=0-6$. 
+    First, we consider the pairs sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that will merge for $z=0-6$. 
     We will then consider two different sets of separation criteria to create separation-selected subsamples.
@@
         We calculate the merger timescales for pairs at each redshift in the simulation from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
-        We include the full catalog of merging orbits at each redshift to calculate the merger timescale. 
+        We include the full catalog of merging orbits at each redshift.% to calculate the merger timescale. 
         % why do we want to consider the entire population of halos? why interesting?
@@
         The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
-        The abrupt decrease to 0 from $z\sim0.6$ to $z=0$ is artificial rather than a true physical feature. 
+        The abrupt decrease is artificial rather than a true physical feature. 
         As $z\to0$, there is an increasingly small fraction of pairs that merge before $z=0$, as seen by the dashed lines in Fig~\ref{fig:numorbits}. 
         These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$.
+These mergers must then proceed on decreasing timescales by definition, as they were selected as pairs that must merge prior to $z=0$ (i.e., a pair chosen by the selection criteria given in Section XYZ at z=0.2 must merge by $z=0$, giving a maximal merger timescale equivalent XYZ Gyr whereas a pair chosen at z=6 can have a maximal merger timescale of XYZ Gyr.).
         This is shown by the dotted black line on the left of Fig.~\ref{fig:timescales}, which shows the lookback time of the simulation as a function of redshift.\footnote{The lookback time at a given redshift $z_n$ is equivalent to the time elapsed between $z_n$ and $z=0$.
@@
 
-    % this is all really discussion, so I can either move this to discussion and reference that we talk about the H(z) line more explicitly in the discussion, or not
-        In principle, merger timescales\footnote{Specifically, the elapsed time between crossing into the virial radius of a subhalo and coalescence with the central galaxy.} for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}.
+        In principle, merger timescales\footnote{Specifically, the elapsed time between crossing into the virial radius of the primary subhalo and coalescence with the primary.} for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}.
         Departures from the idealized case introduce perturbations to that solution, the validity of which has been tested in cosmological hydrodynamic and N-body simulations~\citep{Jiang2008, Boylan_kolchin2008}.
@@
         assuming that there is no (or weak) redshift dependence of the distribution of orbital parameters.
-        Thus, the merger timescale is expected to scale with the Hubble Time at a given redshift.
+        Thus, the merger timescale is expected to scale with the Hubble time at a given redshift.
 
@@
 
-        The first set of separation criteria will select only pairs at a given redshift that have separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, dividing the sample into six pair subsamples.
+        The first set of separation criteria will select only pairs at a given redshift that have physical 3D separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, dividing the sample into six pair subsamples.
         These separation criteria do not vary as a function of time or mass, and are applied equivalently to the low-mass and high-mass samples at all redshifts. 
-        We note that an orbit in the separation bin $10-50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts. 
+        For example, an orbit in the separation bin $10-50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts. 
         Rather, the separation criteria are applied independently at each redshift.
@@
         These traits were also present in Fig.~\ref{fig:timescales}, so they are features that are roughly independent of a separation criteria.
-        As the selection criteria gets larger, the subsample contains more of the full sample, eventually including the full sample at large enough separations.
+        As the selection criteria gets larger, the subsample contains more of the full sample, eventually including the full sample of merging pairs at large enough separations.
 
@@
         The offset between the low-mass and high-mass merger timescale decreases for each successive separation cut, which will each include pairs with larger separation. 
-        In the rightmost top panel, the median merger times overlap almost identically for $z=0-0.5$ and $z>1.5$, almost mimicking our findings from the full sample analysis.
+        In the rightmost top panel, the median merger times converge for $z=0-0.5$ and $z>1.5$, as expected based on results from the full sample shown in Fig.
+\ref{fig:timescales}.
+%almost mimicking our findings from the full sample analysis.
 
@@
 \section{Discussion} \label{sec:discussion}
-    We have calculated the merger timescales, or time until merger, for a sample of low-mass and high-mass pairs that have experienced first infall and merge before $z=0$ to ensure that they are physically collocated.
+    We have calculated the merger timescales, or time until merger, for a sample of physically collocated low-mass and high-mass pairs from $z=0-6$.
     In this section, we will explore the implications and broader impacts of our findings, and put our study in context with previous work. 
@@
 
+    In Sec.~\ref{sec:disc-nonmergers}, we will discuss the non-merger portion of our orbit sample and the difference between low- and high-mass merger fractions.
+    In Sec.~\ref{sec:disc-previous}, we discuss our work in the context of other studies that have sought to reconcile xx xmerger rates and merger timescales via a change in pair selection criteria. 
+    In Sec.~\ref{sec:disc-suggestions}, we discuss the implications of our work for future observational and theoretical merger fraction and merger rate studies and provide suggestions for a self-consistent framework moving forward.
+    Finally, in Sec.~\ref{sec:disc-phys}, we briefly discuss the underlying physical mechanisms implied by our results and the functional form of the merger timescale.
+
+    \subsection{Non-merger population}\label{sec:disc-nonmergers}
+    In our catalog of 22,213 low-mass pairs and 3,029 high-mass pairs, only 16.66\% (3700) of low-mass pairs and 29.71\% (903) of high-mass pairs do not merge by $z=0$.
+    In Sec.~\ref{sec:pairprops-frac}, we found that the merger fraction of low-mass and high-mass pairs at $z>3$ is $>0.95$. 
+    This changes as you move to lower redshifts, where the merger fraction decreases substantially. 
+    While the decline in the merger fraction at $z\lesssim1$ is primarily due to the limited time that a pair will have to merge before the end of the simulation, we also find that between $z=\sim0.6-2.5$ the fraction of low-mass pairs is lower than that of high-mass pairs.
+
+    One potential explanation for this behavior could be that low-mass pairs have shorter merger timescales than high mass pairs, and thus merge at earlier times in the simulations, thus leading to a decrease in the merger fraction at higher redshifts.
+    However, through our analysis, we have shown that the merger timescales of our full low-mass and high-mass pair sample evolve in the same way from $z=0-6$. 
+    This means that the difference in the merger fraction between $z=0.5-2.5$ is not a result of different merger timescales.
+
+    We suspect that the cause of the difference between low- and high-mass merger fractions is not a difference in the merging population itself, but rather in the non-merger population and its evolution over time.
+    As we did not distinguish our non-merging sample into "likely mergers" and "fly-by" interactions for the purposes of the study, we cannot say anything about how these two populations individually contribute to the difference in the merger fraction.
+    However, one possible explanation could be that a larger fraction of infalling low-mass pairs at lower redshift result in fly-by interactions. 
+    This could be the case if the velocity at which secondary subhalo enter the primary's FoF group gets increasingly large compared to the virial velocity\footnote{The virial velocity is the value of the circular velocity at the virial radius.}, resulting in fewer bound systems.
+    We leave this as a focus of future work.
+
+    % Specifically, we believe that the separation and velocity of the pairs that experience first infall at lower redshifts ($z<2.5$) lead to but do not merge in the simulation are driving the difference, and that more of the infalling low-mass pairs result in non-mergers than the high-mass pairs.
+    % More specifically, if low-mass pairs at low redshift ($z<3$) have infall velocities that are lower with respect to the virial velocity than their high-mass counterparts, then the merger timescale for that sample of longer-.
+
+    \subsection{Steps Towards Standardized Pair Selection Criteria}~\label{sec:disc-previous}
+    Comparisons between studies of pair fractions, merger fractions, and merger rates across cosmic time and different mass scales are often made challenging by the use of different sets of selection criteria in each study.
+    On the observational side, pair selection criteria are often set by the observational parameters of the survey itself. 
+    For example, limiting magnitudes and completeness limits can enforce a specific new range of stellar mass ratios to be considered (for example $1-1:6$ instead of $1:10$), while avoiding colliding/overlapping fibers on a spectroscopic survey might lead to very specific separation criteria.
+
+    On the theoretical side, it can be straightforward to adopt a number of different pair selection criteria for comparisons to specific observational studies with identical criteria. 
+    However, this means that theoretical studies typically restrict themselves to a subset of pair selection criteria specific to their target comparison studies.
+
+    A variety of studies in recent years aim to standardize pair selection criteria for future studies, in particular to permit a more fair comparison between theory and observations. 
+    One such study is~\citet{Ventou2017}, which created mock catalogs from the original Illustris-1 simulation to analyze the merger probability of galaxy pairs as a function of projected separation and projected relative velocity.
+    They find that 
+    to provide a new set of selection criteria to define close pairs. 
+    One
+
+    Merger rates/merger fractions in illustris, close pair fractions and merger timescales, etc. 
+    As in our previous work on pair fractions in TNG100, we are unable to make meaningful comparisons with previous observational studies of merger fraction and merger rates, as our study is completed using true physical separations from the simulation.
+    However, 
+    % In~\cite{RG2015}, they quantify the galaxy major merger rate as a function of the stellar mass of the descendent subhalo. 
+    % It is unclear, however, if the major merger rate for galaxies with stellar mass $\ms<10^{10}\Msun$ is actually discrepant between simulations and observations, or if the observational sam
+
+    NOT a lit review of merger topics etc. 
         \begin{itemize}
-        % \item Implications for 
-        \item Sec 1: Comparison to previous works: merger fraction/timescale studies
-        %Why did we pick the separation cuts that we did? 
-        \begin{itemize} 
-            \item What other studies have looked at merger fractions/timescales in hydro simulations, specifically? 
-            \item Have they studied merger fractions as a function of mass and redshift? What did they find?
-            \item Were low-mass pairs expected to have such high merger rates in isolated environments? 
             \item Have they studied merger timescales as a function of mass and redshift? 
@@
             \item specific papers: Ventou 2017, Hopkins 2007, O'Leary 20
+            \item only look at merger rates on the low mass end ? - garreth 
+            \item studies that redefine pair fraction criteria 
+            \item What other studies have looked at merger fractions/timescales in hydro simulations, specifically? 
+            \item Have they studied merger fractions as a function of mass and redshift? What did they find?
+            \item Were low-mass pairs expected to have such high merger rates in isolated environments? 
         \end{itemize}
 
-        \item Sec. 2: Implications for observational merger fractions and merger rates
+    \subsection{Implications for Future Observational and Theoretical Studies of Pairs}~\label{sec:disc-suggestions}
+    In Sec.~\ref{sec:results-timevredshift}, we find that the merger timescales of isolated low-mass and high-mass pairs evolve nearly identically between $z=0-6$, rising from a median of $\sim0.8Gyr$ at $z=6$ to $2.1-2.4\Gyr$ at $z\sim0.6$.
+    From $z=0$ to $z=0.6$, the merger timescale rapidly declines, since the only mergers at low redshift that can proceed before the end of the simulation at $z=0$ are those with very short timescales.
+
+    In our previous work, we studied the impact of different sets of separation criteria on the recovered pair fractions of low-mass and high-mass pairs, and found that only by utilizing scaled separation criteria can the features of the pair fractions' evolution with redshift be distinguished~\citep{Chamberlain2024}.
+    In this work, we explored how these same set of selection criteria impacts the recovered merger timescales.
+
+    We find that all subsamples of our pair catalog selected using a mass and redshift evolving separation criterion, namely the scaled separation given by $\rsep/\Rvir$, result in identical behavior of the low-mass and high-mass pair fractions as a function of redshift.
+    On the other hand, selecting pairs using a static physical separation (such as requiring pairs to have separations between $10-50\kpc$, and using that selection criteria for low- and high-mass pairs at every redshift) results in merger timescales that evolve differently between the two mass scales.
+    For example, the median merger timescale for close pairs selected with $\rsep<70\kpc$ varies by up to $\sim0.6\Gyr$ between low- and high-mass pairs at all considered redshifts, while the merger timescale for close pairs selected via the scaled criteria $\rsep<0.5\Rvir$ have nearly identical median timescales.
+
+    Our results demonstrate the importance of utilizing pair selection criteria that permit fair comparisons of pair properties in future observational studies (i.e.
+pair fractions, merger timescales, merger rates, etc.), particularly when the goal of such studies is to quantify and compare the redshift evolution of pair properties at different mass scales.
+    We thus promote the adoption of separation criteria that vary as a function of mass and redshift for such future studies. 
+    In particular, we suggest employing a separation criteria based on the scaled separation of the system.
+
+    Although this seems to introduce further complications to the  pair selection process in observational studies, the use of a scaled separation as demonstrated here can be applied in a fairly straightforward way.
+    In particular, given an observation of a pair with stellar masses $\ms1$ and $\ms2$, we can first compute the associated dark matter mass of the pair using the SMHM relation. 
+    The SMHM relationship of Moster 2013 accounted for observational stellar mass errors  when fitting their relationship, which means that the impact of using the SMHM to convert from stellar mass to halo mass introduces the same error as it does in this study (but we didn't use it in this study technically...
+think about this).
+
+    Once you have the dark matter masses of the two subhalos, you can compute the total dark matter mass, and get an estimate of the overall virial radius of the group by assuming a FoF mass equivalent to the sum of the dark matter masses.
+    Thus, coming up with virial radius criteria per pair shouldn't in principle be that hard, especially since stellar masses are already needed to be calculated to claculate the stellar mass ratio of observed pairs anyways.
+
+    Our study holds for isolated pairs, however we have not extended this analysis to more high-density environments. 
+    Some strengths of this approach are that we can study the pair evolution behvaior as a function of redshift, and the dynamics will be inherent to the pair rather than 
+    Some weaknesses are that there is a chance that the low-mass and high-mass merger timescale of high density environments vary/evolve completely independently of our findings. 
+    We leave this as the focus of future work.
+
+    Bring up assumption of isolation criteria and it's drawbacks and strengths.
+Reiterate that our findings are for isolated pairs.
+-> Focus of future work to explore the impact of pair selection criteria on pairs in a variety of different galaxy environments.
             \begin{itemize}
@@
             \item specific papers: Lotz 2011, 
+            % As discussed in Chamberlain, this is doable, the abundance matching prescription was formulated with the stellar mass errors already taken into account, so this isn't introducing significantly more error in the analysis
+            % quote the stellar mass errors being folded in at high z 
         \end{itemize}
-        \item Sec 3: The functional form of the merger timescale
+
+   \subsection{The Underlying Physical Behavior of Merger Timescales}~\label{sec:disc-phys}
+
         \begin{itemize}
@@
             \item Scaled separation criteria are akin to picking the same fraction of the total volume of a FoF group, regardless of what redshift it is at, not what mass it has, so it may not be surprising that we got what we did in re: the scaled separation being better, but finding the SAME merger timescales for hi and low is interesting!
-            \item Maybe something to consider in this section: For things that don't merge, what's... going on? 
             \item relevant papers: 
         \end{itemize}
-    \end{itemize}
-
-    Bring up assumption of isolation criteria and it's drawbacks and strengths.
-    Reiterate that our findings are for isolated pairs.
 
@@
     % what we did and how
-    $<$Longer summary of what we did here$>$
+    In this study, we construct a sample of the orbits of low-mass ($10^8 <\ms<5\times10^{9}\Msun$) and high-mass ($5\times10^{9}<\ms<10^{11}\Msun$) major pairs (stellar mass ratio $> 1:4$) from $z=0-6$ in the TNG100 simulation.
+    Orbits of pairs, i.e.
+the separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo  first share a common FoF group (or have experienced 'first infall') to either $z=0$ or the redshift at which the pair merges.
 
-    % set of conclusions from results
-    Our main findings are as follows: 
+    The sample consists of 22,213 unique low-mass orbits and 3,029 unique-high mass orbits, for which we quantify the merger fraction as a function of redshift.
+    From this sample, we calculate the merger timescales of low-mass and high-mass major pairs as a function of separation and separately as a function of redshift. 
+    Our goal is to identify the merger timescales of pairs on cosmologically motivated orbits, and to study and possibly identify differences between the redshift evolution of the merger timescales between low-mass and high-mass pairs.
+
+    Additionally, motivated by our previous work on the pair fractions of low-mass and high-mass pairs, where we showed that the evolution of the pair fraction is sensitive to the separation criteria used to select the pairs, we seek to determine the impact of various selection criteria on the merger timescales.
+    Specifically, we look at two separate sets of selection criteria, one which selects pairs via a static physical separation cut, and the other which selects pairs based on a redshift and mass evolving separation cut.
+    This is especially important for studies that seek to study pair fractions, merger fractions, merger timescales, and merger rates of different galaxy mass scales and as a function of redshift.
+
+    Our main conclusions are as follows: 
     \begin{itemize}
-        \item for a given physical separation (such as 100kpc), high-mass pairs have shorter merger timescales than low-mass pairs. 
-        \item the median time until merger for the full sample of low and high-mass pairs increases as $\sim(1+z)^{-1}$ from $z=6$ to $z\sim0.75$.
-it then decreases to 0 at low redshifts, with an upper bound set by the time remaining in the simulation before $z=0$.
+        % merger fractions
+        \item The merger fraction of low-mass and high-mass physically associated pairs is high ($>0.9$) at $z>\sim1.5$ (see Fig.~\ref{fig:fmerge}.
+However, the merger fraction declines rapidly to 0 as $z\to0$, due to the finite length of the simulation which artificially reduces the number of pairs with enough time left to merge at low redshifts.
+% State whether it's a surprise or not that we see such high merger fractions, or if it is consistent w anything else, if you find something ?
+        % merger timescales at a physical sep
+        \item The merger timescale of a pair at a given redshift increases with increasing pair separation.
+Additionally, for a given physical separation, high-mass pairs have a shorter merger timescale than low-mass pairs at the same redshift with the same separation.
+For example, at $z=2$, high-mass pairs at $100kpc$ merge within $\sim 0-3 \Gyr$, while low-mass pairs at $100\kpc$ merge within $\sim 0.5-5\Gyr$ (see Fig.~\ref{fig:2dhist}).
+        % Median merger timescale
+        \item The median merger timescale for the full sample low-mass and high-mass pairs that merge prior to $z=0$ declines with increasing redshift at a rate proportional to $1/H(z)$ for redshifts $z\gtrsim1$ (see Fig.~\ref{fig:timescales}).
+        The maximum merger timescale for a low-redshift pair ($z<1$) is constrained by the time remaining in the simulation for the pair to merge, which causes the median merger timescale to decline to 0 at $z=0$.
         The decrease in the merger time at low redshifts is artificial and not representative of the true merger timescales that pairs at $z=0$ will have.
-        \item separation criteria that include larger separations will increase the median time until merger.
-        \item static physical separation-selected subsamples of low- and high-mass pairs have different median merger timescales.
+        The median merger timescale peaks around $z\sim0.6$ at $\sim2.1-2.4\Gyr$ for both low-mass and high-mass pairs.
+        % The separation criteria effect 
+        \item \kc{left off here}When pairs are selected via a maximum separation cut
+        static physical separation-selected subsamples of low- and high-mass pairs have different median merger timescales.
         \item scaled separation-selected subsamples yield nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts. 

### 2024-03-29 paper/main.tex
complete discussion and conclusion, and half finished intro!!

@@
 \section{Introduction} \label{sec:intro}
-% Outline draft - 
-% 1 - Galaxy mergers are an importance feature of LCDM theory and are critical for the formation and evolution of galaxies across time. 
-%     - An important feature of galaxy mergers is the merger timescale...
+Galaxy merger rates provide an important pathway for tests of hierarchical assembly from$\Lambda$CDM theory, and are critical for understanding the formation and evolution of galaxies across time~\citep[e.g.][]{Stewart2009, Hopkins2010b,RG2015}.
+In the era of future surveys with JWST, Rubin, and Roman, these tests will be extended to higher redshifts and lower mass scales than were previously accessible~\citep{Gardner2006,Spergel2015,Robertson2019a,Robertson2019b,Behroozi2020,Martin2022}.
 
-% 2 - Merger timescales are important to calibrate 
-%     - They are crucial for understanding the observations of galaxy morphologies and understanding heirarchical evolution, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
-    %- Also important for understanding the few well modelled orbits that we do have, particularly in the LG (might come way later)
+At higher redshift ($z>\kc{xx}$) and at lower masses ($\ms{} \lesssim10^{9.5}$), even isolated galaxies are highly disturbed~\citep{Wuyts2013,Martin2018,RG2019,Martin2021}, making close pair fractions a critical alternative to morphological signatures for merger rate studies.
+However, there is currently no framework for comparing merger rates for low-mass and high-mass galaxies as a function of time, and tensions between predictions from theory and observational studies exist in the literature at low masses~\citep[see e.g.
+Fig.~10 in][]{RG2015}.
 
-% 3 - Merger timescales are also used to get merger rates, both observationally and theoretically, of close pairs.
-%     - How are close pairs defined? 
-%     - How do merger timescales enter merger rate calculations?
-%     - What is important about the merger rates that are used? 
-%     - Maybe need to make distinction here that there are also observability timescales, but they are going to be basically useless for TNG100 studies (see below), but also are low mass pairs have a lot of gas and look super weird.
-
-Galaxy merger rates are a test of LCDM theory.
-In the era of JWST, these tests can push to higher redshifts and lower mass scales.
-At high redshift and at lower masses, even isolated galaxies are highly disturbed, making pair fractions critical for merger rate studies \cite{Martin2018, RG2019}.
-However, there is no framework for comparing merger rates from pair fractions for low mass versus high mass galaxies and tensions exist in the literature at low masses (e.g.
-Vicente paper figure).
-i.e.
-There is a tension at low mass% vicente's work is done consistently, but we're basically showing that if you want to put the observational stuff on there too, you can't just willy nilly use any old pair criteria
-
-This study seeks to establish a framework to enable comparisons of merger rates across mass scales and redshift so that merger rates of low mass galaxies can be used as tests of LCDM theory.
-These results can also be used to interpret observations of galaxy morphologies and their hierarchical evolution.
-, (think galaxy start bursts, formation of gas bridges maybe, fueling agn, BH merger rates, etc)
+This study seeks to establish a framework to enable comparisons of merger rates across mass scales and redshift so that merger rates of low-mass galaxies can be used as tests of $\Lambda$CDM theory. 
+These results can also be used to interpret observations of galaxy morphologies and their hierarchical evolution, including the role of mergers in triggering starbursts, fueling active galactic nuclei (AGN), facilitating the formation of tidal features, etc., in low-mass galaxies~\citep[e.g.][]{Stierwalt2015,Pearson2016,Privon2017,Kristensen2021,Martin2021,Luber2022,Martin2022,GuzmanOrtega2023,ByrneMamahit2024,KadoFong2024}.
 
@@
 %     - Different merger mechanics would be really interesting, i.e. if the DM halo profile (cuspy/core etc) impacted the merger processes
-
-% 5 - Merger timescales have been studied a lot (mostly in obs and theory of high mass pairs)
-%     - Lotz 2011
-%     - RG2015
-%     - Snyder (maybe?) 
-%     - Jiang 2008 \& BK 2008 for all things about the "Chandra" timescale for merging galaxies which scales with rc/Vc and is thus prop to the age of the universe (eq.
-4\&5).
-Also quantified the circularity of orbits as a function of mass ratio $\leftarrow$ might be good to mention
-%     - Jiang 2014
 
@@
 we look at the orbits of the pairs from our previous work, and determine the merger timescale for low and high mass pairs as a function of pair separation and redshift.
-In sec.
-1...
-etc.
 
 %%% OUTLINE
+In Sec.~\ref{sec:methods}, we detail our selection criteria for isolated low-mass and high-mass orbits in the TNG100 simulation. 
+We examine the evolution of the number of pairs and merger fraction of our sample as a function of redshift in Sec.~\ref{sec:pairprops}.
+In Sec.~\ref{sec:results}, we present our findings on the separation and redshift dependence of the merger timescales of low-mass and high-mass pairs, and study the impact of physical and scaled separation criteria for pair selection.
+Finally, we discuss the implications of this work and provide suggestions for a self-consistent way of studying pairs across different redshifts and mass scales in Sec.~\ref{sec:discussion}, and present our final conclusions in Sec.~\ref{sec:conclusions}.
+
 \section{Methodology} \label{sec:methods}
@@
 
-We require that subhalos constituting a pair meet a minimum subhalo mass 
-% GB added Mhalo
-($\Mhalo$)
-criteria of 
+We require that subhalos constituting a pair meet a minimum subhalo mass ($\Mhalo$) criteria of 
 \begin{equation*}
@@
 
-% GB added 
-As in \chambe{}, 
-stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}.
+As in \chambe{}, stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
+Utilizing abundance matching to assign stellar masses allows us to circumvent simulation-specific stellar mass effects, and also results in a more straightforward process for applying our methodology to observational studies (see Sec.~\ref{sec:disc-suggestions} for more details).
 The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo via the abundance matching prescription  $\ms{}=f(\Mpeak,z)$, where $z$ is the redshift of consideration.
@@
 We calculate the number of pairs by selecting all pairs at a given redshift that have not merged.
-% The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
-%For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$.
-A single, non-merging pair with first infall at $z=3$ will thus contribute to the number of pairs at all redshifts from $z=0-3$.
+A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.
 
@@
 As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
+Note, this definition of the merger fraction differs from that typically used in observational studies, where merger fractions are computed from close pair fractions with a correction term.
 
@@
 Between $z=0.5-2.5$, the high-mass merger fraction is higher than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift. 
-This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$, while the number of high-mass pairs continues increasing to $z\sim1$.
-At these redshifts, the number of merging pairs begins declining rapidly. 
+We discuss this difference in more detail in Sec.~\ref{sec:disc-nonmergers}.
+% This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$, while the number of high-mass pairs continues increasing to $z\sim1$.
 % When the number of pairs stops increasing and beings to decrease, pairs that will not have time to merge before $z=0$ are being added, while many merging pairs are being removed from the sample, thus decreasing the merger fraction.
-Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.
+% Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.
 % move the above to discussion and end with something shorter: 
 % This COULD be that the timescales. This will answer this with this study and discuss this in detail in Disc.~\ref. 
+% could this be related to the number of non-mergers? is there a larger fraction of those starting at z=2 for the low mass galaxies?  these wouldn't be accounted for in Figure 6 right?
+% So fig 6 shows that this isn't due to the merger time being significantly different.
+if anything the higher mass pairs merge longer.
+you can state that at face value this difference in the knee may point to a difference in merger timescales, which you will illustrate in this paper is not the culprit.
 
@@
 
-\section{Results: The Mass and Redshift Dependence of Merger Timescales}
+\section{Results: The Mass and Redshift Dependence of Merger Timescales}\label{sec:results}
     Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
-    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the redshift at which pairs are selected and the merger redshift.\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
+    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the redshift at which pairs are selected and the merger redshift.\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}\textsuperscript{,}\footnote{We choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing.
+However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.}
     Only the merging pairs will be considered for the remainder of the analysis.
@@
     We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
-    % Pairs with small separations may be undergoing a pericentric passage, or very close to merging if on a radial orbit.
     Binning the pairs by separation, we can study how the merger timescale changes for pairs selected at different points in their orbits.
@@
     We find that the time until merger for low-mass and high-mass pairs is positively correlated with the separation of the pair at each redshift. 
-    However, the slope of this correlation decreases at higher redshifts. 
-    % For the same physical separation, the merger timescale is lower than for low-mass 
-    % This does that the overall merger timescales are lower for high-mass systems. 
-    % This is discussed in detail in the next section.
+    % However, the slope of this correlation decreases at higher redshifts. 
+    For the same physical separation, the merger timescale of high-mass pairs is shorter than for low-mass. 
     Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.
@@
     The high-mass pairs also have a larger spread in the distribution of merger timescales and separations at low redshift than at higher redshift.
-    This is due to the growth of halos over time, such that more pairs are identified at low redshifts where the virial radius of FoF groups is larger.
+    This is due to the growth of halos over time, and thus an increasing virial radius for FoF groups at lower redshifts.
 
@@
         The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$.
+        \kc{add h(z) line}
         }
@@
         As $z\to0$, there is an increasingly small fraction of pairs that merge before $z=0$, as seen by the dashed lines in Fig~\ref{fig:numorbits}. 
-        These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$.
-These mergers must then proceed on decreasing timescales by definition, as they were selected as pairs that must merge prior to $z=0$ (i.e., a pair chosen by the selection criteria given in Section XYZ at z=0.2 must merge by $z=0$, giving a maximal merger timescale equivalent XYZ Gyr whereas a pair chosen at z=6 can have a maximal merger timescale of XYZ Gyr.).
+        % These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$. 
+        These mergers must also proceed on shorter timescales by definition, as they were selected as pairs that must merge prior to $z=0$ (i.e., a pair at $z=0.1$ that merges by $z=0$ can have a maximal merger timescale of $1.47\Gyr$, whereas a pair at $z=1$ can have a maximal merger timescale of $8\Gyr$).
         This is shown by the dotted black line on the left of Fig.~\ref{fig:timescales}, which shows the lookback time of the simulation as a function of redshift.\footnote{The lookback time at a given redshift $z_n$ is equivalent to the time elapsed between $z_n$ and $z=0$.
-Thus, merging pairs cannot have merger timescales larger than the lookback time of the simulation at a given redshift.} % this could go in the methods section, if we think it's more necessary there.
-I don't mind it here for more detail
+Thus, merging pairs cannot have merger timescales larger than the lookback time of the simulation at a given redshift.}
 
         In principle, merger timescales\footnote{Specifically, the elapsed time between crossing into the virial radius of the primary subhalo and coalescence with the primary.} for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}.
-        Departures from the idealized case introduce perturbations to that solution, the validity of which has been tested in cosmological hydrodynamic and N-body simulations~\citep{Jiang2008, Boylan_kolchin2008}.
+        Departures from the idealized case introduce perturbations to that solution, the validity of which has been tested in cosmological hydrodynamic and N-body simulations~\citep{Jiang2008, BoylanKolchin2008}.
 
@@
 
-        In Fig.~\ref{fig:timescales}, the black dashed line represents a merger timescale that scales with $1/H(z)$,
-        \footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation. Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
-        and which is consistent with our findings for redshifts between $z=1.5-6$.
+        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by 0.25. 
+        We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the overall scaling.
+        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=1.5-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation.
+Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
+
+        It is notable that such a simple scaling holds for both low-mass and high-mass pairs, especially in a complex simulation like TNG100 where subhalos are far from isotropic spherical distributions of mass, and that low-mass and high-mass pairs have the same merger timescales.
+        We believe this is due to the dominance of dynamical friction in determining the redshift behavior of merger timescales. 
+        % Thus, the primary mechanics that determine merger timescales are likely dynamical friction
+
+        % maybe discuss Poulton 2023 somewhere in here -- add after send to prelim committee
 
@@
         Both separation selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \subfind{} catalogs.
-        This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates (see Sec.~\ref{sec:discussion} for further detail).
+        This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates~\citep[see e.g.][and observational studies cited therein]{Lotz2011,Snyder2017,Besla2018,Chamberlain2024}.
 
@@
 \section{Discussion} \label{sec:discussion}
-    We have calculated the merger timescales, or time until merger, for a sample of physically collocated low-mass and high-mass pairs from $z=0-6$.
+    We have calculated the merger timescales, or time until merger, for a sample of physically co-located low-mass and high-mass pairs from $z=0-6$.
     In this section, we will explore the implications and broader impacts of our findings, and put our study in context with previous work.
-    %In Sec.~\ref{}, etc...
 
     In Sec.~\ref{sec:disc-nonmergers}, we will discuss the non-merger portion of our orbit sample and the difference between low- and high-mass merger fractions.
-    In Sec.~\ref{sec:disc-previous}, we discuss our work in the context of other studies that have sought to reconcile xx xmerger rates and merger timescales via a change in pair selection criteria. 
-    In Sec.~\ref{sec:disc-suggestions}, we discuss the implications of our work for future observational and theoretical merger fraction and merger rate studies and provide suggestions for a self-consistent framework moving forward.
-    Finally, in Sec.~\ref{sec:disc-phys}, we briefly discuss the underlying physical mechanisms implied by our results and the functional form of the merger timescale.
+    In Sec.~\ref{sec:disc-pairfrac}, we return to one of the key questions posed in our previous work: what is the underlying physical mechanisms that drives the difference in pair fraction evolution between low- and high-mass pairs.
+    Finally, in Sec.~\ref{sec:disc-suggestions}, we discuss the implications of our work for future observational and theoretical merger fraction and merger rate studies, and provide suggestions for a self-consistent framework moving forward.
 
@@
         In Sec.~\ref{sec:pairprops-frac}, we found that the merger fraction of low-mass and high-mass pairs at $z>3$ is $>0.95$. 
-    This changes as you move to lower redshifts, where the merger fraction decreases substantially. 
+        The merger fraction decreases substantially as you move to lower redshifts. 
         While the decline in the merger fraction at $z\lesssim1$ is primarily due to the limited time that a pair will have to merge before the end of the simulation, we also find that between $z=\sim0.6-2.5$ the fraction of low-mass pairs is lower than that of high-mass pairs.
 
-    One potential explanation for this behavior could be that low-mass pairs have shorter merger timescales than high mass pairs, and thus merge at earlier times in the simulations, thus leading to a decrease in the merger fraction at higher redshifts.
+        One potential explanation for this behavior could be that low-mass pairs have shorter merger timescales than high-mass pairs, and thus merge at earlier times in the simulations, thus leading to a decrease in the merger fraction at higher redshifts.
         However, through our analysis, we have shown that the merger timescales of our full low-mass and high-mass pair sample evolve in the same way from $z=0-6$. 
@@
         We suspect that the cause of the difference between low- and high-mass merger fractions is not a difference in the merging population itself, but rather in the non-merger population and its evolution over time.
-    As we did not distinguish our non-merging sample into "likely mergers" and "fly-by" interactions for the purposes of the study, we cannot say anything about how these two populations individually contribute to the difference in the merger fraction.
-    However, one possible explanation could be that a larger fraction of infalling low-mass pairs at lower redshift result in fly-by interactions. 
-    This could be the case if the velocity at which secondary subhalo enter the primary's FoF group gets increasingly large compared to the virial velocity\footnote{The virial velocity is the value of the circular velocity at the virial radius.}, resulting in fewer bound systems.
-    We leave this as a focus of future work.
+        As we did not distinguish our non-merging sample into ``likely mergers" and ``fly-by" interactions for this study, we cannot say anything about how these two populations individually contribute to the difference in the merger fraction.
+        However, one possible explanation could be that a larger fraction of infalling low-mass pairs at lower redshift results in fly-by interactions. 
+        This could be the case if the velocity at which the secondary subhalo enters the primary's FoF group becomes increasingly large compared to the virial velocity\footnote{The virial velocity is the value of the circular velocity at the virial radius.}, resulting in fewer bound systems.
+        % We leave this as a focus of future work.
 
@@
 
-    \subsection{Steps Towards Standardized Pair Selection Criteria}~\label{sec:disc-previous}
-    Comparisons between studies of pair fractions, merger fractions, and merger rates across cosmic time and different mass scales are often made challenging by the use of different sets of selection criteria in each study.
+    \subsection{Connections to Pair Fractions}
+        In our previous work~\citep{Chamberlain2024}, we found that the pair fractions of isolated low-mass and high-mass pairs evolve very differently from $z=0-2.5$. 
+        Low-mass major pair fractions are constant for $z>2.2$, but decrease by almost 70\% from $z=3$ to $z=0$ (see their Fig.~3). 
+        High-mass major pair fractions slowly increase from $z=6$ to $z=0$, with a more significant increase from $z=0.25$ to $z=0$.
+
+        One of the hypotheses we laid out in~\chambe{} for the difference in the behavior of the pair fractions is that perhaps low- mass and high-mass pairs have different merger timescales, which could lead to a difference in pair fractions from high to low redshift.
+
+        However, in this study, we have shown that the merger timescales of low-mass and high-mass pairs are approximately equal at all redshifts, particularly when considering the full pair sample or the scaled separation selected sample.
+        This means that different merger timescales are not responsible for the difference in the evolution of pair fractions. 
+        Rather, we suspect that the build-up of larger structures due to hierarchical formation under $\Lambda$CDM leads to a reduction in the total number of low-mass groups at lower redshift $z\lesssim2$.
+
+        Additionally through this study, we have found that a very large fraction of the pairs selected in our previous work do indeed merge ($>90\%$ of pairs for $z>1.6$), so that the rate at which mergers occur is likely larger than the rate at which isolated low-mass pairs form, particularly at low redshift, thus leading to a quickly declining pair fraction measured in our previous work.
+
+    \subsection{Implications and Suggestions for Future Observational and Theoretical Studies of Pairs}~\label{sec:disc-suggestions}
+        While we are unable to make meaningful comparisons with previous observational studies of merger fractions and merger rates, as our study is completed using true physical separations from the simulation rather than projected separations, we can still draw meaningful conclusions about the application of our results to future studies of merger timescales and merger rates.
+
+        \subsubsection{Pair Selection Criteria and Merger Rates in Illustris}
+            Comparisons between studies of pair fractions, merger fractions, and merger rates, especially across cosmic time and different mass scales, are often challenged by the use of different sets of selection criteria in each study.
+
             On the observational side, pair selection criteria are often set by the observational parameters of the survey itself. 
-    For example, limiting magnitudes and completeness limits can enforce a specific new range of stellar mass ratios to be considered (for example $1-1:6$ instead of $1:10$), while avoiding colliding/overlapping fibers on a spectroscopic survey might lead to very specific separation criteria.
-
-    On the theoretical side, it can be straightforward to adopt a number of different pair selection criteria for comparisons to specific observational studies with identical criteria. 
+            For example, limiting magnitudes and completeness limits can enforce a specific new range of stellar mass ratios to be considered (for example $1-1:6$ instead of $1:10$), while avoiding fiber collisions on a spectroscopic survey might lead to very specific separation criteria.
+            On the theoretical side, it can be more straightforward to adopt a number of different pair selection criteria for comparisons to specific observational studies with identical criteria. 
             However, this means that theoretical studies typically restrict themselves to a subset of pair selection criteria specific to their target comparison studies.
 
-    A variety of studies in recent years aim to standardize pair selection criteria for future studies, in particular to permit a more fair comparison between theory and observations. 
-    One such study is~\citet{Ventou2017}, which created mock catalogs from the original Illustris-1 simulation to analyze the merger probability of galaxy pairs as a function of projected separation and projected relative velocity.
-    They find that 
-    to provide a new set of selection criteria to define close pairs. 
-    One
+            In recent years, there have been works that aim to standardize pair selection criteria for future studies, in particular to facilitate a fair comparison between theory and observations. 
+            One such study is~\citet{Ventou2019}, which created mock catalogs from the original Illustris-1 simulation to analyze the merger probability of galaxy pairs as a function of (both physical and projected) separation and relative velocity.
+            They find that pairs with projected separations between $5\leq r_{proj} \leq 50\,kpc$ and projected relative velocities of $v_{proj}\leq300\,\kms$ selects all pairs that have greater than 30\% change of merging before $z=0$, and provide a weighting scheme to determine the likelihood of merger as a function of projected separation and velocity.
+            These selection criteria are thus calibrated to select a population of pairs that are fairly likely to merge, and a way to calibrate merger fractions more accurately from close pair fractions.
+            However, \kc{I think I need to say something here about what the shortcoming of this framework in the above paper is (if there is one) }.
 
-    Merger rates/merger fractions in illustris, close pair fractions and merger timescales, etc. 
-    As in our previous work on pair fractions in TNG100, we are unable to make meaningful comparisons with previous observational studies of merger fraction and merger rates, as our study is completed using true physical separations from the simulation.
-    However, 
-    % In~\cite{RG2015}, they quantify the galaxy major merger rate as a function of the stellar mass of the descendent subhalo. 
-    % It is unclear, however, if the major merger rate for galaxies with stellar mass $\ms<10^{10}\Msun$ is actually discrepant between simulations and observations, or if the observational sam
+            % \citet{OLeary2021a} used the \textsc{Emerge} empirical model to show that the true merger rates of galaxy pairs in their simulation cannot be captured by using pair selection criteria that are too restrictive.
+            % Although their findings come from an approximation for observability timescales calibrated from merger timescales,  rather than merger timescales specifically, they find that projected separation criteria of $5\leq r_{\rm proj}<30\kpc$ underestimates the merger rate for $z=0-5$.
 
-    NOT a lit review of merger topics etc. 
-        \begin{itemize}
-            \item Have they studied merger timescales as a function of mass and redshift? 
-            \item What is unique about our study?
-            \item Are comparisons with previous works fair, or are they not really comparable?
-            \item specific papers: Ventou 2017, Hopkins 2007, O'Leary 20
-            \item only look at merger rates on the low mass end ? - garreth 
-            \item studies that redefine pair fraction criteria 
-            \item What other studies have looked at merger fractions/timescales in hydro simulations, specifically? 
-            \item Have they studied merger fractions as a function of mass and redshift? What did they find?
-            \item Were low-mass pairs expected to have such high merger rates in isolated environments? 
-        \end{itemize}
+            % \textit{RG2015} 
+            % the reason I want to talk about the RG15 paper is because its an example of trying to put galaxy merger rates on the same plots, and using the same selection criteria at high and low masses may have also resulted in such a discrepancy, IN ADDITION to the problems that low mass morphology studies face when trying to use morphological signatures for low-mass merger fraction studies.
 
-    \subsection{Implications for Future Observational and Theoretical Studies of Pairs}~\label{sec:disc-suggestions}
+            \citet{RG2015} constructed galaxy merger trees to calculate the galaxy merger rate as a function of the stellar mass of the descendent subhalo and redshift. 
+            They find that the galaxy merger rate increases with increasing redshift for descendent galaxy masses $\ms{}>10^{10}\Msun$. 
+            These findings are consistent with predictions of the galaxy merger rate from semi-empirical models~\citep{Stewart2009,Hopkins2010}, and with merger rates derived from observational merger fraction measurements calibrated by the corrected observability timescales of \citet{Lotz2011}.
+            They additionally find that the major merger rate at $z\sim0.1$ increases with increasing descendant subhalo mass.
+            While the redshift evolution of merger rates are broadly consistent with observations, they find significant discrepancy with major merger rates of galaxies with $\ms<10^{10}\Msun$ from~\citet{Casteels2014}, which used morphological merger signatures to estimate merger rates at low $z$.
+            Since the~\citet{RG2015} work is done self consistently at all redshifts and descendant masses, they suggest that the discrepancy arises from an extrapolation of gas fraction data available for $\ms<10^{10}\Msun$ to galaxies with lower mass, which will be reflected in observability timescales with significant uncertainty in the lower mass regime.
+
+            These results imply that the timescales used to convert observational merger fractions (determined either by weighted close pair fractions or by morphologically selected pairs) are sensitive to the mass ranges and redshifts at which they are valid, and attempting to make comparisons between theoretical predictions and observational measurements can lead to seemingly incompatible results.
+            Our work provides a unique framework that can allow for more robust comparisons of theoretical and observational samples in the future, since the observational samples themselves will be able to also self-consistently apply merger timescales to close pair fractions using our suggested scaled separation criteria.
+
+        \subsubsection{Implementing Scaled Separation Pair Selection Criteria in Future Studies}
+            In our previous work, we studied the impact of different sets of separation criteria on the recovered pair fractions of low-mass and high-mass pairs, and found that only by utilizing scaled separation criteria can the features of the pair fractions' evolution with redshift be distinguished~\citep{Chamberlain2024}.
+            In this work, we explored how these same set of selection criteria impacts the recovered merger timescales.
+
             In Sec.~\ref{sec:results-timevredshift}, we find that the merger timescales of isolated low-mass and high-mass pairs evolve nearly identically between $z=0-6$, rising from a median of $\sim0.8Gyr$ at $z=6$ to $2.1-2.4\Gyr$ at $z\sim0.6$.
             From $z=0$ to $z=0.6$, the merger timescale rapidly declines, since the only mergers at low redshift that can proceed before the end of the simulation at $z=0$ are those with very short timescales.
-
-    In our previous work, we studied the impact of different sets of separation criteria on the recovered pair fractions of low-mass and high-mass pairs, and found that only by utilizing scaled separation criteria can the features of the pair fractions' evolution with redshift be distinguished~\citep{Chamberlain2024}.
-    In this work, we explored how these same set of selection criteria impacts the recovered merger timescales.
 
@@
             For example, the median merger timescale for close pairs selected with $\rsep<70\kpc$ varies by up to $\sim0.6\Gyr$ between low- and high-mass pairs at all considered redshifts, while the merger timescale for close pairs selected via the scaled criteria $\rsep<0.5\Rvir$ have nearly identical median timescales.
+            This is because selection criteria based on scaled separations take the equivalent \textit{fraction} of the full volume of each FoF group, regardless of mass or redshift, whereas physical separation criteria will select the same volume around each primary, with no account for the growth of dark matter halos or spatial distributions of satellites.
 
-    Our results demonstrate the importance of utilizing pair selection criteria that permit fair comparisons of pair properties in future observational studies (i.e.
-pair fractions, merger timescales, merger rates, etc.), particularly when the goal of such studies is to quantify and compare the redshift evolution of pair properties at different mass scales.
-    We thus promote the adoption of separation criteria that vary as a function of mass and redshift for such future studies. 
-    In particular, we suggest employing a separation criteria based on the scaled separation of the system.
+            These results translate to important implications for observational studies of merger rates, which are typically determined using physical separation selected close pairs and averaged merger timescales.
+            Our work has shown that both of these quantities can vary significantly for different pair selection criteria, and that physical separation criteria can wipe out the distinguishing features of pair fractions or the equivalence of merger timescales between low-mass and high-mass pairs.
+
+            We therefore promote the adoption of separation criteria that vary as a function of mass and redshift for future pair studies. 
+            In particular, we point out the importance of utilizing pair selection criteria that permit fair comparisons of pair properties in observational studies (i.e.
+pair fractions, merger timescales, merger rates, etc.), particularly when the goal of such a study is to quantify and compare the redshift evolution of pair properties at different mass scales.
+            Specifically, we suggest employing a separation criteria based on the scaled separation of the system.
 
             Although this seems to introduce further complications to the pair selection process in observational studies, the use of a scaled separation as demonstrated here can be applied in a fairly straightforward way.
-    In particular, given an observation of a pair with stellar masses $\ms1$ and $\ms2$, we can first compute the associated dark matter mass of the pair using the SMHM relation. 
-    The SMHM relationship of Moster 2013 accounted for observational stellar mass errors  when fitting their relationship, which means that the impact of using the SMHM to convert from stellar mass to halo mass introduces the same error as it does in this study (but we didn't use it in this study technically...
-think about this).
+            In particular, given an observation of a pair with stellar masses $\ms1$ and $\ms2$, we can use the SMHM relation to calculate the associated dark matter masses, $M_{h1}$ and $M_{h2}$, of the galaxy pair.
+            This introduces the same systematic error to an observational study as it does when assigning stellar masses to our dark matter halos that utilizes the abundance matching prescription of~\cite{Moster2013}, since they have accounted for observational stellar mass errors when constructing their SMHM relationship.
+            Indeed, the SMHM relationship can be sampled in a similar fashion as in~\chambe{} to get a spread on the error associated with the abundance matching process.
 
-    Once you have the dark matter masses of the two subhalos, you can compute the total dark matter mass, and get an estimate of the overall virial radius of the group by assuming a FoF mass equivalent to the sum of the dark matter masses.
-    Thus, coming up with virial radius criteria per pair shouldn't in principle be that hard, especially since stellar masses are already needed to be calculated to claculate the stellar mass ratio of observed pairs anyways.
+            One can determine an approximate virial radius for a given observed isolated pair by finding the virial radius of a subhalo with total mass $M_{h1}+M_{h2}$.\footnote{We showed in~\chambe{} that the virial radius of a halo with mass given by the combined dark matter halo masses of the primary and secondary recovers the virial radius of the FoF group with approximately 98\% accuracy.}
+            Since observed stellar masses already need to be computed in order to quantify the mass ratio of each pair, no further information from the observations are needed to create the mass and redshift evolving separation cuts.
 
-    Our study holds for isolated pairs, however we have not extended this analysis to more high-density environments. 
-    Some strengths of this approach are that we can study the pair evolution behvaior as a function of redshift, and the dynamics will be inherent to the pair rather than 
-    Some weaknesses are that there is a chance that the low-mass and high-mass merger timescale of high density environments vary/evolve completely independently of our findings. 
-    We leave this as the focus of future work.
+            As a concluding note, we want to remind readers that our results are explicitly for isolated pairs. 
+            One strength of our approach to only consider isolated pairs is that we have been able to study dynamic that are inherent to the pairs themselves, rather than those that are a product of the environment.
+            It is possible that low-mass and high-mass pair merger timescales in high density environments evolve completely differently than those found in our findings. 
+            We leave the extension of this analysis to more high-density environments as the focus of future work.
 
-    Bring up assumption of isolation criteria and it's drawbacks and strengths.
-Reiterate that our findings are for isolated pairs.
--> Focus of future work to explore the impact of pair selection criteria on pairs in a variety of different galaxy environments.
-            \begin{itemize}
-            \item How did we motivate separation criteria?(motivated already in Chambe+ 2024, restate here about close pair definitions from lit.)
-            \item Pair fractions can vary wildly depending on separation criteria
-            \item This is important too for merger rates, since they are a function of pair fractions and merger timescales, and you should really be selecting pairs at each stage as self-consistently as possible
-            \item What is our suggestion moving forwards?
-            \item specific papers: Lotz 2011, 
-            % As discussed in Chamberlain, this is doable, the abundance matching prescription was formulated with the stellar mass errors already taken into account, so this isn't introducing significantly more error in the analysis
-            % quote the stellar mass errors being folded in at high z 
-        \end{itemize}
-
-   \subsection{The Underlying Physical Behavior of Merger Timescales}~\label{sec:disc-phys}
-
-        \begin{itemize}
-            \item What can we say about the fact that the merger timescales follow a $(1+z)^{-1}$ relation at $z>0.5-1$? Physical explanation for this behaviour? 
-            \item Scaled separation criteria are akin to picking the same fraction of the total volume of a FoF group, regardless of what redshift it is at, not what mass it has, so it may not be surprising that we got what we did in re: the scaled separation being better, but finding the SAME merger timescales for hi and low is interesting!
-            \item relevant papers: 
-        \end{itemize}
-
-\section{Summary and Conclusions}
+\section{Summary and Conclusions}\label{sec:conclusions}
     % what we did and how
-    In this study, we construct a sample of the orbits of low-mass ($10^8 <\ms<5\times10^{9}\Msun$) and high-mass ($5\times10^{9}<\ms<10^{11}\Msun$) major pairs (stellar mass ratio $> 1:4$) from $z=0-6$ in the TNG100 simulation.
+    In this study, we construct a sample of the orbits of low-mass ($10^8 <\ms{}<5\times10^{9}\Msun$) and high-mass ($5\times10^{9}<\ms{}<10^{11}\Msun$) major pairs (stellar mass ratio $> 1:4$) from $z=0-6$ in the TNG100 simulation.
     Orbits of pairs, i.e.
-the separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo  first share a common FoF group (or have experienced 'first infall') to either $z=0$ or the redshift at which the pair merges.
+the separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo  first share a common FoF group (or have experienced `first infall') to either $z=0$ or the redshift at which the pair merges.
 
@@
         % merger fractions
-        \item The merger fraction of low-mass and high-mass physically associated pairs is high ($>0.9$) at $z>\sim1.5$ (see Fig.~\ref{fig:fmerge}.
+        \item The merger fraction of low-mass and high-mass physically associated pairs is high ($>0.9$) at $z\gtrsim1.5$ (see Fig.~\ref{fig:fmerge}.
 However, the merger fraction declines rapidly to 0 as $z\to0$, due to the finite length of the simulation which artificially reduces the number of pairs with enough time left to merge at low redshifts.
@@
 Additionally, for a given physical separation, high-mass pairs have a shorter merger timescale than low-mass pairs at the same redshift with the same separation.
-For example, at $z=2$, high-mass pairs at $100kpc$ merge within $\sim 0-3 \Gyr$, while low-mass pairs at $100\kpc$ merge within $\sim 0.5-5\Gyr$ (see Fig.~\ref{fig:2dhist}).
+For example, at $z=2$, high-mass pairs at $100\kpc$ merge within $\sim 0-3 \Gyr$, while low-mass pairs at $100\kpc$ merge within $\sim 0.5-5\Gyr$ (see Fig.~\ref{fig:2dhist}).
         % Median merger timescale
         \item The median merger timescale for the full sample low-mass and high-mass pairs that merge prior to $z=0$ declines with increasing redshift at a rate proportional to $1/H(z)$ for redshifts $z\gtrsim1$ (see Fig.~\ref{fig:timescales}).
-        The maximum merger timescale for a low-redshift pair ($z<1$) is constrained by the time remaining in the simulation for the pair to merge, which causes the median merger timescale to decline to 0 at $z=0$.
+        The maximum merger timescale for a low-redshift pair ($z\lesssim1$) is constrained by the time remaining in the simulation for the pair to merge, which causes the median merger timescale to decline to 0 at $z=0$.
         The decrease in the merger time at low redshifts is artificial and not representative of the true merger timescales that pairs at $z=0$ will have.
@@
         % The separation criteria effect 
-        \item \kc{left off here}When pairs are selected via a maximum separation cut
-        static physical separation-selected subsamples of low- and high-mass pairs have different median merger timescales.
-        \item scaled separation-selected subsamples yield nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts. 
+        \item When pairs are selected via a scaled separation criteria, namely the pair separation scaled by the virial radius $\Rvir$ of the FoF group of the primary, low-mass and high-mass merger timescales are nearly identical at all redshifts.
+This holds for all scaled separations considered $[0.25,0.5,1,1.5,2,3]\,\Rvir$.
+At $z=[0, 1, 2, 3, 4]$, a scaled separation of $1\,\Rvir$ corresponds to an average physical separation of $[134, 85, 59, 43, 33]\kpc$ for low-mass pairs and $[348, 206, 134, 97, 76] \kpc$ for high-mass pairs.
+        \item When pairs are selected via a physical separation criteria (one that does not vary with redshift or with mass), the median merger timescales of low-mass and high-mass pairs differ by up to $0.8\Gyr$.
+Thus, using the same merger timescale for low-mass and high-mass close pair samples selected via a physical separation cut will result in incorrect merger rate measurements.
     \end{itemize}
 
-    % set of conclusions for discussion
-    - We find that the merger timescale of low-mass or high-mass merging pairs is sensitive to the selection criteria that are used to define the pair sample.\\
+    Studies have found that the merger rates of galaxy pairs varies with redshift and with the mass of the pair~\citep{Stewart2009,Hopkins2010a,RG2015}.
+    In our previous work, we found that the pair fraction of isolated pairs in TNG100 likewise evolves with redshift and mass~\citep{Chamberlain2024}. 
+    In this paper, we show that the merger timescales of major pairs in TNG100 vary with redshift, but that low-mass and high-mass pairs have equal merger timescales at all redshifts from $z=0-6$ if the correct separation selection criteria is used to pick equivalent samples of pairs.
 
-    - We recommend that future studies that seek to compare or compute pair fractions, merger fractions, merger timescales, or merger rates, define selected pair samples using separation criteria that evolve with redshift and mass.
-
-    - Why is this important?
-Cause future studies are gonna get so many pairs, especially low mass, and now we have a framework to compare low and high mass pairs in a variety of ways, from pair fractions to merger fractions to merger timescales!
-Woo!
-
-    Further work that would be nice to do would be to see if all the timescale stuff stays the same for minor mergers, and for smaller bins in range 1-1:4
+    These works together give a comprehensive framework for inferring merger rates via close pair fractions and merger timescales at redshifts from at least $z=0-6$, though likely higher redshift, and for quantifying the differences between low-mass and high-mass pairs.
+    Indeed, separation selection criteria that scale with the mass and redshift of the target system are crucial for interpreting pair properties in a self-consistent way. 
+    In the future, observatories like Rubin Observatory, the Roman Space Telescope, JWST, and future ELTs, will detect an abundance of low-mass pairs at a wide range of redshifts, and our theoretical framework for interpretting these observations will be more critical than ever.
 

### 2024-03-30 paper/main.tex
edits after full read through

@@
 
-\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time
-}
+\title{A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time}
 \shorttitle{Merger Timescales in TNG100}
@@
     % Statement
-    % Evolutionary differences between low-mass ($\rm 10^8<M_*<5\times10^9\,\Msun$) and high-mass ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) galaxy pairs have been studied in detail for the last few decades, with studies finding differences in star formation rates, gas fractions, and dark matter halo profiles.
-% could remove this line
-    % Problem 
-    The merger rates of high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively, both observationally and theoretically. 
-    Since low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) are comparatively under studied, it is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems.
-    However, it is unclear if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift.
+    The merger timescales of isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) on cosmologically motivated orbits have not been studied in detail, though isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively.
+    It is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems, however it is unclear if their merger timescales are similar, nor if they evolve similarly with redshift.
     % Our solution
-    Here, we present a physically-motivated framework by which merger timescales for high-mass and low-mass pairs can be compared equivalently, 
-    %and converted from one system to another 
-    %by scaling 
-    % GB added below
-    wherein separation criteria are scaled
-    by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 R_{\rm vir}$). 
-    We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass major pairs as a function of cosmic time, and compare to the merger timescales of similarly-selected high-mass major pairs.
-    Using our framework to select pairs via a scaled separation criterion leads to equivalent merger timescales for low- and high-mass systems at redshifts $z=0-6$. 
+    We use the Illustris TNG100 simulation to quantify the merger timescales of isolated low-mass and high-mass major pairs as a function of cosmic time, and explore how different selection criteria impact the mass and redshift dependence of merger timescales.
     % What are the timescales? 
-    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts, particularly for the closest pairs ($\rsep<150\kpc$), leads to timescales that differ by up to $\sim1\Gyr$ between low- and high-mass pairs.
-    % Why it matters 
-    Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system.
-    As a result, applying the same merger timescales to different mass systems will lead to a bias that systematically (over/under) predicts low-mass galaxy merger rates. 
+    We find that the merger timescales for low- and high-mass pairs are the same from $z=0-6$ in the absence of any separation criteria.
+    We then present a physically-motivated framework for selecting pairs via a scaled separation criteria, wherein pair separations are scaled by the virial radius of the primary's FoF group halo ($r_{\mathrm{sep}}< 1 \Rvir$).
+    Applying these scaled separation criteria recovers equivalent merger timescales for both mass scales at all redshifts.
+    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts leads a difference in merger rates of up to $\sim1\Gyr$ between low- and high-mass pairs, particularly for $\rsep<150\kpc$.
+    % Importance
+    As a result, applying the same merger timescales to physical separation-selected pairs will lead to a bias that systematically over-predicts low-mass galaxy merger rates. 
+    % could replace the above line with something softer suggesting the application of scaled selection criteria for studies of merger properties of pairs at different mass scales and redshifts, or something like the below:
+    % Our findings suggest that the comparison of low-mass and high-mass galaxy pair fractions, merger timescales, and merger rates is self-consistent using separation criteria that scales as a function of mass and redshift of the system.
+    % if using the same separation criteria to select low-mass galaxy pairs for such studies, or if using the same separation criteria at different redshifts, permits an equitable comparison between galaxies of different mass and redshift.
+
 \end{abstract}
@@
 \section{Introduction} \label{sec:intro}
-Galaxy merger rates provide an important pathway for tests of hierarchical assembly from$\Lambda$CDM theory, and are critical for understanding the formation and evolution of galaxies across time~\citep[e.g.][]{Stewart2009, Hopkins2010b,RG2015}.
+Galaxy merger rates provide an important pathway for tests of hierarchical assembly from $\Lambda$CDM theory, and are critical for understanding the formation and evolution of galaxies across time~\citep[e.g.][]{Stewart2009, Hopkins2010b,RG2015}.
 In the era of future surveys with JWST, Rubin, and Roman, these tests will be extended to higher redshifts and lower mass scales than were previously accessible~\citep{Gardner2006,Spergel2015,Robertson2019a,Robertson2019b,Behroozi2020,Martin2022}.
 
-At higher redshift ($z>\kc{xx}$) and at lower masses ($\ms{} \lesssim10^{9.5}$), even isolated galaxies are highly disturbed~\citep{Wuyts2013,Martin2018,RG2019,Martin2021}, making close pair fractions a critical alternative to morphological signatures for merger rate studies.
+At higher redshift ($z\gtrsim1$) and at lower masses ($\ms{} \lesssim10^{9.5}$), even isolated galaxies are highly disturbed~\citep{Wuyts2012,Wuyts2013,Martin2018,RG2019,Martin2021}, making close pair fractions a critical alternative to morphological signatures for merger rate studies.
 However, there is currently no framework for comparing merger rates for low-mass and high-mass galaxies as a function of time, and tensions between predictions from theory and observational studies exist in the literature at low masses~\citep[see e.g.
@@
 
-% 4 - Comparing the merger rates of high mass and low mass pairs is important why?
-%     - Is important for our understanding of the build up of higher mass things in LCDM, and thus a test of LCDM
-%     - Different merger mechanics would be really interesting, i.e. if the DM halo profile (cuspy/core etc) impacted the merger processes
+In our previous work, we show that the redshift evolution of pair fractions of isolated low-mass and high-mass pairs in Illustris TNG100 differ significantly, particularly at $z<3$ where pair fractions of low-mass pairs decrease up to $75\%$ between $z=3$ and $z=0$, while high-mass pair fractions peak at $z=0$~\citep{Chamberlain2024}.
+We suggested that differences in the merger timescales of low-mass and high-mass pairs might be the cause of the pair fraction evolution differences.
+In our following analysis, we will investigate this question directly and answer whether merger timescales are responsible for these pair fraction differences.
 
-% 6 - Problem is, we don't have a lot of studies on low mass pair merger rates? 
-%     - This is not ideal because it means that we typically extend merger rates calibrated for high mass down to low mass, but this means that we aren't able to test LCDM and draw conclusions about the impact
+Additionally, we found in \citet{Chamberlain2024} that recovering the pair fraction differences seen between the two mass scales is sensitive to the selection criteria adopted during pair selection.
+Physical separation cuts applied equivalently to low-mass and high-mass pairs at all redshifts eliminate the ability to distinguish the behavior of the true pair fraction evolution (see their Figure~5), even for separation cuts as large as $\rsep<300\kpc$.
+Alternatively, when separation criteria vary with the mass and redshift of each system, in particular when the separation is scaled by the virial radius $\Rvir$ of the pair's FoF group halo, the ability to distinguish the underlying pair fraction behavior was recovered for scaled separations as low as $\rsep < 0.5\,\Rvir$ (see their Figure~4).
+Employing scaled separation criteria then permits the equivalent comparison of pair fractions between low-mass and high-mass pairs, and robust comparisons across redshifts from $z=0-4$. 
+Since merger rate estimates are derived from close pair fractions and merger timescales, the results from our previous work imply that careful consideration of separation criteria is required for merger timescale and merger rate studies at different mass scales and redshifts as well.
 
-7 - In addition, our previous work showed that even the evolution of the fractions of high and low mass pairs can't be robustly studied and compared if even the selection criteria for the pairs are off.
-    - This means that the selection of pairs for pair fractions, merger fractions, and for merger timescales, are ALL going to be impacted by the selection criteria that's used. 
-    - This is also a problem in observations already, since different samples will have different completeness down to different projected separations, and with different mass ratios, etc. 
-    - O'Leary 2021 merging systems can be missed if you don't use wide enough sep criteria, thus leading to under predict merger rate
-
-8 - Fortunately, our previous work showed that using scaled criteria for pair selection as a function of mass and redshift allows you to fairly compare samples. 
-    - This motivated us to study the merger fraction and merger timescales of our galaxy pair sample from the first study, to see how merger timescales might ALSO depend on the selection criteria that are used to pick the pair sample.
-
-9 - Here, we aim to build a framework to understand the pair selection criteria that are used to select samples, and the merger timescales that result from various selections.
-we look at the orbits of the pairs from our previous work, and determine the merger timescale for low and high mass pairs as a function of pair separation and redshift.
+In order to investigate the impact of pair selection criteria on merger timescales specifically, we will track the pairs selected in \chambe{} forwards and backwards in time to construct orbits for each pair.
+We will then study the merger timescales of isolated low-mass and high-mass pairs for the same set of physical and scaled separation criteria used in \chambe{} to determine whether merger timescales of isolated pairs are robust to the selection criteria used to determine the pair samples.
 
 %%% OUTLINE
-In Sec.~\ref{sec:methods}, we detail our selection criteria for isolated low-mass and high-mass orbits in the TNG100 simulation. 
-We examine the evolution of the number of pairs and merger fraction of our sample as a function of redshift in Sec.~\ref{sec:pairprops}.
+In this paper, we aim to extend the framework for pair selection criteria presented in \chambe{}.
+%to examine the merger timescales that result from these various separation criteria.
+In Sec.~\ref{sec:methods}, we detail our selection criteria for isolated low-mass and high-mass orbits in the Illustris TNG100 simulation. 
+We examine the redshift evolution of the number of pairs and merger fraction of our sample in Sec.~\ref{sec:pairprops}.
 In Sec.~\ref{sec:results}, we present our findings on the separation and redshift dependence of the merger timescales of low-mass and high-mass pairs, and study the impact of physical and scaled separation criteria for pair selection.
-Finally, we discuss the implications of this work and provide suggestions for a self-consistent way of studying pairs across different redshifts and mass scales in Sec.~\ref{sec:discussion}, and present our final conclusions in Sec.~\ref{sec:conclusions}.
+Finally, we discuss the implications of this work and provide suggestions for a self-consistent way of studying pairs across redshifts and mass scales in Sec.~\ref{sec:discussion}, and present our final conclusions in Sec.~\ref{sec:conclusions}.
 
@@
 %recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
-We utilize the group catalogs, produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009}, and the merger tree catalogs, generated by the \sublink{} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
+We utilize the group catalogs produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009} and the merger tree catalogs generated by the \sublink{} algorithm~\citep{RG2015} from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
 The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).
 
-Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated, but physically associated, as in \citet{Chamberlain2024}.
+Our sample consists of major low-mass and high-mass pairs (stellar mass ratio $1/4 \leq M_{*2}/M_{*1}\leq 1$) that are isolated, but physically associated, as in \citet{Chamberlain2024}.
 From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.
@@
 
-At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive subhalos (by assigned stellar mass, using abundance matching; see below) from FoF groups with virial mass
+At each snapshot, low-mass and high-mass pairs are chosen by first selecting the two most massive subhalos (by assigned stellar mass, using abundance matching; see below) from FoF groups with virial mass $\MG$
 % GB added M_G
-($\rm M_{G}$)\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass.
+\footnote{We use \texttt{Group\_M\_TopHat200} from the TNG100 Group Catalogs as the FoF Group virial mass.
 This mass is defined to be the mass enclosed by a sphere with mean density $\Delta_c *\rho_c$, where $\Delta_c$ is the overdensity constant from~\citet{Bryan1998} and $\rho_c$ is the critical density of the universe at the time calculated.
@@
 \begin{align*}
-        \mbox{\textbf{low mass:}}&\,\rm M_{G} = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
-        \mbox{\textbf{high mass:}}&\, \rm M_{G}=1\times 10^{12}- 6.5\times10^{12}\,\Msun.
+        \mbox{\textbf{low mass:}}&\,\rm \MG = 8\times 10^{10}- 5\times 10^{11}\,\Msun \\ 
+        \mbox{\textbf{high mass:}}&\, \rm \MG =1\times 10^{12}- 6.5\times10^{12}\,\Msun.
 \end{align*}
@@
 
-We require that subhalos constituting a pair meet a minimum subhalo mass ($\Mhalo$) criteria of 
+We require the subhalos that constitute a pair to meet a minimum subhalo mass, \Mhalo, criteria of 
 \begin{equation*}
@@
 
-Primary subhalos are defined as the subhalo with the highest assigned stellar mass ($M_{*1}$) in their FoF group, and secondaries are defined as the second most massive subhalo with stellar mass $M_{*2}$.
+Primary subhalos are defined as the subhalo with the highest assigned stellar mass, $M_{*1}$ in the FoF group, and secondaries are defined as the second most massive subhalo with stellar mass $M_{*2}$.
 Our sample of major pairs then consists of all pairs of primary and secondary subhalos with 
 \begin{align*} 
-\mbox{\textbf{low mass primaries:}}&\, 10^{8}< \rm M_{*1} < 5\times10^{9} \Msun \\ 
-\mbox{\textbf{high mass primaries:}}&\, 5\times 10^{9}< \rm M_{*1} < 10^{11} \Msun\\
+\mbox{\textbf{low mass primaries:}}&\, 10^{8}< M_{*1} < 5\times10^{9} \Msun \\ 
+\mbox{\textbf{high mass primaries:}}&\, 5\times 10^{9}< M_{*1} < 10^{11} \Msun\\
 \mbox{\textbf{stellar mass ratio:}}&\,      
@@
 A primary or secondary subhalo can only be a member of one single pair at a given snapshot, such that a collection of $N$ pairs consists of $N$ unique primaries and $N$ unique secondaries.
-
 A subhalo can belong to different pairs at different redshifts. 
@@
 \subsection{Mergers} \label{subsec:mergers}
-Here we describe how subhalo mergers are identified. 
-Note that this definition is specific to subhalo mergers and therefore results are not guaranteed to hold for galaxy mergers, especially in cases where subhalos and galaxies have different centers of mass~\citep[see e.g.,]{RG2015}.
+Here we describe how mergers are identified. 
+Note that this definition is specific to subhalo mergers and therefore results are not guaranteed to hold for galaxy mergers~\citep[see e.g.,][]{RG2015,Patton2024}, especially in cases where dark matter subhalos and galaxies have different centers of mass, as is the case in our own Galaxy~\citep{Gomez2015,GaravitoCamargo2019,Petersen2021,Chamberlain2023}.
 % might  note that this is discussed in more deets in discussion - I think Patton 2024 addresses this a bit?
@@
 
-We utilize the \texttt{DescendantID} field of the \sublink{} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
+We utilize the \texttt{DescendantID} field of the \sublink{} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when each merger occurs. 
 The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
@@
 In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift.
-We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group (also in $\kpc$).\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs.
+We use the definition of scaled separation $\rsep/\Rvir$ from~\cite{Chamberlain2024}, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group (also in $\kpc$).\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs.
 If the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
-The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
-This ``scaled separation" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.
+The virial radius of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
+This ``scaled separation" is, by construction, a function of mass and redshift.
+%, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.
 
@@
 
-Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$.
-The top panel shows the orbits of five pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo.
+Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected from \chambe{} at $z=1.5$.
+The top panel shows the orbits of five pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit before the secondary and primary ever share a FoF group halo.
 Even amongst galaxies of the same approximate stellar mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\,\Gyr$ from first infall to merger (triangles).
 
-For some pairs, the secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo than that of their primary.
-Because all of our mergers occur when the subhalos are in the same FoF group by definition, in these cases, we use the full orbit starting when the subhalos are first identified as members of the same FoF group through when they merge, as described in Section \ref{subsec:mergers}.
+For some pairs, the secondary subhalos can experience first infall, then after one pericentric passage can return to large distances where they are temporarily assigned a different FoF group halo than that of their primary.
+In these cases, we use the full orbit beginning at first infall, including the sections of the orbit where the secondary is assigned to another FoF group temporarily,  through to merger as described in Section \ref{subsec:mergers}.
 This definition is more robust for considering the full interaction timescales of merging pairs.
@@
 The bottom panel shows the orbits of ``Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
-Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.
+Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of selected orbits.
 
@@
 Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts after infall, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists.
-However, we only want to keep a single instance of any given orbit in our catalog so we can track the merger timescale of the pair.
+However, we only want to keep a single instance of any given orbit in our catalog, since the orbits with the longest timescales will be selected many times which would skew our data to longer merger timescales.
 
@@
 \section{Pair Sample Properties}\label{sec:pairprops}
-\begin{figure}[htb]
+\begin{figure}[tb]
     \begin{center}
     \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
-    \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
-    The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
-    We only count the number of orbits ``at" a given redshift if the pairs have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
-    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
+    \caption{The number of unique low-mass (green) and high-mass (pink) pairs in our orbit catalog as a function of redshift.
+Orbits are only defined between first infall and the merger redshift (for merging pairs) or $z=0$ (for non-merging pairs).
+    The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers), while dashed lines show the number of pairs at a given redshift that will merge before $z=0$.
+    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.\kc{"total"}
     }
@@
     %At $z>1$, the fraction of selected pairs that merge before $z=0$ is upwards of 80\% for both low mass and high mass isolated pairs. 
-    The sharp decline of the merger fraction to 0 at low redshift is a non-physical feature of the simulation ending at $z=0$.
+    The sharp decline of the merger fraction to zero at low redshift is a non-physical feature of the simulation ending at $z=0$.
     Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
@@
 \subsection{Number of Pairs}\label{sec:pairprops-num}
-We calculate the number of pairs by selecting all pairs at a given redshift that have not merged.
+We calculate the total number of pairs by selecting all pairs at a given redshift, including merger and non-merger pairs.
 A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.
@@
 Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass pairs as a function of redshift from $z=0-6$. 
-Low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $z=0-1$.
+Low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while high-mass pairs (pink solid line) are most numerous between $z=0-1$.
 The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
@@
 
-As more subhalos experience first infall into a FoF, the number of pairs increases with decreasing redshift. 
-On the other hand, as mergers occur, the number of pairs decreases with decreasing redshift, as pairs that have merged are no longer members of the pairs sample.
+As subhalos experience first infall into a FoF group, the number of pairs increases. 
+On the other hand, mergers cause a decrease in the number of pairs.
 % The difference between high and low mass
-The number of low-mass pairs continues to increase from $z=6$ until around $z=2$, at which point the number of added pairs at a given redshift slows. 
+The number of low-mass pairs increases from $\sim600$ pairs at $z=6$ to $\sim6,600$ at $z=2$. 
 The decrease in the number of pairs after $z=2$ means that the number of low-mass pairs that are merging at each redshift is larger than the rate at which pairs are being added to the sample.
 
-The number of high-mass pairs continues increasing until $z=1$, at which point it remains approximately constant from $z=0-1$ at $\sim1000$ pairs. 
+The number of high-mass pairs also increases until $z=1$, at which point it remains approximately constant from $z=0-1$ at $\sim1000$ pairs. 
 There is a slight decrease in the number of high mass pairs at the very lowest redshifts $z<0.1$, where the mergers begin to outnumber the new pairs being added at each redshift.
@@
 \subsection{Merger Fraction}\label{sec:pairprops-frac}
-We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift, including all pairs that do and do not merge.
+    We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the total number of merging and non-merging orbits (solid lines) at a given redshift.
     As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
@@
     At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
-The merger fraction for both mass ranges declines to zero at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
+    The merger fraction for both mass ranges declines to zero at $z=0$, due to the very small fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.
 
     % difference/similiarity between low and high
-The merger fraction of low-mass and high-mass pairs is the same from $z=0-0.5$ and $z>2.5$. 
-Between $z=0.5-2.5$, the high-mass merger fraction is higher than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift. 
+    The merger fraction of low-mass and high-mass pairs is the same from $z=0-0.5$ and $z=2.5-6$. 
+    Between $z=0.5-2.5$, the high-mass merger fraction is larger than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift. 
     We discuss this difference in more detail in Sec.~\ref{sec:disc-nonmergers}.
@@
 
-\begin{figure*}[htb]
+\begin{figure*}[htb!]
     \begin{center}
-    \includegraphics[width=0.75\textwidth]{plots/bet-on-it/8_2Dhist.png}
-    \caption{The distribution of times until merger for low-mass (top) and high-mass (bottom) pairs as a function of physical separation at $z=0.5,1,1.5,2,3,4,5\mbox{and }6$. 
-    There is a positive correlation between separation at the selected redshift and the time until merger. 
-    The lowest separation pairs tend to have the shortest merger timescales.
-    Of low-mass pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$, while high-mass pairs have separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$.
-    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
+    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/8_2Dhist.png}
+    \caption{The distribution of merger timescales for low-mass (top) and high-mass (bottom) pairs as a function of physical separation at $z=0.5,1,1.5,2,3,4,5\mbox{and }6$. 
+    The colorbars show the percentage of the total number of pairs at each redshift (given in the bottom right corner of each panel) that are in a given bin.
+    In each of the first panels, the grey horizontal dashed line shows the time remaining in the simulation, above which there are no mergers. 
+    %
+    We find:
+    1.) a positive correlation between separation and the time until merger at each selected redshift -- the lowest separation pairs tend to have the shortest merger timescales; 
+    2.) given a redshift and separation, the low-mass merger timescale is longer than the high-mass merger timescale, i.e.
+low-mass pairs at $z=1$ with separations $\rsep<100\,\kpc$ merge within $6\, \Gyr$, while high-mass pairs in the same separation range merge within $3\, \Gyr$;
+    3.) merger timescales are longer for lower redshift pairs; 
+    4.) the spread of merger timescales and separations increase with decreasing redshift.
+    %
     This plot can be used to get estimates for the merger timescale of an isolated pair at a given redshift. 
-    For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
-    On the other hand, a high-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-2\,\Gyr$, with a most likely time to merger of around $0-1\Gyr$.
+    For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-5\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
+    On the other hand, a high-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-2\,\Gyr$, with a most likely merger timescale of around $0-1\Gyr$.
     }
@@
     Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
-    The merger timescale of a pair is defined to be the amount of time that elapses in the simulation between the redshift at which pairs are selected and the merger redshift.\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}\textsuperscript{,}\footnote{We choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing.
-However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.}
+    The merger timescale of a pair is defined to be the amount of time that elapses between the redshift at which pairs are selected and the merger redshift.
     Only the merging pairs will be considered for the remainder of the analysis.
+\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
+    Note that we choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing.
+However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.
 
-    In Sec.~\ref{sec:results-timevsep}, we explore how the merger timescale changes as a function of the separation of the pair across redshifts, and how the merger timescales differ between the low-mass and high-mass pairs.
-    In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all isolated pairs as a function of redshift. 
+    In Sec.~\ref{sec:results-timevsep}, we explore how the merger timescale changes as a function of pair separation across redshifts for low-mass and high-mass pairs. 
+    In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all pairs as a function of redshift. 
     We will additionally examine the merger timescale's dependence on a variety of separation criteria. 
@@
 separation for low-mass (top) and high-mass (bottom) pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$.
-    The colorbars to the right of each figure show the percentage of the population of pairs at that redshift that are in each bin, and the number of pairs at each redshift is shown by the number in the bottom right corner of each panel.
+    The colorbars to the right of each figure show the percentage of the population of pairs at that redshift that are in each bin.
+    The number of pairs at each redshift is printed in the bottom right corner of each panel.
     The horizontal line in the first panel shows the time remaining in the simulation until $z=0$, above which no merging pairs exist.
@@
     We find that the time until merger for low-mass and high-mass pairs is positively correlated with the separation of the pair at each redshift. 
+    Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$. 
+    For the same physical separation, the merger timescale of high-mass pairs is shorter than for low-mass at each redshift. 
     % However, the slope of this correlation decreases at higher redshifts.
-    For the same physical separation, the merger timescale of high-mass pairs is shorter than for low-mass. 
-    Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.
 
-    In addition, the spread of times until merger and separations is smaller at higher redshift, for both low- and high-mass pairs. 
-    The spread of merger timescales for low-mass pairs at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
+    In addition, the spreads in the merger timescales and pair separations are smaller at higher redshift, for both low- and high-mass pairs. 
+    For example, the spread of merger timescales for low-mass pairs at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
     Likewise, the spread of separations at $z=1$ is $0-200\kpc$, but at $z=4$ the spread is smaller, between $10-125\kpc$. 
@@
     % 
-    \input{table-septime.tex}
-    A compilation of more specific data is available in Table~\ref{tab:timevssepvsz}.
-    Here, we list the fraction of pairs within three separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction of pairs within each of those separation bins that merge in four different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$.
-    We present these data for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$.
+    % \input{table-septime.tex}
+    % A compilation of more specific data is available in Table~\ref{tab:timevssepvsz}.
+    % Here, we list the fraction of pairs within three separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction of pairs within each of those separation bins that merge in four different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$.
+    % We present these data for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$.
 
-    Data from the table can be interpreted in the following way: at $z=1.5$, there are 5,787 pairs that merge prior to $z=0$. 
-    Of those, 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 (38.1\%) of those $10-50\,\kpc$ pairs will merge in the next $2\,\Gyr$.
+    % Data from the table can be interpreted in the following way: at $z=1.5$, there are 5,787 pairs that merge prior to $z=0$. 
+    % Of those, 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 (38.1\%) of those $10-50\,\kpc$ pairs will merge in the next $2\,\Gyr$.
 
-    \begin{figure}[htb]
+\subsection{Redshift Dependence of Merger Timescales}\label{sec:results-timevredshift}
+    % subsections:
+    % 1 - the full sample 
+    % 2 - physical separation cuts
+    % 3 - scaled separation cuts
+    In this subsection, we study the merger timescales of low-mass and high-mass pairs as a function of redshift.
+
+    First, we consider the pair sample as a whole, and quantify the median merger timescale for all merging low- and high-mass pairs from $z=0-6$. 
+    We then consider two sets of separation criteria to create separation-selected subsamples, as in \chambe{}, and study the impact of different selection criteria on the resulting merger timescales.
+
+    \subsubsection{Full sample}
+        We calculate the merger timescales for pairs at each redshift in the simulation from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
+        We include the full catalog of merging orbits at each redshift, only excluding separations $\rsep<10\kpc$ to limit the impact of subhalos becoming indistinguishable in the \subfind{} catalogs.% to calculate the merger timescale.
+        % why do we want to consider the entire population of halos? why interesting?
+
+        \begin{figure}[tb]
             \centering
             \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale_mod.png}
-        \caption{The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs that will merge by $z=0$.
+            \caption{The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs in our merger sample. 
             Shaded regions represent the 1st and 3rd quartile spread on the median. 
-%Pairs at each redshift are selected if they have experienced first infall and are pre-merger.
             % 
-        The black dashed line shows the fraction of the Hubble time as a function of redshift, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
-        The dotted black line shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
-        The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.\kc{only one of the dashed lines will remain in this plot once I discuss w Paul; likely remove age of sim}}
+            The thick dotted black line at low redshifts shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the merger timescale that a merging pair can have.
+            The thin black line shows the Hubble time as a function of redshift multiplied by a constant $XX$, which approximates the median time until merger from $z=6$ to $z\sim1$.
+            %
+            The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.\kc{update}}
             \label{fig:timescales}
@@
 
-    \begin{figure*}[htb]
-        \centering
-        \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger.png}
-        \caption{(Top) The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $(50,70,100,150,200,\mbox{and }300)\,\kpc$ from left to right.
-        (Bottom) The median time until merger as a function of redshift for pairs with physical separations greater than $10\,\kpc$ and less than $(0.25, 0.5, 1, 1.5, 2,\mbox{and }2.5)\,\Rvir$ from left to right.
-        % 
-        The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$.
-        \kc{add h(z) line}
-        }
-        \label{fig:timescales-sep}
-    \end{figure*}
-
-\subsection{Redshift Dependence of Merger Timescales}\label{sec:results-timevredshift}
-    % The behavior of time until merger as a function of redshift for low-mass and high-mass pairs:
-    % 1 - the full sample 
-    % 2 - the physical and static kpc separation-selected sample
-    % 3 - the mass and redshift evolving scaled separation-selected sample
-    In this subsection, we study the merger timescale of low-mass and high-mass pairs as a function of redshift.
-
-    First, we consider the pairs sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that will merge for $z=0-6$. 
-    We will then consider two different sets of separation criteria to create separation-selected subsamples.
-    Since pair samples are typically picked via separation criteria, in both simulations and observations, we study the impact of different separation-based selection criteria on inferred merger timescales.
-
-    \subsubsection{Full sample}
-        We calculate the merger timescales for pairs at each redshift in the simulation from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
-        We include the full catalog of merging orbits at each redshift.% to calculate the merger timescale. 
-        % why do we want to consider the entire population of halos? why interesting?
-
-        Figure~\ref{fig:timescales} presents the median of the time until merger for post-infall merging pairs from $z=0-6$. 
+        Figure~\ref{fig:timescales} presents the median merger timescales for merging pairs from $z=0-6$. 
         The 1st and 3rd quartiles are indicated with shaded regions. 
@@
         % redshift evolution
-        The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.
-        The abrupt decrease is artificial rather than a true physical feature. 
+        The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to zero at $z=0$.
+        The abrupt decrease is artificial rather than a true physical feature of merger timescales. 
         As $z\to0$, there is an increasingly small fraction of pairs that merge before $z=0$, as seen by the dashed lines in Fig~\ref{fig:numorbits}. 
         % These mergers must then proceed on timescales less than the elapsed time difference between a given redshift and $z=0$. 
-        These mergers must also proceed on shorter timescales by definition, as they were selected as pairs that must merge prior to $z=0$ (i.e., a pair at $z=0.1$ that merges by $z=0$ can have a maximal merger timescale of $1.47\Gyr$, whereas a pair at $z=1$ can have a maximal merger timescale of $8\Gyr$).
+        These mergers must proceed on shorter timescales by definition, as they were selected as pairs that must merge prior to $z=0$ (i.e., a pair at $z=0.1$ that merges by $z=0$ can have a maximal merger timescale of $1.47\Gyr$, whereas a pair at $z=1$ can have a maximal merger timescale of $8\Gyr$).
         This is shown by the dotted black line on the left of Fig.~\ref{fig:timescales}, which shows the lookback time of the simulation as a function of redshift.\footnote{The lookback time at a given redshift $z_n$ is equivalent to the time elapsed between $z_n$ and $z=0$.
@@
 
-        In principle, merger timescales\footnote{Specifically, the elapsed time between crossing into the virial radius of the primary subhalo and coalescence with the primary.} for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}.
+        To understand the functional form of merger timescales between $z=1-6$, which is shown by the thin black line in Fig.~\ref{fig:timescales}, we consider the following derivation. 
+        In principle, merger timescales for a body moving through a homogeneous field of collision-less matter can be analytically derived from the Chandrasekhar formula for dynamical friction~\citep{Binney2008}.\footnote{Specifically, this derivation considers the merger timescale to be the time that elapses between the secondary subhalo crossing into the virial radius of the primary subhalo and the secondary's coalescence with the primary.}
         Departures from the idealized case introduce perturbations to that solution, the validity of which has been tested in cosmological hydrodynamic and N-body simulations~\citep{Jiang2008, BoylanKolchin2008}.
@@
         where $V_{\rm circ}(\Rvir)$ is the circular velocity at the virial radius of the primary subhalo. 
-        Note that the dynamical time can be rewritten as 
+        Note that the dynamical time can thus be rewritten as 
         \begin{equation}
@@
 
-        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by 0.25. 
-        We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the overall scaling.
-        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=1.5-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation.
+        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by XXX. 
+        We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the specific values of the merger timescale.
+        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=\sim1-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation, from~\citet{Planck2018}.
 Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
 
-        It is notable that such a simple scaling holds for both low-mass and high-mass pairs, especially in a complex simulation like TNG100 where subhalos are far from isotropic spherical distributions of mass, and that low-mass and high-mass pairs have the same merger timescales.
+        It is notable that such a simple scaling holds for a complex simulation like TNG100 where subhalos are far from isotropic spherical distributions of mass.
+        It is particularly interesting as well that low-mass and high-mass pairs have the same merger timescales \textit{and} follow the $1/H(z)$ redshift scaling. 
         We believe this is due to the dominance of dynamical friction in determining the redshift behavior of merger timescales. 
@@
     \label{sec:results-phys}
-        Both separation selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \subfind{} catalogs.
+        Since pair samples are typically picked via separation criteria in both simulations and observations, we study the impact of different separation-based selection criteria on inferred merger timescales.
+        Separation selected samples have a lower separation criteria of $10\,\kpc$.
         This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates~\citep[see e.g.][and observational studies cited therein]{Lotz2011,Snyder2017,Besla2018,Chamberlain2024}.
 
-        The first set of separation criteria will select only pairs at a given redshift that have physical 3D separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, dividing the sample into six pair subsamples.
+        The first set of separation criteria will select only pairs at a given redshift that have physical 3D separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, yielding six pair subsamples.
+        Pairs that are selected via the $\rsep<50\kpc$ separation cut will be included in the $\rsep<70\kpc$ subsample, and so on.
         These separation criteria do not vary as a function of time or mass, and are applied equivalently to the low-mass and high-mass samples at all redshifts. 
-        For example, an orbit in the separation bin $10-50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts. 
-        Rather, the separation criteria are applied independently at each redshift.
+        The separation criteria are applied independently at each redshift.
+        For example, an orbit in the separation bin $<50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts.
 
         % figure description 
-        The top panel of Fig.~\ref{fig:timescales-sep} shows the time until merger versus redshift for each of the low-mass (green) and high-mass (pink) pair subsamples with separations less than the physical separation listed above each column.
+        The top panel of Fig.~\ref{fig:timescales-sep} shows the time until merger versus redshift for each of the low-mass (green) and high-mass (pink) pair subsamples with separations less than the physical separation listed above each panel.
+        The same Hubble time redshift scaling from Sec.~\ref{sec:results-timevredshift} is shown by the thin black line in each panel.
         % redshift evolution of each subsample
@@
         In addition, all subsamples show the same decline of mean merger time as $z\to0$, as discussed in the previous section. 
-        These traits were also present in Fig.~\ref{fig:timescales}, so they are features that are roughly independent of a separation criteria.
-        As the selection criteria gets larger, the subsample contains more of the full sample, eventually including the full sample of merging pairs at large enough separations.
+        These traits were also present in Fig.~\ref{fig:timescales}, meaning they are features that are roughly independent of separation criteria.
+
+        %, eventually including the full sample of merging pairs at large enough separations.
 
@@
         The subsamples with maximum separations of $[50, 70, \mbox{and } 100]\,\kpc$ result in median merger timescales that are higher for low-mass pairs than high-mass pairs at all redshifts. 
-        The difference in the median time until merger is up to $0.8\,\Gyr$ longer for low-mass pairs than high-mass pairs at the same redshift for the same separation criteria. 
-        The offset between the low-mass and high-mass merger timescale decreases for each successive separation cut, which will each include pairs with larger separation. 
-        In the rightmost top panel, the median merger times converge for $z=0-0.5$ and $z>1.5$, as expected based on results from the full sample shown in Fig.
-\ref{fig:timescales}.
-%almost mimicking our findings from the full sample analysis.
+        The difference in the merger timescale is up to $0.8\,\Gyr$ longer for low-mass pairs than high-mass pairs at the same redshift for the same separation cut. 
+        The offset between the low-mass and high-mass merger timescale decreases for the largest separation cut of $\rsep<300\kpc$. 
+        In the top rightmost panel, the median merger times converge for early all redshifts, similar to the merger timescales from the full sample shown in Fig. \ref{fig:timescales}.
 
         % impact of the separation criteria
-        When the maximum separation is increased (from left to right), more of the full sample is used in the calculation for the median merger timescale. 
-        Indeed, the merger timescale for both low-mass and high-mass pairs increases with an increasing maximum separation cut.  
+        As the selection criteria increases (from left to right), each subsample contains a larger fraction of the full sample.
+        % When the maximum separation is increased (from left to right), more of the full sample is used in the calculation for the median merger timescale. 
+        The merger timescale for both low-mass and high-mass pairs increases with an increasing maximum separation cut.  
         This follows directly from Sec.~\ref{sec:results-timevsep}, where we found that the time until merger is positively correlated with increasing separation for all pairs. 
-        Thus, including a higher fraction of larger-separation systems in our merger time calculation tends to increase the median time until merger.
+        Including a higher fraction of larger-separation systems in each subsample increases the median time until merger at all redshifts.
 
-        The median merger timescale for low-mass pairs does not change significantly for any separation criteria $>150\,\kpc$. 
-        On the other hand, high-mass pairs tend to see an increase in the merger timescale for each larger separation cut. 
-        This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
+        The median merger timescale for low-mass pairs does not change significantly for any separation cuts above $150\,\kpc$. 
+        On the other hand, high-mass merger timescale tends to increase for each larger separation cut. 
+        % This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
         High-mass pairs tend to have higher separations than low-mass pairs, as shown in~\ref{sec:results-timevsep}. 
-
+    \begin{figure*}[htb]
+        \centering
+        \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger.png}
+        \caption{(Top) The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $50,70,100,150,200,\mbox{and }300\,\kpc$ from left to right.
+        (Bottom) The median time until merger as a function of redshift for pairs with physical separations greater than $10\,\kpc$ and less than $0.25, 0.5, 1, 1.5, 2,\mbox{and }2.5\,\Rvir$ from left to right.
+        % 
+        The thin black line is the same as shown in Fig.~\ref{fig:timescales}, and goes as the Hubble time $1/H(z)$.
+        The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$.
+        \kc{H(z), "merger timescale"}
+        }
+        \label{fig:timescales-sep}
+    \end{figure*}  
     \subsubsection{Scaled Separation Selected Pairs}
     \label{sec:results-scal}
-        We calculate the merger timescales for six additional subsamples of pairs, this time applying separation criteria that scale with both redshift and mass.
+        We calculate the merger timescales for six additional subsamples of pairs, applying separation criteria that scale with both redshift and mass.
         The scaled separation, which we define in Section~\ref{sec:methods}, is the separation of a pair divided by the virial radius of the pair's FoF group, $\Rvir$. 
-        The scaled separation selected pairs have separations greater than $10\,\kpc$ and less than $<[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$. 
+        Scaled separation criteria select the equivalent \textit{fraction} of the full volume of each FoF group, regardless of mass or redshift, whereas physical separation criteria will select the same volume around each primary, with no account for the growth of dark matter halos or spatial distributions of satellites.
+        The scaled separation selected pairs have separations greater than $10\,\kpc$ and less than $[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$. 
         As in the previous section, the selection criteria are applied at each redshift independently.
@@
         As detailed in \chambe{}, the median virial radius for low-mass FoF groups at $z=[0,1,2,3,4]$ is approximately $[134, 85, 59, 43, 33]\,\kpc$, and for high-mass FoF groups is approximately $[348, 206, 134, 97, 76]\,\kpc$.
-        Thus, choosing a scaled separation cut of $<1\,\Rvir$ at $z=1$ will select high-mass pairs with separations $\lesssim205\,\kpc$ and low-mass pairs with separations $\lesssim85\,\kpc$.
+        Thus, choosing a scaled separation cut of $\rsep<1\,\Rvir$ at $z=1$ will select low-mass pairs with separations $\rsep\lesssim85\,\kpc$ and high-mass pairs with separations $\rsep\lesssim206\,\kpc$.
 
@@
         The bottom panel of Fig.~\ref{fig:timescales-sep} shows the median time until merger versus redshift for low-mass (green) and high-mass (pink) pair subsamples with scaled separations less than those listed at the top of each plot.
-        We find that the median time until merger at $z=1$ is $[0.64, 0.94,1.25,1.51-1.58,1.76-1.91,1.91-2.05]\,\Gyr$ for each scaled separation cut respectively, for both low-mass and high-mass samples.
-%\kc{gonna have to triple check this cause I get the SAME median time til merger for low and high mass for Rvir 0.25, 0.5, and 1, down to like 14 decimals...}
+        As in the top row, the thin black line in each panel shows the Hubble time scaling derived in Sec.~\ref{sec:results-timevredshift}.
+
+        Quite unlike the physical separation selected subsamples in the top panels, using a scaled separation criteria results in nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts.
+        We find that the median time until merger at $z=1$ for each scaled separation cut is about $\sim[0.64, 0.94,1.25,1.55,1.85,1.96]\,\Gyr$  respectively, for both low-mass and high-mass samples.
+\kc{triple check these nums}
 
         % redshift evolution of each subsample
-        Similar to the static physical separation-selected subsamples in the top panel, the median time until merger starts off small at $z>4$, then increases to a peak around $z\sim0.5$, before quickly decreasing to 0 at $z=0$.
-        Additionally, the median time until merger at a given redshift increases for separation criteria that include a larger fraction of the virial radius, and thus larger separation pairs.
+        The median merger timescale starts off small at $z>4$, then increases to a peak around $z\sim0.6$, before quickly decreasing to zero at $z=0$. 
+        This redshift evolution is similar to that of the full sample in Fig.~\ref{fig:timescales}.
 
         % low mass vs high mass
-        However, quite unlike the physical separation selected subsamples in the top panels, using a scaled separation criteria results in nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts.
-        The low-mass and high-mass merger timescales increase with each larger scaled separation cut, but recover the behavior of the full sample by $<2\,\Rvir$.
+        The median time until merger at a given redshift increases .
+        The low-mass and high-mass merger timescales increase for separation criteria that include a larger fraction of the virial radius, and thus larger separation pairs, but recover the behavior of the full sample by a separation cut of $2\,\Rvir$.
 
-        In addition, the slope of the time until merger as a function of redshift for $z>0.5$ is steeper for high-mass pairs in the scaled separation subsamples than the physical separation subsamples.
-        This is especially noticeable in the first panel of both the top and bottom rows, where the merger timescale of high-mass pairs is approximately flat for $<50\,\kpc$.
+        In addition, the slope of the merger timescale from $z=6\to1$ is steeper for high-mass pairs in the scaled separation subsamples than the physical separation subsamples.
+        This is especially noticeable in the first panel of the top and bottom rows, where the merger timescale of high-mass pairs is approximately flat for the physical separation cut $\rsep<50\,\kpc$.
 
@@
 \section{Discussion} \label{sec:discussion}
-    We have calculated the merger timescales, or time until merger, for a sample of physically co-located low-mass and high-mass pairs from $z=0-6$.
-    In this section, we will explore the implications and broader impacts of our findings, and put our study in context with previous work.
+    %We have calculated the merger timescales, or time until merger, for a sample of physically co-located low-mass and high-mass pairs from $z=0-6$.
+    In this section, we will explore the implications and broader impacts of our merger timescale analysis, and put our study in context with previous work.
 
-    In Sec.~\ref{sec:disc-nonmergers}, we will discuss the non-merger portion of our orbit sample and the difference between low- and high-mass merger fractions.
-    In Sec.~\ref{sec:disc-pairfrac}, we return to one of the key questions posed in our previous work: what is the underlying physical mechanisms that drives the difference in pair fraction evolution between low- and high-mass pairs.
-    Finally, in Sec.~\ref{sec:disc-suggestions}, we discuss the implications of our work for future observational and theoretical merger fraction and merger rate studies, and provide suggestions for a self-consistent framework moving forward.
+    In Sec.~\ref{sec:disc-pairfrac}, we return to one of the key questions posed in our previous work: what are the underlying physical mechanisms that drive the difference in pair fraction evolution between low- and high-mass pairs?
+    In Sec.~\ref{sec:disc-nonmergers}, we will discuss the non-merger portion of our orbit sample and the difference between the merger fractions in the low-mass and high-mass regimes.
+    Finally, in Sec.~\ref{sec:disc-suggestions}, we discuss the implications of our work for future observational and theoretical merger fraction and merger rate studies and provide suggestions for a self-consistent framework moving forward.
+
+    \subsection{Connections to Pair Fractions}\label{sec:disc-pairfrac}
+        In our previous work~\citep{Chamberlain2024}, we found that the pair fractions of isolated low-mass and high-mass pairs evolve very differently from $z=0-2.5$ (see their Fig.~3). 
+        Low-mass major pair fractions are constant for $z>2.2$, but decrease by almost 70\% from $z=3$ to $z=0$. 
+        High-mass major pair fractions slowly increase from $z=6$ to $z=0$, with a more significant increase from $z=0.25$ to $z=0$.
+        In short, the redshift evolution of pair fractions is opposite for low-mass pairs (which have a decreasing pair fraction with decreasing redshift) and high-mass pairs (which have an increasing pair fraction with decreasing redshift) from $z=3\to0$.
+
+        \chambe{} suggested a hypothesis for the difference in the behavior of the pair fractions:  that perhaps low- mass and high-mass pairs have different merger timescales, which could lead to a difference in pair fractions from high to low redshift.
+
+        However in the present study, we have shown that the merger timescales of the full sample of low-mass and high-mass pairs are approximately equal at all redshifts, meaning that the opposing behaviour of low-mass and high-mass pair fractions is not a result of different merger timescales.
+        Rather, we suspect that the build-up of larger structures due to hierarchical formation under $\Lambda$CDM leads to a reduced number of low-mass groups that enter the analysis at lower redshift.
+        Additionally through this study, we have found that a significant fraction of the pairs selected in our previous work do indeed merge ($>90\%$ of pairs for $z>1.6$), so that the rate at which mergers occur is likely larger than the rate at which isolated low-mass pairs form, particularly at low redshift, thus leading to the quickly declining pair fraction measured in our previous work.
 
     \subsection{Non-merger population}\label{sec:disc-nonmergers}
-        In our catalog of 22,213 low-mass pairs and 3,029 high-mass pairs, only 16.66\% (3700) of low-mass pairs and 29.71\% (903) of high-mass pairs do not merge by $z=0$.
-        In Sec.~\ref{sec:pairprops-frac}, we found that the merger fraction of low-mass and high-mass pairs at $z>3$ is $>0.95$. 
-        The merger fraction decreases substantially as you move to lower redshifts. 
+        In our catalog of 22,213 low-mass pairs and 3,029 high-mass pairs, only 3700 (16.66\%) low-mass pairs and 903 (29.71\%) high-mass pairs do not merge by $z=0$ (i.e.
+are classified here as ``non-mergers").
+        In Sec.~\ref{sec:pairprops-frac}, we found that the merger fraction of low-mass and high-mass pairs at $z>3$ is $>0.95$ and the merger fraction decreases substantially at $z<3$. 
         While the decline in the merger fraction at $z\lesssim1$ is primarily due to the limited time that a pair will have to merge before the end of the simulation, we also find that between $z=\sim0.6-2.5$ the fraction of low-mass pairs is lower than that of high-mass pairs.
 
-        One potential explanation for this behavior could be that low-mass pairs have shorter merger timescales than high-mass pairs, and thus merge at earlier times in the simulations, thus leading to a decrease in the merger fraction at higher redshifts.
-        However, through our analysis, we have shown that the merger timescales of our full low-mass and high-mass pair sample evolve in the same way from $z=0-6$. 
-        This means that the difference in the merger fraction between $z=0.5-2.5$ is not a result of different merger timescales.
+        As in the previous subsection, one possible hypothesis that could explain the merger fraction differences is that low-mass pairs have shorter merger timescales than high-mass pairs, such that their mergers would occur at earlier times in the simulation leading to a decrease in the merger fraction at higher redshifts compared to high-mass pairs.
+        However, we have shown that the merger timescales of our full low-mass and high-mass pair sample evolve similarly for $z=0-6$ which means that the difference in the low- and high-mass merger fraction between $z=0.5-2.5$ is not a result of a difference in merger timescales.
 
         We suspect that the cause of the difference between low- and high-mass merger fractions is not a difference in the merging population itself, but rather in the non-merger population and its evolution over time.
-        As we did not distinguish our non-merging sample into ``likely mergers" and ``fly-by" interactions for this study, we cannot say anything about how these two populations individually contribute to the difference in the merger fraction.
+        As we do not categorize our non-merging sample into ``likely mergers" and ``fly-by interactions" subsamples for this study (see Fig.
+\ref{fig:example-orbits}), we cannot comment on how these two populations individually contribute to the difference in merger fractions.
         However, one possible explanation could be that a larger fraction of infalling low-mass pairs at lower redshift results in fly-by interactions. 
         This could be the case if the velocity at which the secondary subhalo enters the primary's FoF group becomes increasingly large compared to the virial velocity\footnote{The virial velocity is the value of the circular velocity at the virial radius.}, resulting in fewer bound systems.
-        % We leave this as a focus of future work.
+        This will be studied in future work.
 
@@
 
-    \subsection{Connections to Pair Fractions}
-        In our previous work~\citep{Chamberlain2024}, we found that the pair fractions of isolated low-mass and high-mass pairs evolve very differently from $z=0-2.5$. 
-        Low-mass major pair fractions are constant for $z>2.2$, but decrease by almost 70\% from $z=3$ to $z=0$ (see their Fig.~3). 
-        High-mass major pair fractions slowly increase from $z=6$ to $z=0$, with a more significant increase from $z=0.25$ to $z=0$.
-
-        One of the hypotheses we laid out in~\chambe{} for the difference in the behavior of the pair fractions is that perhaps low- mass and high-mass pairs have different merger timescales, which could lead to a difference in pair fractions from high to low redshift.
-
-        However, in this study, we have shown that the merger timescales of low-mass and high-mass pairs are approximately equal at all redshifts, particularly when considering the full pair sample or the scaled separation selected sample.
-        This means that different merger timescales are not responsible for the difference in the evolution of pair fractions. 
-        Rather, we suspect that the build-up of larger structures due to hierarchical formation under $\Lambda$CDM leads to a reduction in the total number of low-mass groups at lower redshift $z\lesssim2$.
-
-        Additionally through this study, we have found that a very large fraction of the pairs selected in our previous work do indeed merge ($>90\%$ of pairs for $z>1.6$), so that the rate at which mergers occur is likely larger than the rate at which isolated low-mass pairs form, particularly at low redshift, thus leading to a quickly declining pair fraction measured in our previous work.
-
     \subsection{Implications and Suggestions for Future Observational and Theoretical Studies of Pairs}~\label{sec:disc-suggestions}
-        While we are unable to make meaningful comparisons with previous observational studies of merger fractions and merger rates, as our study is completed using true physical separations from the simulation rather than projected separations, we can still draw meaningful conclusions about the application of our results to future studies of merger timescales and merger rates.
+        While we are unable to make direct comparisons with previous observational studies of merger fractions and merger rates, as our study uses true physical separations from the simulation rather than projected separations, we can still draw meaningful conclusions about the application of our results to future studies of merger timescales and merger rates.
 
         \subsubsection{Pair Selection Criteria and Merger Rates in Illustris}
-            Comparisons between studies of pair fractions, merger fractions, and merger rates, especially across cosmic time and different mass scales, are often challenged by the use of different sets of selection criteria in each study.
+            Comparisons between studies of pair fractions, merger fractions, and merger rates, especially across cosmic time and different mass scales, are often challenged by the implementation of different selection criteria in each study.
 
-            On the observational side, pair selection criteria are often set by the observational parameters of the survey itself. 
-            For example, limiting magnitudes and completeness limits can enforce a specific new range of stellar mass ratios to be considered (for example $1-1:6$ instead of $1:10$), while avoiding fiber collisions on a spectroscopic survey might lead to very specific separation criteria.
-            On the theoretical side, it can be more straightforward to adopt a number of different pair selection criteria for comparisons to specific observational studies with identical criteria. 
-            However, this means that theoretical studies typically restrict themselves to a subset of pair selection criteria specific to their target comparison studies.
+            In some observational studies, pair selection criteria are often set by the observational parameters of the survey itself. 
+            For example, limiting magnitudes and completeness limits can dictate which range of stellar masses and stellar mass ratios are considered~\citep{Patton2000,Lotz2008,Man2016,Ventou2019}.
+Additionally, specific separation criteria may be chosen to avoid fiber collisions in a spectroscopic survey~\citep[as in][]{Patton2011,Besla2018}.
 
-            In recent years, there have been works that aim to standardize pair selection criteria for future studies, in particular to facilitate a fair comparison between theory and observations. 
+            In theoretical studies, it is more straightforward to adopt several different pair selection criteria simultaneously for comparisons to specific observational studies~\citep[i.e.][]{Lotz2011,RG2015,Besla2018,Snyder2023}.
+            However, this then limits theoretical studies to a subset of pair selection criteria specific to their target comparison studies.
+
+            In recent years, some work has aimed to standardize pair selection criteria for future studies, in particular, to facilitate a fair comparison between theory and observations. 
             One such study is~\citet{Ventou2019}, which created mock catalogs from the original Illustris-1 simulation to analyze the merger probability of galaxy pairs as a function of (both physical and projected) separation and relative velocity.
-            They find that pairs with projected separations between $5\leq r_{proj} \leq 50\,kpc$ and projected relative velocities of $v_{proj}\leq300\,\kms$ selects all pairs that have greater than 30\% change of merging before $z=0$, and provide a weighting scheme to determine the likelihood of merger as a function of projected separation and velocity.
-            These selection criteria are thus calibrated to select a population of pairs that are fairly likely to merge, and a way to calibrate merger fractions more accurately from close pair fractions.
-            However, \kc{I think I need to say something here about what the shortcoming of this framework in the above paper is (if there is one) }.
+            They find that selecting pairs with projected separations between $5\leq r_{\rm proj} \leq 50\,\kpc$ and projected relative velocities of $v_{\rm proj}\leq300\,\kms$ selects all pairs that have greater than 30\% chance of merging before $z=0$.
+            These selection criteria are calibrated to select a population of pairs that are fairly likely to merge.
+            Using their projected separation vs velocity analysis of merger probabilities, they also develop a weighting scheme to determine the likelihood of merger as a function of projected separation and velocity, which provides a way to more accurately calibrate merger fractions from close pair fractions.
+            While their study considers the impact of stellar mass and redshift on the probability of merger, their suggested pair selection criteria do not vary with mass or redshift, which we have shown is crucial for interpretting and comparing pair fractions and merger timescales.
 
@@
 
-            \citet{RG2015} constructed galaxy merger trees to calculate the galaxy merger rate as a function of the stellar mass of the descendent subhalo and redshift. 
-            They find that the galaxy merger rate increases with increasing redshift for descendent galaxy masses $\ms{}>10^{10}\Msun$. 
-            These findings are consistent with predictions of the galaxy merger rate from semi-empirical models~\citep{Stewart2009,Hopkins2010}, and with merger rates derived from observational merger fraction measurements calibrated by the corrected observability timescales of \citet{Lotz2011}.
-            They additionally find that the major merger rate at $z\sim0.1$ increases with increasing descendant subhalo mass.
-            While the redshift evolution of merger rates are broadly consistent with observations, they find significant discrepancy with major merger rates of galaxies with $\ms<10^{10}\Msun$ from~\citet{Casteels2014}, which used morphological merger signatures to estimate merger rates at low $z$.
-            Since the~\citet{RG2015} work is done self consistently at all redshifts and descendant masses, they suggest that the discrepancy arises from an extrapolation of gas fraction data available for $\ms<10^{10}\Msun$ to galaxies with lower mass, which will be reflected in observability timescales with significant uncertainty in the lower mass regime.
+            \citet{RG2015} constructed galaxy merger trees to calculate the galaxy merger rate as a function of the stellar mass of the descendent subhalo and redshift in the original Illustris simulation.
+            They find that the galaxy merger rate increases with increasing redshift for descendent galaxy masses $\ms{}>10^{10}\,\Msun$. 
+            These findings are consistent with predictions of the galaxy merger rate from semi-empirical models~\citep{Stewart2009,Hopkins2010a}, and with merger rates derived from observational merger fraction measurements calibrated by the corrected observability timescales of \citet{Lotz2011}.
+            Additionally, \citet{RG2015} find that the major merger rate at $z\sim0.1$ increases with increasing descendant subhalo mass.
+            While the redshift evolution of merger rates are broadly consistent with observations, they find significant discrepancies with major merger rates of galaxies with $\ms\, <10^{10}\,\Msun$ from~\citet{Casteels2014}, which used morphological merger signatures to estimate merger rates at low $z$.
+            As the~\citet{RG2015} work is self-consistently carried out at all redshifts and descendant masses, \citet{RG2015} suggest that the discrepancy arises from significant uncertainty in the observability timescales of galaxies with lower mass $\ms{}<10^{10}\,\Msun$, since the timescales in \citet{Casteels2014} are calibrated using an extrapolation of gas fraction data that is only available for $\ms{}>10^{10}\,\Msun$.
 
             These results imply that the timescales used to convert observational merger fractions (determined either by weighted close pair fractions or by morphologically selected pairs) are sensitive to the mass ranges and redshifts at which they are valid, and attempting to make comparisons between theoretical predictions and observational measurements can lead to seemingly incompatible results.
-            Our work provides a unique framework that can allow for more robust comparisons of theoretical and observational samples in the future, since the observational samples themselves will be able to also self-consistently apply merger timescales to close pair fractions using our suggested scaled separation criteria.
+            % Since our observational samples themselves will be able to also self-consistently apply merger timescales to close pair fractions using our suggested scaled separation criteria.
+            Our work provides a unique framework that will allow for more robust comparisons of theoretical and observational samples.
 
         \subsubsection{Implementing Scaled Separation Pair Selection Criteria in Future Studies}
-            In our previous work, we studied the impact of different sets of separation criteria on the recovered pair fractions of low-mass and high-mass pairs, and found that only by utilizing scaled separation criteria can the features of the pair fractions' evolution with redshift be distinguished~\citep{Chamberlain2024}.
-            In this work, we explored how these same set of selection criteria impacts the recovered merger timescales.
+            In our previous work, we studied the impact of different sets of separation criteria on the recovered pair fractions of low-mass and high-mass pairs, and found that features of the redshift evolution of the pair fractions can only be distinguished by employing scaled separation criteria~\citep{Chamberlain2024}.
+            In the present work, we explore how these same set of selection criteria impact the recovered merger timescales. 
+            We find that the merger timescales of isolated low-mass and high-mass pairs evolve nearly identically between $z=0-6$.
+            All subsamples of our pair catalog from scaled separation cuts result in identical behavior of the low-mass and high-mass merger timescales as a function of redshift as well. 
+            On the other hand, selecting pairs using a static physical separation criteria results in merger timescales that evolve differently between the two mass scales, and can differ by $0.8\Gyr$.
 
-            In Sec.~\ref{sec:results-timevredshift}, we find that the merger timescales of isolated low-mass and high-mass pairs evolve nearly identically between $z=0-6$, rising from a median of $\sim0.8Gyr$ at $z=6$ to $2.1-2.4\Gyr$ at $z\sim0.6$.
-            From $z=0$ to $z=0.6$, the merger timescale rapidly declines, since the only mergers at low redshift that can proceed before the end of the simulation at $z=0$ are those with very short timescales.
+            % condensed and moved around
+            %, rising from a median of $\sim0.8\Gyr$ at $z=6$ to $2.1-2.4\Gyr$ at $z\sim0.6$.
+            %From $z=0$ to $z=0.6$, the merger timescale rapidly declines, since the only mergers at low redshift that can proceed before the end of the simulation at $z=0$ are those with very short timescales.
+            % We find that all subsamples of our pair catalog selected using the scaled separation $\rsep/\Rvir$, result in identical behavior of the low-mass and high-mass pair fractions as a function of redshift.
+            % On the other hand, selecting pairs using a static physical separation (such as requiring pairs to have separations between $10-50\kpc$, and using that selection criteria for low- and high-mass pairs at every redshift) results in merger timescales that evolve differently between the two mass scales.
+            % For example, the median merger timescale for close pairs selected with $\rsep<70\kpc$ varies by up to $\sim0.6\Gyr$ between low- and high-mass pairs at all considered redshifts, while the merger timescale for close pairs selected via the scaled criteria $\rsep<0.5\Rvir$ have nearly identical median timescales.
+            % This is because selection criteria based on scaled separations take the equivalent \textit{fraction} of the full volume of each FoF group, regardless of mass or redshift, whereas physical separation criteria will select the same volume around each primary, with no account for the growth of dark matter halos or spatial distributions of satellites.
 
-            We find that all subsamples of our pair catalog selected using a mass and redshift evolving separation criterion, namely the scaled separation given by $\rsep/\Rvir$, result in identical behavior of the low-mass and high-mass pair fractions as a function of redshift.
-            On the other hand, selecting pairs using a static physical separation (such as requiring pairs to have separations between $10-50\kpc$, and using that selection criteria for low- and high-mass pairs at every redshift) results in merger timescales that evolve differently between the two mass scales.
-            For example, the median merger timescale for close pairs selected with $\rsep<70\kpc$ varies by up to $\sim0.6\Gyr$ between low- and high-mass pairs at all considered redshifts, while the merger timescale for close pairs selected via the scaled criteria $\rsep<0.5\Rvir$ have nearly identical median timescales.
-            This is because selection criteria based on scaled separations take the equivalent \textit{fraction} of the full volume of each FoF group, regardless of mass or redshift, whereas physical separation criteria will select the same volume around each primary, with no account for the growth of dark matter halos or spatial distributions of satellites.
-
-            These results translate to important implications for observational studies of merger rates, which are typically determined using physical separation selected close pairs and averaged merger timescales.
-            Our work has shown that both of these quantities can vary significantly for different pair selection criteria, and that physical separation criteria can wipe out the distinguishing features of pair fractions or the equivalence of merger timescales between low-mass and high-mass pairs.
+            These results translate to important implications for observational studies of merger rates, which are typically determined using physical separation-selected close pairs and merger timescales.
+            Our work has shown that both of these quantities can vary significantly for different pair selection criteria and that physical separation criteria can eliminate the distinguishing features of pair fractions or the equivalence of merger timescales between low-mass and high-mass pairs.
 
@@
 pair fractions, merger timescales, merger rates, etc.), particularly when the goal of such a study is to quantify and compare the redshift evolution of pair properties at different mass scales.
-            Specifically, we suggest employing a separation criteria based on the scaled separation of the system.
 
-            Although this seems to introduce further complications to the pair selection process in observational studies, the use of a scaled separation as demonstrated here can be applied in a fairly straightforward way.
-            In particular, given an observation of a pair with stellar masses $\ms1$ and $\ms2$, we can use the SMHM relation to calculate the associated dark matter masses, $M_{h1}$ and $M_{h2}$, of the galaxy pair.
-            This introduces the same systematic error to an observational study as it does when assigning stellar masses to our dark matter halos that utilizes the abundance matching prescription of~\cite{Moster2013}, since they have accounted for observational stellar mass errors when constructing their SMHM relationship.
-            Indeed, the SMHM relationship can be sampled in a similar fashion as in~\chambe{} to get a spread on the error associated with the abundance matching process.
+            Specifically, we suggest employing separation criteria based on the scaled separation of each pair. 
+            The application of a scaled separation as demonstrated here can be straightforwardly applied to observational studies. 
+            Since estimates of a pair's stellar masses are needed to quantify the mass ratio of the pair (which is commonly computed for merger rate studies), no further information from the observations is needed to create the mass and redshift evolving separation cuts.
+            First, compute the associated dark matter masses from the observed stellar masses of the pair using a SMHM relation~\citep[in this study, we employ that of][]{Moster2013}.
+            % In particular, given an observation pair with stellar masses $\ms1$ and $\ms2$, we can use the SMHM relation to calculate the associated dark matter masses, $M_{h1}$ and $M_{h2}$, of the galaxy pair.
+            Next, calculate the virial radius of a dark matter halo with a virial mass equal to the sum of the dark matter masses of the pair.\footnote{We showed in~\chambe{} that the virial radius of a halo with mass given by the combined dark matter halo masses of the primary and secondary recovers the virial radius of the FoF group with approximately 98\% accuracy.}
+            Finally, determine the physical separation for each pair individually that corresponds to the scaled separation criteria adopted in the study.
 
-            One can determine an approximate virial radius for a given observed isolated pair by finding the virial radius of a subhalo with total mass $M_{h1}+M_{h2}$.\footnote{We showed in~\chambe{} that the virial radius of a halo with mass given by the combined dark matter halo masses of the primary and secondary recovers the virial radius of the FoF group with approximately 98\% accuracy.}
-            Since observed stellar masses already need to be computed in order to quantify the mass ratio of each pair, no further information from the observations are needed to create the mass and redshift evolving separation cuts.
+            The process of computing the dark matter halo masses of the observed galaxies typically introduces systematic error to an observational study.
+            However, we have already accounted for the error introduced by the spread in the SMHM relationship, since we use abundance matching to assign stellar masses to our dark matter halos and the SMHM relationship of~\cite{Moster2013} was constructed to account for observational stellar mass errors.
+            When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled in a similar fashion as performed in~\chambe{} to derive the spread associated with the abundance matching process.
 
-            As a concluding note, we want to remind readers that our results are explicitly for isolated pairs. 
-            One strength of our approach to only consider isolated pairs is that we have been able to study dynamic that are inherent to the pairs themselves, rather than those that are a product of the environment.
-            It is possible that low-mass and high-mass pair merger timescales in high density environments evolve completely differently than those found in our findings. 
+            As a concluding note, we remind readers that our results are explicitly for isolated pairs. 
+            One strength of this approach is that we have been able to study dynamics that are inherent to the pairs themselves, rather than those that are a product of the environment. 
+            Low-mass and high-mass pair merger timescales in high density environments may evolve differently than in our findings. 
             We leave the extension of this analysis to more high-density environments as the focus of future work.
@@
     Orbits of pairs, i.e.
-the separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo  first share a common FoF group (or have experienced `first infall') to either $z=0$ or the redshift at which the pair merges.
+the separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo first share a common FoF group (i.e., `first infall') to either $z=0$ or the redshift at which the pair merges.
 
     The sample consists of 22,213 unique low-mass orbits and 3,029 unique-high mass orbits, for which we quantify the merger fraction as a function of redshift.
-    From this sample, we calculate the merger timescales of low-mass and high-mass major pairs as a function of separation and separately as a function of redshift. 
-    Our goal is to identify the merger timescales of pairs on cosmologically motivated orbits, and to study and possibly identify differences between the redshift evolution of the merger timescales between low-mass and high-mass pairs.
+    We calculate the merger timescales of low-mass and high-mass major pairs as a function of separation and separately as a function of redshift. 
+    Our goal is to identify the merger timescales of pairs in a cosmological framework and to identify differences in the redshift evolution of the merger timescale between low-mass and high-mass pairs.
 
-    Additionally, motivated by our previous work on the pair fractions of low-mass and high-mass pairs, where we showed that the evolution of the pair fraction is sensitive to the separation criteria used to select the pairs, we seek to determine the impact of various selection criteria on the merger timescales.
-    Specifically, we look at two separate sets of selection criteria, one which selects pairs via a static physical separation cut, and the other which selects pairs based on a redshift and mass evolving separation cut.
-    This is especially important for studies that seek to study pair fractions, merger fractions, merger timescales, and merger rates of different galaxy mass scales and as a function of redshift.
+    Additionally, motivated by our previous work on the pair fractions of low-mass and high-mass pairs in \citet{Chamberlain2024} where we showed that the evolution of the pair fraction is sensitive to the separation criteria used to select the pairs, we seek to determine the corresponding impact of various selection criteria on the merger timescales of pairs.
+    Specifically, we look at two sets of selection criteria, one which selects pairs via a static physical separation cut, and the other which selects pairs based on a separation cut that evolves with redshift and mass.
+    This is especially important for studies that seek to study pair fractions, merger fractions, merger timescales, and merger rates at different mass scales and/or as a function of redshift.
 
@@
         % merger fractions
-        \item The merger fraction of low-mass and high-mass physically associated pairs is high ($>0.9$) at $z\gtrsim1.5$ (see Fig.~\ref{fig:fmerge}.
-However, the merger fraction declines rapidly to 0 as $z\to0$, due to the finite length of the simulation which artificially reduces the number of pairs with enough time left to merge at low redshifts.
+        \item The merger fraction of physically associated low-mass and high-mass pairs is high ($>0.9$) at $z\gtrsim1.5$ (see Fig.~\ref{fig:fmerge}).
+However, the merger fraction declines rapidly to zero as $z\to0$, due to the finite length of the simulation, which artificially reduces the number of pairs with enough time left to merge at low redshifts.
 % State whether it's a surprise or not that we see such high merger fractions, or if it is consistent w anything else, if you find something ?
@@
         \item The merger timescale of a pair at a given redshift increases with increasing pair separation.
-Additionally, for a given physical separation, high-mass pairs have a shorter merger timescale than low-mass pairs at the same redshift with the same separation.
-For example, at $z=2$, high-mass pairs at $100\kpc$ merge within $\sim 0-3 \Gyr$, while low-mass pairs at $100\kpc$ merge within $\sim 0.5-5\Gyr$ (see Fig.~\ref{fig:2dhist}).
+Additionally, for a given physical separation, high-mass pairs have a shorter merger timescale than low-mass pairs at the same redshift.
+For example, low-mass pairs at $z=1$ with separations $\rsep<100\,\kpc$ merge within $6\, \Gyr$, while high-mass pairs in the same separation range merge within $3\, \Gyr$ (see Fig.~\ref{fig:2dhist}).
         % Median merger timescale
-        \item The median merger timescale for the full sample low-mass and high-mass pairs that merge prior to $z=0$ declines with increasing redshift at a rate proportional to $1/H(z)$ for redshifts $z\gtrsim1$ (see Fig.~\ref{fig:timescales}).
-        The maximum merger timescale for a low-redshift pair ($z\lesssim1$) is constrained by the time remaining in the simulation for the pair to merge, which causes the median merger timescale to decline to 0 at $z=0$.
-        The decrease in the merger time at low redshifts is artificial and not representative of the true merger timescales that pairs at $z=0$ will have.
+        \item 
         The median merger timescale peaks around $z\sim0.6$ at $\sim2.1-2.4\Gyr$ for both low-mass and high-mass pairs.
+        At redshifts $z\gtrsim1$, the median merger timescale for the full sample of low-mass and high-mass pairs that merge prior to $z=0$ declines with increasing redshift at a rate proportional to $1/H(z)$ (see Fig.~\ref{fig:timescales}).
+        At $z\lesssim1$, the maximum merger timescale is constrained by the time remaining in the simulation for the pair to merge, which causes the median merger timescale to decline to zero at $z=0$.
+        The decrease in the merger time at low redshifts is artificial and not representative of the true merger timescales of pairs at low redshift.
         % The separation criteria effect 
         \item When pairs are selected via a scaled separation criteria, namely the pair separation scaled by the virial radius $\Rvir$ of the FoF group of the primary, low-mass and high-mass merger timescales are nearly identical at all redshifts.
-This holds for all scaled separations considered $[0.25,0.5,1,1.5,2,3]\,\Rvir$.
+This holds for all scaled separation criteria considered ($0.25,0.5,1,1.5,2,\mbox{and }3\,\Rvir$).
 At $z=[0, 1, 2, 3, 4]$, a scaled separation of $1\,\Rvir$ corresponds to an average physical separation of $[134, 85, 59, 43, 33]\kpc$ for low-mass pairs and $[348, 206, 134, 97, 76] \kpc$ for high-mass pairs.
         \item When pairs are selected via a physical separation criteria (one that does not vary with redshift or with mass), the median merger timescales of low-mass and high-mass pairs differ by up to $0.8\Gyr$.
-Thus, using the same merger timescale for low-mass and high-mass close pair samples selected via a physical separation cut will result in incorrect merger rate measurements.
+Thus, using the same merger timescale for low-mass and high-mass close pair samples selected via a physical separation cut will result in biased merger rate estimates.
     \end{itemize}
 
-    Studies have found that the merger rates of galaxy pairs varies with redshift and with the mass of the pair~\citep{Stewart2009,Hopkins2010a,RG2015}.
+    Studies have found that merger rates of galaxy pairs vary with redshift and with the mass of the pair~\citep{Stewart2009,Hopkins2010a,RG2015}.
     In our previous work, we found that the pair fraction of isolated pairs in TNG100 likewise evolves with redshift and mass~\citep{Chamberlain2024}. 
@@
 
-    These works together give a comprehensive framework for inferring merger rates via close pair fractions and merger timescales at redshifts from at least $z=0-6$, though likely higher redshift, and for quantifying the differences between low-mass and high-mass pairs.
+    These works together give a comprehensive framework for inferring merger rates via pair fractions and merger timescales at redshifts from at least $z=0-6$, and for quantifying the differences between low-mass and high-mass pairs.
     Indeed, separation selection criteria that scale with the mass and redshift of the target system are crucial for interpreting pair properties in a self-consistent way. 
-    In the future, observatories like Rubin Observatory, the Roman Space Telescope, JWST, and future ELTs, will detect an abundance of low-mass pairs at a wide range of redshifts, and our theoretical framework for interpretting these observations will be more critical than ever.
+    In the future, observatories such as the Rubin Observatory, the Roman Space Telescope, JWST, and future ELTs, will detect an abundance of low-mass pairs at a wide range of redshifts, and our theoretical framework for interpreting these observations will be more critical than ever.
 

### 2024-03-31 paper/main.tex
Updates from Overleaf

@@
     Tucson, AZ 85721, USA}
-\newcommand{\affuofu}{Department of Astronomy, University of Utah, Salt Lake City, UT 84112, USA}
+\newcommand{\affuofu}{Department of Physics and Astronomy, University of Utah, 115 South 1400 East, Salt Lake City, Utah 84112, USA}
 
@@
     % Statement
-    The merger timescales of isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) on cosmologically motivated orbits have not been studied in detail, though isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively.
+    The merger timescales of isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) on cosmologically motivated orbits have not yet been studied in detail, though isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) have been studied extensively.
+    % While the merger timescales of isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) in cosmological environments have been studied extensively, isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) have not yet been studied in detail.
     It is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems, however it is unclear if their merger timescales are similar, nor if they evolve similarly with redshift.
@@
     Applying these scaled separation criteria recovers equivalent merger timescales for both mass scales at all redshifts.
-    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts leads a difference in merger rates of up to $\sim1\Gyr$ between low- and high-mass pairs, particularly for $\rsep<150\kpc$.
+    Alternatively, static physical separation selections applied equivalently to all galaxy pairs at all redshifts leads to a difference in merger rates of up to $\sim1\Gyr$ between low- and high-mass pairs, particularly for $\rsep<150\kpc$.
     % Importance
@@
 
-In \chambe{}, 
+In \chambe{}, the
 abundance matching prescription was sampled 1000 times for each subhalo to account for the spread in the Stellar Mass -- Halo Mass (SMHM) relation. 
@@
 A primary or secondary subhalo can only be a member of one single pair at a given snapshot, such that a collection of $N$ pairs consists of $N$ unique primaries and $N$ unique secondaries.
-A subhalo can belong to different pairs at different redshifts. 
+A subhalo can belong to multiple different pairs at different redshifts. 
 For example, the primary of a pair that merges at $z=2$ can be selected with a different secondary at $z=1$, constituting a new pair. 
-More detail regarding the uniqueness of pairs and orbits is discussed in~\ref{sec:methods-unique}.
+More detail regarding the uniqueness of pairs and orbits is discussed in Sec.~\ref{sec:methods-unique}.
 
@@
 % might  note that this is discussed in more deets in discussion - I think Patton 2024 addresses this a bit?
-The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \subfind{} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift.
+
+The primary and secondary subhalos of each pair have a \texttt{SubfindID} in the \subfind{} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift. 
 The \sublink{} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.
@@
 We utilize the \texttt{DescendantID} field of the \sublink{} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when each merger occurs. 
-The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
-The \sublink{} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one.
+The \texttt{DescendantID} field provides the \texttt{SubhaloID}\footnote{Note that the \texttt{SubhaloID} is distinct from the \texttt{SubfindID}, and is unique for every subhalo in the merger trees. 
+The \sublink{} catalogs provide the associated \texttt{SubfindID} of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one.
 
@@
 
-For each merging pair, we define the `merger redshift' as the redshift that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
-For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place some time between $z=2$ and $z=1.9$, which is the next redshift corresponding to a snapshot in the simulation.
+For each merging pair, we define the \emph{merger redshift} as the redshift that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
+For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place between $z=2$ and $z=1.9$, which is the next redshift corresponding to a snapshot in the simulation.
 In this example, we take $z=1.9$ to be the merger redshift.
@@
 
-We extract orbits for all mergers and non-mergers in our base pair sample. 
+We extract orbits for all mergers and non-mergers in our base pair sample using the merger tree catalogs. 
 An orbit for a single pair is defined to be the physical separation between the primary and secondary subhalo as a function of redshift (or lookback time).
@@
 In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift.
-We use the definition of scaled separation $\rsep/\Rvir$ from~\cite{Chamberlain2024}, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group (also in $\kpc$).\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs.
+We use the definition of scaled separation $\rsep/\Rvir$ from~\cite{Chamberlain2024}, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Given by \texttt{Group\_R\_TopHat200} in the group catalogs.
 If the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.}
@@
 For some pairs, the secondary subhalos can experience first infall, then after one pericentric passage can return to large distances where they are temporarily assigned a different FoF group halo than that of their primary.
-In these cases, we use the full orbit beginning at first infall, including the sections of the orbit where the secondary is assigned to another FoF group temporarily,  through to merger as described in Section \ref{subsec:mergers}.
-This definition is more robust for considering the full interaction timescales of merging pairs.
+In these cases, we use the full orbit beginning at first infall, including the segments of the orbit where the secondary is assigned to another FoF group temporarily,  through to merger as described in Section \ref{subsec:mergers}.
+This definition is most robust for considering the full interaction timescales of merging pairs.
 %the post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same.
@@
 These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
-The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with many close encounters throughout, and three close passages in just the past $2\,\Gyr$.
+The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with three close passages in just the past $2\,\Gyr$.
 
@@
 Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of selected orbits.
+We further discuss the role of non-mergers and fly-bys on our results in Sec.~\ref{sec:disc-nonmergers}.
 
@@
 \label{sec:methods-unique}
-Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts after infall, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists.
-However, we only want to keep a single instance of any given orbit in our catalog, since the orbits with the longest timescales will be selected many times which would skew our data to longer merger timescales.
+Since we collect the orbit for each pair at all redshifts after infall, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists.
+However, it is only necessary to keep a single instance of any given orbit in our catalog to avoid skewing our data artificially to longer merger timescales.
+%since the orbits with the longest timescales will be selected many times which would skew our data to longer merger timescales.
 
 To distinguish the collection of unique orbits, each pair is assigned a `pairkey' while constructing their orbit. 
-The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the \sublink{} catalogs, and is unique for each pair of halos. 
+The pairkey is created by concatenating the earliest \texttt{SubhaloID} of the primary and secondary subhalo from the \sublink{} catalogs, and is unique for each pair of halos. 
 After each pair is assigned a unique pairkey, we only keep one instance of an orbit per pair to avoid double/multi-counting in our orbit catalog.
@@
 For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
-In this case, both orbits (of the original low-mass pair and the new pair which includes the primary subhalo from the previous merger) are retained in the orbit catalog, since each orbit is unique.
+In this case, both orbits (of the original low-mass pair and the new pair which includes the primary subhalo from the previous merger) are retained in the orbit catalog.%, since each orbit is unique.
 
-The total number of pairs (including double counting the same pair at multiple redshifts) is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
+The total number of orbits, before removing the redundant orbits,  %(including double counting the same pair at multiple redshifts) 
+is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
 However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits that each correspond to a unique pair of subhalos.
+The collection of all unique orbits constitutes our orbit sample and is the dataset that will be used for the remainder of this analysis.
 
-The collection of all unique orbits constitutes our orbit sample and is the dataset that will be used for the remainder of the paper.
+Note that we choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing.
+However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.
 
@@
     \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
-    \caption{The fraction of low-mass (green) and high-mass (pink) post-infall orbits that result in a merger before z=0 as a function of redshift. 
+    \caption{The fraction of low-mass (green) and high-mass (pink) orbits that merge before z=0 shown as a function of redshift. 
     The horizontal line shows a merger fraction of 0.9, or 90\%. 
@@
     The sharp decline of the merger fraction to zero at low redshift is a non-physical feature of the simulation ending at $z=0$.
-    Only orbits with sufficiently short merger timescales will be able to merge at these redshifts.}
+    Only orbits with sufficiently short merger timescales will merge at redshifts $z<\sim1$.}
     \label{fig:fmerge}
@@
 \subsection{Number of Pairs}\label{sec:pairprops-num}
-We calculate the total number of pairs by selecting all pairs at a given redshift, including merger and non-merger pairs.
+The total number of pairs at a given redshift is equal to the number of orbits at that redshift, including both merger and non-merger pairs. 
+%We calculate the total number of pairs by selecting all pairs at a given redshift, including merger and non-merger pairs.
 A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.
@@
 As subhalos experience first infall into a FoF group, the number of pairs increases. 
-On the other hand, mergers cause a decrease in the number of pairs.
+On the other hand, mergers simultaneously lead to a decrease in the number of pairs.
 % The difference between high and low mass
 The number of low-mass pairs increases from $\sim600$ pairs at $z=6$ to $\sim6,600$ at $z=2$. 
-The decrease in the number of pairs after $z=2$ means that the number of low-mass pairs that are merging at each redshift is larger than the rate at which pairs are being added to the sample.
+The decrease in the number of pairs after $z=2$ means that the number of low-mass pairs that are merging at each redshift is larger than the rate at which pairs are added to the sample.
 
@@
     As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.
-    Note, this definition of the merger fraction differs from that typically used in observational studies, where merger fractions are computed from close pair fractions with a correction term.
+    Note, this definition of the merger fraction differs from that typically used in observational studies, where merger fractions are computed from close pair fractions with a correction term~\cite[e.g.,][]{Ventou2019}.
 
@@
     \caption{The distribution of merger timescales for low-mass (top) and high-mass (bottom) pairs as a function of physical separation at $z=0.5,1,1.5,2,3,4,5\mbox{and }6$. 
-    The colorbars show the percentage of the total number of pairs at each redshift (given in the bottom right corner of each panel) that are in a given bin.
+    The colorbars show the percentage of the total number of pairs at each redshift (given in the bottom right corner of each panel) that are in a given separation bin.
     In each of the first panels, the grey horizontal dashed line shows the time remaining in the simulation, above which there are no mergers. 
@@
     We find:
-    1.) a positive correlation between separation and the time until merger at each selected redshift -- the lowest separation pairs tend to have the shortest merger timescales; 
-    2.) given a redshift and separation, the low-mass merger timescale is longer than the high-mass merger timescale, i.e.
+    i.) a positive correlation between separation and the time until merger at each selected redshift -- the lowest separation pairs tend to have the shortest merger timescales; 
+    ii.) given a redshift and separation, the low-mass merger timescale is longer than the high-mass merger timescale, i.e.
 low-mass pairs at $z=1$ with separations $\rsep<100\,\kpc$ merge within $6\, \Gyr$, while high-mass pairs in the same separation range merge within $3\, \Gyr$;
-    3.) merger timescales are longer for lower redshift pairs; 
-    4.) the spread of merger timescales and separations increase with decreasing redshift.
-    %
-    This plot can be used to get estimates for the merger timescale of an isolated pair at a given redshift. 
-    For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-5\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
-    On the other hand, a high-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-2\,\Gyr$, with a most likely merger timescale of around $0-1\Gyr$.
-    }
+    iii.) merger timescales are longer for lower redshift pairs; 
+    iv.) the spread of merger timescales and separations increase with decreasing redshift. }
     \label{fig:2dhist}
@@
 \section{Results: The Mass and Redshift Dependence of Merger Timescales}\label{sec:results}
-    Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
+    Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in sample. 
     The merger timescale of a pair is defined to be the amount of time that elapses between the redshift at which pairs are selected and the merger redshift.
     Only the merging pairs will be considered for the remainder of the analysis. 
-\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
-    Note that we choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing.
-However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.
+    %\footnote{Since our orbits are only defined after infall, the merger timescales will likewise only be calculated for pairs that are post-infall at the given separation.}
 
@@
 
-    We find that the time until merger for low-mass and high-mass pairs is positively correlated with the separation of the pair at each redshift. 
+    We find that the merger timescale for low-mass and high-mass pairs is positively correlated with the separation of the pair at a given redshift. 
+    The merger timescale also increases with decreasing redshift for a given separation.
     Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$. 
@@
     In addition, the spreads in the merger timescales and pair separations are smaller at higher redshift, for both low- and high-mass pairs. 
-    For example, the spread of merger timescales for low-mass pairs at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$. 
+    For example, the spread of merger timescales for low-mass pairs at $z=1$ (4273 pairs) goes from $0-8\,\Gyr$, while pairs at $z=4$ (2218 pairs) have merger timescales between $0-3\,\Gyr$. 
     Likewise, the spread of separations at $z=1$ is $0-200\kpc$, but at $z=4$ the spread is smaller, between $10-125\kpc$. 
@@
     This is due to the growth of halos over time, and thus an increasing virial radius for FoF groups at lower redshifts.
+
+    Fig.~\ref{fig:2dhist} can be used to estimate the merger timescale of an isolated pair at a given redshift. 
+    For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-5\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
+    On the other hand, a high-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-2\,\Gyr$, with a most likely merger timescale of around $0-1\Gyr$.
 
@@
     \subsubsection{Full sample}
-        We calculate the merger timescales for pairs at each redshift in the simulation from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
+        We calculate the merger timescales for pairs at each redshift, and quantify the median and spread on the merger timescale as a function of redshift in Fig.~\ref{fig:timescales}. 
         We include the full catalog of merging orbits at each redshift, only excluding separations $\rsep<10\kpc$ to limit the impact of subhalos becoming indistinguishable in the \subfind{} catalogs.% to calculate the merger timescale.
@@
 
-        Figure~\ref{fig:timescales} presents the median merger timescales for merging pairs from $z=0-6$. 
-        The 1st and 3rd quartiles are indicated with shaded regions. 
+        Figure~\ref{fig:timescales} also shows
+        the 1st and 3rd quartiles are indicated with shaded regions. 
         % low vs high
@@
         It is particularly interesting as well that low-mass and high-mass pairs have the same merger timescales \textit{and} follow the $1/H(z)$ redshift scaling. 
-        We believe this is due to the dominance of dynamical friction in determining the redshift behavior of merger timescales. 
-        % Thus, the primary mechanics that determine merger timescales are likely dynamical friction
+        %We believe this is due to the 
+        % Thus, the primary mechanism that determines merger timescales is gravity.... and not dynamical friction? is this true?
 
@@
         Separation selected samples have a lower separation criteria of $10\,\kpc$.
-        This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates~\citep[see e.g.][and observational studies cited therein]{Lotz2011,Snyder2017,Besla2018,Chamberlain2024}.
+        This lower separation criteria is also commonly applied to observationally selected pairs in studies of merger fractions and merger rates~\citep[see e.g.][and observational studies cited therein]{Lotz2011,Snyder2017,Besla2018}.
 
-        The first set of separation criteria will select only pairs at a given redshift that have physical 3D separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, yielding six pair subsamples.
+        The first set of separation criteria selects only pairs at a given redshift that have physical 3D separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$, yielding six pair subsamples.
         Pairs that are selected via the $\rsep<50\kpc$ separation cut will be included in the $\rsep<70\kpc$ subsample, and so on.
         These separation criteria do not vary as a function of time or mass, and are applied equivalently to the low-mass and high-mass samples at all redshifts. 
-        The separation criteria are applied independently at each redshift.
+        %The separation criteria are applied independently at each redshift.
         For example, an orbit in the separation bin $<50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts.
@@
         % figure description 
-        The top panel of Fig.~\ref{fig:timescales-sep} shows the time until merger versus redshift for each of the low-mass (green) and high-mass (pink) pair subsamples with separations less than the physical separation listed above each panel.
+        The top panel of Fig.~\ref{fig:timescales-sep} shows the merger timescale versus redshift for each of the low-mass (green) and high-mass (pink) pair subsamples with separations less than the physical separation listed above each panel.
         The same Hubble time redshift scaling from Sec.~\ref{sec:results-timevredshift} is shown by the thin black line in each panel.
@@
         In addition, all subsamples show the same decline of mean merger time as $z\to0$, as discussed in the previous section. 
-        These traits were also present in Fig.~\ref{fig:timescales}, meaning they are features that are roughly independent of separation criteria.
+        These traits are shown in Fig.~\ref{fig:timescales}, meaning they are features that are independent of separation criteria.
 
@@
         The offset between the low-mass and high-mass merger timescale decreases for the largest separation cut of $\rsep<300\kpc$. 
-        In the top rightmost panel, the median merger times converge for early all redshifts, similar to the merger timescales from the full sample shown in Fig. \ref{fig:timescales}.
+        In the top rightmost panel, the median merger times converge for nearly all redshifts, similar to the merger timescales from the full sample shown in Fig. \ref{fig:timescales}.
 
@@
 
-        The median merger timescale for low-mass pairs does not change significantly for any separation cuts above $150\,\kpc$. 
-        On the other hand, high-mass merger timescale tends to increase for each larger separation cut. 
+        The median merger timescale for low-mass pairs does not change significantly for any separation cuts above $150\,\kpc$, which is larger than the average virial radius of low-mass systems at all redshifts.
+        On the other hand, high-mass merger timescales tend to increase as separation increases.
         % This is because selecting all pairs with $\rsep<200\,\kpc$ in a low-mass system is including a larger volume of the FoF halo, and thus a more complete selection of the full sample.
-        High-mass pairs tend to have higher separations than low-mass pairs, as shown in~\ref{sec:results-timevsep}. 
+        High-mass pairs tend to have higher separations than low-mass pairs, as shown in Sec.~\ref{sec:results-timevsep}, due to the larger size of the high-mass subhalos. 
     \begin{figure*}[htb]
@@
         % low mass vs high mass
-        The median time until merger at a given redshift increases .
+        The median time until merger at a given redshift increases.
         The low-mass and high-mass merger timescales increase for separation criteria that include a larger fraction of the virial radius, and thus larger separation pairs, but recover the behavior of the full sample by a separation cut of $2\,\Rvir$.
@@
         This could be the case if the velocity at which the secondary subhalo enters the primary's FoF group becomes increasingly large compared to the virial velocity\footnote{The virial velocity is the value of the circular velocity at the virial radius.}, resulting in fewer bound systems.
-        This will be studied in future work.
+        This phenomena will be tested in future work.
 
@@
             These findings are consistent with predictions of the galaxy merger rate from semi-empirical models~\citep{Stewart2009,Hopkins2010a}, and with merger rates derived from observational merger fraction measurements calibrated by the corrected observability timescales of \citet{Lotz2011}.
+
             Additionally, \citet{RG2015} find that the major merger rate at $z\sim0.1$ increases with increasing descendant subhalo mass.
@@
         \subsubsection{Implementing Scaled Separation Pair Selection Criteria in Future Studies}
-            In our previous work, we studied the impact of different sets of separation criteria on the recovered pair fractions of low-mass and high-mass pairs, and found that features of the redshift evolution of the pair fractions can only be distinguished by employing scaled separation criteria~\citep{Chamberlain2024}.
+            In our previous work, we studied the impact of different separation criteria on the recovered pair fractions of low-mass and high-mass pairs, and found that features of the redshift evolution of the pair fractions can only be distinguished by employing scaled separation criteria~\citep{Chamberlain2024}.
             In the present work, we explore how these same set of selection criteria impact the recovered merger timescales. 
@@
             The application of a scaled separation as demonstrated here can be straightforwardly applied to observational studies. 
-            Since estimates of a pair's stellar masses are needed to quantify the mass ratio of the pair (which is commonly computed for merger rate studies), no further information from the observations is needed to create the mass and redshift evolving separation cuts.
+            Since estimates of a pair's stellar masses are needed to quantify the mass ratio of the pair (which is commonly computed for merger rate studies), no further observational information is needed to develop mass and redshift evolving separation cuts.
+
+            We provide a step by step example of this application as follows.
 First, compute the associated dark matter masses from the observed stellar masses of the pair using a SMHM relation~\citep[in this study, we employ that of][]{Moster2013}.
@@
             The process of computing the dark matter halo masses of the observed galaxies typically introduces systematic error to an observational study.
-            However, we have already accounted for the error introduced by the spread in the SMHM relationship, since we use abundance matching to assign stellar masses to our dark matter halos and the SMHM relationship of~\cite{Moster2013} was constructed to account for observational stellar mass errors.
+            However, we have already accounted for the error introduced by the spread in the SMHM relationship, since we use abundance matching to assign stellar masses to our dark matter halos.
+Furthermore, the SMHM relationship of~\cite{Moster2013} was constructed to account for observational stellar mass errors.
             When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled in a similar fashion as performed in~\chambe{} to derive the spread associated with the abundance matching process.
@@
 
-% ** 
-% \appendix
-% adfadfa
-
 \bibliography{refs}{}

### 2024-03-31 paper/main.tex
Updates from Overleaf

@@
             \centering
-            \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale_mod.png}
+            \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale.png}
             \caption{The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs in our merger sample. 
@@
             The thick dotted black line at low redshifts shows the time to $z=0$ (the lookback time) as a function of redshift, which sets the upper bound for the merger timescale that a merging pair can have.
-            The thin black line shows the Hubble time as a function of redshift multiplied by a constant $XX$, which approximates the median time until merger from $z=6$ to $z\sim1$.
+            The thin black line shows the Hubble time as a function of redshift multiplied by a constant $\beta=0.35$, which approximates the median time until merger from $z=6$ to $z\sim1$.
             %
-            The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.\kc{update}}
+            The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
             \label{fig:timescales}
@@
 
-        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by XXX. 
+        In Fig.~\ref{fig:timescales}, the black dashed line shows the Hubble time $1/H(z)$ multiplied by $\beta=0.35$. 
         We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the specific values of the merger timescale.
@@
         \centering
-        \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger.png}
+        \includegraphics[width=\textwidth]{plots/bet-on-it/8_timescale_sepcut.png}
         \caption{(Top) The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $50,70,100,150,200,\mbox{and }300\,\kpc$ from left to right.
@@
         % 
-        The thin black line is the same as shown in Fig.~\ref{fig:timescales}, and goes as the Hubble time $1/H(z)$.
+        The thin black line is the same as shown in Fig.~\ref{fig:timescales}, and goes as the Hubble time $1/H(z)$ multiplied by 0.35 (see Sec.~\ref{sec:results-timevredshift} for additional details).
         The time until merger increases from $z=4$ to $z\sim0.5-0.75$, at which point the time until merger decreases to zero since all low redshift mergers must have short merger timescales to merge before the end of the simulation at $z=0$.
-        \kc{H(z), "merger timescale"}
         }

### 2024-05-13 paper/main.tex
Updates from Overleaf

@@
 \newcommand{\affuofu}{Department of Physics and Astronomy, University of Utah, 115 South 1400 East, Salt Lake City, Utah 84112, USA}
+\newcommand{\affuva}{Department of Astronomy, University of Virginia, 530 McCormick Road, Charlottesville, VA 22904, USA}
 
@@
 
+\author[0000-0002-5653-0786]{Paul Torrey}
+\affiliation{\affuva}
+
 \author{others}
@@
     % While the merger timescales of isolated high-mass pairs ($\rm 5\times10^9<M_*<10^{11}\,\Msun$) in cosmological environments have been studied extensively, isolated low-mass pairs ($\rm 10^8<M_*<5\times10^9\,\Msun$) have not yet been studied in detail.
-    It is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems, however it is unclear if their merger timescales are similar, nor if they evolve similarly with redshift.
+    It is common to apply the same separation criteria and expected merger timescales of high-mass pairs to low-mass systems, however it is unclear if their merger timescales are similar, 
+    %nor
+    % GB edited
+    or
+    if they evolve similarly with redshift. 
     % Our solution
@@
 \section{Methodology} \label{sec:methods}
-%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
-We utilize the group catalogs produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009} and the merger tree catalogs generated by the \sublink{} algorithm~\citep{RG2015} from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
+The IllustrisTNG simulation suite~\citep{TNG1,TNG2,TNG3,TNG4,TNG5} is a set of large volume dark-matter-only and full magnetohydrodynamical cosmological simulations consistent with the \textit{Planck 2015} \lcdm{} cosmology~\citep{Planck2015}.
+
+Following~\chambe{}, we use TNG100-~1 (hereafter TNG100), which is the highest resolution full physics run of the ($100\Mpc)^3$ volume.
+In particular, we utilize the group catalogs produced by the \subfind{} algorithm~\citep{Springel2001b,Dolag2009} and the merger tree catalogs generated by the \sublink{} algorithm~\citep{RG2015}.
 The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).
@@
 Our sample consists of major low-mass and high-mass pairs (stellar mass ratio $1/4 \leq M_{*2}/M_{*1}\leq 1$) that are isolated, but physically associated, as in \citet{Chamberlain2024}.
-From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.
+From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales from $z=0-6$.
 
@@
 \end{align*}
-Requiring pairs to belong to the same FoF group ensures that the pairs are isolated from other massive nearby systems that could perturb the dynamical state of the pair.
+Requiring pairs to belong to the same FoF group ensures that the pairs are distant from other massive nearby systems that could perturb the dynamical state of the pair.  
+In addition, for each pair, we calculated the Hill Radius of every FoF group with a higher mass within $10Mpc$ of the pair, as a proxy to discern whether the pair lies outside the gravitational sphere of influence of these more massive groups.
+We found that over 99\% of our pairs at $z=0$ are more distant than two times the Hill Radius of each FoF group.
 
@@
     \includegraphics[width=\columnwidth]{plots/bet-on-it/5_exampleorbits.png}
+    % [Example Orbits for Galaxy Pairs in TNG100 Selected at $z=1.5$]
     \caption{Selection of example orbits of low-mass major pairs that pass the pair selection criteria at $z=1.5$, showing the separation between the primary and secondary as a function of lookback time.
@@
 % We collect the following data for each pair at each redshift where the primary and secondary are both defined: the virial mass of the primary's FoF group \texttt{Group\_M\_TopHat200}, the virial radius of the primary's FoF group , the physical separation (in $\kpc$), and the scaled separation (dimensionless).
+
+Note that we choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing.
+However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.
 
@@
 
-Note that we choose not to interpolate the orbits to get a more precise determination of the merger timescale, as the spread on the merger timescales in our study will be dominated by the variation of orbital configurations, rather than the uncertainty due to snapshot spacing.
-However,~\citet{Patton2024} recently showed that for studies needing more precise determination of, e.g., pericenter passages or merger times, their 6D interpolation scheme can improve timing uncertainties from $\pm80\Myr$ to $\pm3.3\Myr$.
-
 \section{Pair Sample Properties}\label{sec:pairprops}
@@
     \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
+    % [Number of Unique Pairs in Orbit Catalog]
     \caption{The number of unique low-mass (green) and high-mass (pink) pairs in our orbit catalog as a function of redshift.
@@
     The solid lines show the total number of pairs at a given redshift (including non-mergers and mergers), while dashed lines show the number of pairs at a given redshift that will merge before $z=0$.
-    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.\kc{"total"}
+    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
     }
@@
     \includegraphics[width=\columnwidth]{plots/bet-on-it/6_mergerfraction.png}
+    %[Merger Fraction of Pairs as a Function of Redshift]
     \caption{The fraction of low-mass (green) and high-mass (pink) orbits that merge before z=0 shown as a function of redshift. 
@@
 The total number of pairs at a given redshift is equal to the number of orbits at that redshift, including both merger and non-merger pairs. 
-%We calculate the total number of pairs by selecting all pairs at a given redshift, including merger and non-merger pairs.
 A single, non-merging pair with first infall at $z=3$ will contribute to the number of pairs at all redshifts from $z=0-3$.
@@
     \begin{center}
-    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/8_2Dhist.png}
+    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/10_2Dhist_fits.png}
+    %[Distribution of Time Until Merger as a Function of Separation at Redshifts $z=0.5,1,2,3,4,5,6$]
     \caption{The distribution of merger timescales for low-mass (top) and high-mass (bottom) pairs as a function of physical separation at $z=0.5,1,1.5,2,3,4,5\mbox{and }6$. 
@@
     The merger timescale also increases with decreasing redshift for a given separation.
-    Pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$. 
+    For example, pairs at $z=4$ with separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$. 
     For the same physical separation, the merger timescale of high-mass pairs is shorter than for low-mass at each redshift. 
@@
     On the other hand, a high-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-2\,\Gyr$, with a most likely merger timescale of around $0-1\Gyr$.
+
+    \input{paper/table-slopes}
 
@@
             \includegraphics[width=\columnwidth]{plots/bet-on-it/8_timescale.png}
+            % [Median Merger Timescale of Low-mass and High-mass Pairs as a Function of Redshift]
             \caption{The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs in our merger sample. 
@@
         We did not perform a fit for this multiplicative constant, since we are interested in investigating the behavior of the merger timescale as a function of redshift rather than the specific values of the merger timescale.
-        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=\sim1-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation, from~\citet{Planck2018}.
+        The redshift evolution of our median merger timescales is consistent with the scaling with $1/H(z)$ between $z=\sim1-6$.\footnote{$H(z)$ is calculated using the same cosmology as the TNG100 simulation, from~\citet{Planck2015}.
 Specifically, $\Omega_M=0.31$ and $\Omega_{\Lambda}=0.69$.}
@@
         \includegraphics[width=\textwidth]{plots/bet-on-it/8_timescale_sepcut.png}
+        % [Median Merger Timescale as a Function of Redshift for Pairs Selected by Physical and Scaled Separation Criteria]
         \caption{(Top) The median merger timescale as a function of redshift for low-mass (green) and high-mass (pink) pairs with 3D physical separations greater than $10\,\kpc$ and less than $50,70,100,150,200,\mbox{and }300\,\kpc$ from left to right.
@@
             Using their projected separation vs velocity analysis of merger probabilities, they also develop a weighting scheme to determine the likelihood of merger as a function of projected separation and velocity, which provides a way to more accurately calibrate merger fractions from close pair fractions.
-            While their study considers the impact of stellar mass and redshift on the probability of merger, their suggested pair selection criteria do not vary with mass or redshift, which we have shown is crucial for interpretting and comparing pair fractions and merger timescales.
+            While their study considers the impact of stellar mass and redshift on the probability of merger, their suggested pair selection criteria do not vary with mass or redshift, which we have shown is crucial for interpreting and comparing pair fractions and merger timescales.
 
@@
             % Although their findings come from an approximation for observability timescales calibrated from merger timescales,  rather than merger timescales specifically, they find that projected separation criteria of $5\leq r_{\rm proj}<30\kpc$ underestimates the merger rate for $z=0-5$.
-
-            % \textit{RG2015} 
-            % the reason I want to talk about the RG15 paper is because its an example of trying to put galaxy merger rates on the same plots, and using the same selection criteria at high and low masses may have also resulted in such a discrepancy, IN ADDITION to the problems that low mass morphology studies face when trying to use morphological signatures for low-mass merger fraction studies.
 
@@
             These results imply that the timescales used to convert observational merger fractions (determined either by weighted close pair fractions or by morphologically selected pairs) are sensitive to the mass ranges and redshifts at which they are valid, and attempting to make comparisons between theoretical predictions and observational measurements can lead to seemingly incompatible results.
-            % Since our observational samples themselves will be able to also self-consistently apply merger timescales to close pair fractions using our suggested scaled separation criteria.
             Our work provides a unique framework that will allow for more robust comparisons of theoretical and observational samples.
@@
 First, compute the associated dark matter masses from the observed stellar masses of the pair using a SMHM relation~\citep[in this study, we employ that of][]{Moster2013}.
-            % In particular, given an observation pair with stellar masses $\ms1$ and $\ms2$, we can use the SMHM relation to calculate the associated dark matter masses, $M_{h1}$ and $M_{h2}$, of the galaxy pair.
-            Next, calculate the virial radius of a dark matter halo with a virial mass equal to the sum of the dark matter masses of the pair.\footnote{We showed in~\chambe{} that the virial radius of a halo with mass given by the combined dark matter halo masses of the primary and secondary recovers the virial radius of the FoF group with approximately 98\% accuracy.}
+            % Paul's comment:
+            % you could make this more straight forward, if you want.
+Specifically, there is a M_* to R_vir correction that exists in TNG, and you could parameterize and "give" this to the observers.
+            Next, calculate the virial radius of a dark matter halo with a virial mass equal to the sum of the dark matter masses of the pair.
+            \footnote{We showed in~\chambe{} that the virial radius of a halo with mass given by the combined dark matter halo masses of the primary and secondary recovers the virial radius of the FoF group with approximately 98\% accuracy.
+The virial radius can be calculated from
+            \begin{equation}\label{eq}
+                \Rvir = \sqrt[3]{\frac{3\rm M_{vir}}{4\pi \Delta_{c}\rho_{c}}},
+            \end{equation} 
+            where $\rho_c$ is the critical density of the Universe, and $\Delta_{c}$ is the overdensity constant~\citep[see][]{Binney2008}.}
             Finally, determine the physical separation for each pair individually that corresponds to the scaled separation criteria adopted in the study.
@@
             The process of computing the dark matter halo masses of the observed galaxies typically introduces systematic error to an observational study.
-            However, we have already accounted for the error introduced by the spread in the SMHM relationship, since we use abundance matching to assign stellar masses to our dark matter halos.
+            When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled, i.e.
+by computing many realizations of the dark matter mass of each galaxy, in a similar fashion as performed in~\chambe{}, to derive the associated spread introduced by the abundance matching process.
+
+            % However, we have already accounted for the error introduced by the spread in the SMHM relationship, since we use abundance matching to assign stellar masses to our dark matter halos.
 Furthermore, the SMHM relationship of~\cite{Moster2013} was constructed to account for observational stellar mass errors.
-            When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled in a similar fashion as performed in~\chambe{} to derive the spread associated with the abundance matching process.
+            % When converting from stellar mass to dark matter halo mass, the SMHM relationship can be sampled in a similar fashion as performed in~\chambe{} to derive the spread associated with the abundance matching process.
 
-            As a concluding note, we remind readers that our results are explicitly for isolated pairs. 
+            As a concluding note, we remind readers that our results are explicitly for isolated pairs, as outlined in Sec.~\ref{sec:methods}. 
             One strength of this approach is that we have been able to study dynamics that are inherent to the pairs themselves, rather than those that are a product of the environment. 
@@
     % what we did and how
-    In this study, we construct a sample of the orbits of low-mass ($10^8 <\ms{}<5\times10^{9}\Msun$) and high-mass ($5\times10^{9}<\ms{}<10^{11}\Msun$) major pairs (stellar mass ratio $> 1:4$) from $z=0-6$ in the TNG100 simulation.
+    In this study, we construct a sample of the orbits of isolated low-mass ($10^8 <\ms{}<5\times10^{9}\Msun$) and high-mass ($5\times10^{9}<\ms{}<10^{11}\Msun$) major pairs (stellar mass ratio $> 1:4$) from $z=0-6$ in the TNG100 simulation.
     Orbits of pairs, i.e.
-the separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo first share a common FoF group (i.e., `first infall') to either $z=0$ or the redshift at which the pair merges.
+the 3D physical separation between the pair as a function of time, are defined from the redshift at which the primary and secondary subhalo first share a common FoF group (i.e., `first infall') to either $z=0$ or the redshift at which the pair merges.
 
